### 2010-02-05 trunk/docs/www2010/paper.tex
adding all the images for the paper

% This is "www2010-sample.tex" copied from "www2005-sample.tex" V1.2 January 26 2004
% This file should be compiled with V1.4 of "www2010-submission.class"
%
% This example file demonstrates the use of the 'www2010-submission.cls'
% V1.4 LaTeX2e document class file. It is for those submitting
% articles to the WWW'04 Conference WHO DO NOT WISH TO 
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'www2010-submission.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V1.4) produces:
%       1) NO Permission Statement
%       2) WWW'04-specific conference (location) information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Julie Goetz (goetz@acm.org) or Adrienne Griscti (griscti@acm.org)
%
% Technical questions only to
% Gerald Murray (murray@acm.org)
% ===============================================================
%
% For tracking purposes - this is V1.2 - January 26 2004
\documentclass{www2010-submission}

\begin{document}
%
\title{QueryMed: An Intuitive Federated SPARQL Query Builder for Biomedical RDF Data}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the "boxing"
% and alignment of the authors under the title, and to add
% a section for authors number 4 through n.
%
% Up to the first three authors are aligned under the title;
% use the \alignauthor commands below to handle those names
% and affiliations. Add names, affiliations, addresses for
% additional authors as the argument to \additionalauthors;
% these will be set for you without further effort on your
% part as the last section in the body of your article BEFORE
% References or any Appendices.

\numberofauthors{2}
%
% Put no more than the first THREE authors in the \author command

% NOTE: All authors should be on the first page. For instructions
% for more than 3 authors, see:
% http://www.acm.org/sigs/pubs/proceed/sigfaq.htm#a18

\author{
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
%% e-mail address with \email.
\alignauthor Oshani Seneviratne\\
       \affaddr{Massachusetts Institute of Technology}\\
       \affaddr{Cambridge, MA}\\
       \affaddr{USA}\\
       \email{oshani@csail.mit.edu}
\alignauthor Rachel Sealfon\\
       \affaddr{Massachusetts Institute of Technology}\\
       \affaddr{Cambridge, MA}\\
       \affaddr{USA}\\
       \email{rsealfon@csail.mit.edu}
}
\date{15 Feb 2010}

\maketitle
\begin{abstract}
We have developed an open-source SPARQL query builder and result set visualizer for biomedical data, QueryMed, that allows end users to easily construct and run translational medicine queries across multiple data sources. QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, runs federated queries across multiple SPARQL endpoints, and is designed to be accessible to users who do not know the structure of the underlying ontologies or the SPARQL query language. The system allows users to select the data sources that they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.  Users can add additional data sources if they are interested in querying endpoints that are not in the default list. After retrieval of the initial result set, query results can be filtered to improve their relevance. As an advanced querying feature, the system also allows the user to exploit the underlying structure of the RDF data to improve query results. 
\end{abstract}

% A category with only the three required fields
\category{J.3}{Life and Medical Sciences}{Computer Applications}
\category{H.3.3}{Information Search and Retrieval}{Information Systems}

\keywords{Biomedical Ontologies, SPARQL, Query Federation, Query Building, Semantic Web, User Interfaces}

\section{Introduction}

RDF data in the semantic web, like the data in relational databases, is modeled in a structured format that can be queried using a standardized query language. The biomedical domain is among the early successes of the semantic web, due to the rapidity with which the biomedical community has made its data available in RDF triple stores. However, although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.

For example, a physician may know her patient’s personal information, symptoms, current medications, and genotype. She may wish to determine the patient’s treatment plan and identify clinical trials for which the patient is eligible.  Although the physician has a single question--based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?--there is no single data source that the physician can use to answer this question. The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and GO [Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, Gene Ontology]. Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.  For example, to address her question, the physician might be interested in information on “coronary artery disease” (which she thinks is the cause of her patient’s symptoms). The information about the disease can be found in Diseasome. She might also want to know the set of drugs in DailyMed and Drugbank that can be used to treat coronary artery disease and that also will not react adversely with the patient’s current medications.  Information on clinical trials available for her patient can be found in LinkedCT.  Due to the large number of databases  that the physician needs to query in order to find an answer to her single question, she is likely to find useful a system that can automatically run queries over multiple data sources. Also, the physician may not know SPARQL query syntax, the location of the SPARQL endpoints, or the structure of the relevant ontologies. She is likely to want an intuitive way to query and to display the query result.  Developing intuitive ways to query multiple data sources and display results is both an important and a challenging problem.

This paper is organized as follows: Section 2 provides background information on the semantic web and its relevance for the biomedical domain. Section 3 describes our system. Section 4 discusses related work and illustrates how QueryMed differs from previous systems. Finally, section 5 outlines future work and summarizes the contributions of our system.

\section{Background}

\subsection{Biomedical Ontologies and Semantic Web}

The semantic web can be viewed as global database system for the data available on the world wide web. Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using a standardized query language, SPARQL [Berners-Lee, 2001]. The addition of structure to web data is intended to allow inferences to be automatically made by intelligent agents drawing on data from multiple sources.

RDF data is described by a directed, labeled graph, with relationships among data items represented  in triples containing a subject, an object and a predicate.  For example, we might want to express the concept, "Sam is a professor for 6.830." The corresponding triple expresses the relationship "is a professor for" between two entities, “Sam” and “6.830”. The relationship "is a professor for" is known as the predicate of the triple; "Sam" is the subject while "6.830" is the object.  In order for this statement to be machine readable, each of the entities in the statement ("Sam," "is a professor for," and "6.830") will be described by a Uniform Resource Identifier (URI) [Manola, 2004].  The Web Ontology Language (OWL) is a language that permits the definition of ontologies, allowing additional properties and constraints on relationships among data to be represented [Smith, 2004]. Some parallels between relational databases and RDF data on the semantic web are illustrated in table 1 [Berners-Lee, 1998].

Although most web data are not currently available in semantic web formats, the biomedical community has been extremely proactive in making structured biomedical data available in RDF. Thus, the biomedical knowledge domain has been an early success of the semantic web [Yip, 2009]. Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.   Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists, and these semantic web resources represent valuable standardized repositories of information that can be automatically mined for biological knowledge.  The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain [HCLSIG].

However, although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources of maximal use to physicians, patients, and life scientists.  One challenge is constructing systems that allow end users to run intuitive queries on biomedical data.  Users of biomedical resources are likely to have extensive domain knowledge, but be unfamiliar with the SPARQL query language syntax and with the structure of biomedical ontologies. It is important to design user-friendly systems that allow these users to take advantage of the wealth of structured biological knowledge available on the semantic web.  Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data is often distributed among many sources [Pasquier, 2008].

\subsection{SPARQL Query Language}

A SPARQL query performs a pattern-matching search on the underlying RDF graph of the desired resource.  Most SPARQL queries are constructed from basic graph patterns.  Like RDF triples, a basic graph pattern contains a subject, an object, and a predicate; however, in the basic graph pattern, the subject, object, or predicate may be replaced by variables.  The query result is the set of tuples matching the query pattern [Prud'hommeaux, 2008].

The SPARQL query language is designed to query semantic web data, just as SQL is designed to query relational data. However unlike SQL, SPARQL supports multiple query forms appropriate for  retrieving information from structured graph data. In SPARQL, the following types of queries are possible:
1. "ASK" query to test whether a graph contains  specific data of interest.  This type of query returns a boolean result.
2. "DESCRIBE" query to generate an RDF description of one or more resources.  This type of query returns a set of RDF triples.
3. "CONSTRUCT" query to create a custom RDF graph based on a query criteria. This can also be used to transform RDF data.  This type of query returns a graph matching the query criteria.
4. “SELECT” query to identify triples matching a query pattern.  This type of query returns a set of matching triples.

Our system supports a subset of queries of the “SELECT” type.

\subsection{The Query Federation Problem}

In the relational database world, data is contained within "data silos."  Schemas for different databases are likely to differ, even if the data that is stored is similar.  For example, suppose that two companies, Company A and Company B, construct independent databases,   Database 1 and Database 2 respectively, to describe all known drugs.  Company A may include “brand name,” “generic name,” “active compound,” “cost,” “indications,” and “protein target” information in its schema, while Company B may define the corresponding table in its database using a different schema.  In addition to discrepancies in table schemas, the data itself may be represented differently by the two companies. For example, the protein target of a drug to treat cardiovascular disease may be described as “ABCA1” in Database 1, and by a synonym, “ABC1,” in Database 2.  Therefore, it will be difficult to integrate the data between Database 1 and Database 2 without writing extensive application logic, even though the two databases contain highly overlapping information. In contrast, RDF data is represented in a form that is amenable for data integration spanning multiple data sources, because unique entities are described by unique Uniform Resource Identifiers (URIs).

However, because of the scale free nature of the web, locating the appropriate data endpoint to execute a specific query is a challenging problem. This is known as the query federation problem. While the SPARQL query language has some support for expressing queries that span multiple data sources, there is a need for improved systems able to orchestrate queries across multiple sources.

\section{QueryMed}

\subsection{Architecture overview}

A general overview of QueryMed architecture is shown in Figure 1. The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.

QueryMed is implemented in Java in the backend and JavaScript in the frontend.  The JQuery library [Resig] was used to develop an attractive user interface.  In the backend, the Jena library [McBride] is used to run the SPARQL queries. QueryMed relies on the following remote resources:
SPARQL Endpoints:  QueryMed runs queries on multiple remote SPARQL endpoints that expose biomedical data.
Sources List: The default list of endpoints available to the user is stored in a source list on the proxy server.

\begin{figure}
\centering
%\epsfig{file=fly.eps, height=1in, width=1in}
\caption{QueryMed Architecture Overview}
\end{figure}

\subsection{}

\section{}

\subsection{}

\subsection{}

\subsection{Tables}

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
%\epsfig{file=fly.eps}
\caption{A sample black and white graphic (.eps format).}
\end{figure}

\begin{figure}
\centering
%\epsfig{file=fly.eps, height=1in, width=1in}
\caption{A sample black and white graphic (.eps format)
that has been resized with the \texttt{epsfig} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
\begin{figure*}
\centering
%\epsfig{file=flies.eps}
\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}
and don't forget to end the environment with
{figure*}, not {figure}!

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\begin{figure}
\centering
%\psfig{file=rosette.ps, height=1in, width=1in,}
\caption{A sample black and white graphic (.ps format) that has
been resized with the \texttt{psfig} command.}
\end{figure}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the \texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}

\section{Acknowledgments}

This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
%
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The www2010-submission.cls file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
\balancecolumns % GM July 2000
% That's all folks!
\end{document}

### 2010-02-05 trunk/docs/www2010/paper.tex
copy pasted all the stuff from the db submission

@@
 Although the physician has a single question--based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?--there is no single data source that the physician can use to answer this question.
-The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and GO [Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, Gene Ontology].
+The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and GO \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
 Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
@@
 The semantic web can be viewed as global database system for the data available on the world wide web.
-Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using a standardized query language, SPARQL [Berners-Lee, 2001].
+Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using a standardized query language, SPARQL \cite{Berners-Lee}.
 The addition of structure to web data is intended to allow inferences to be automatically made by intelligent agents drawing on data from multiple sources.
@@
 The relationship "is a professor for" is known as the predicate of the triple; "Sam" is the subject while "6.830" is the object.
-In order for this statement to be machine readable, each of the entities in the statement ("Sam," "is a professor for," and "6.830") will be described by a Uniform Resource Identifier (URI) [Manola, 2004].
-The Web Ontology Language (OWL) is a language that permits the definition of ontologies, allowing additional properties and constraints on relationships among data to be represented [Smith, 2004].
-Some parallels between relational databases and RDF data on the semantic web are illustrated in table 1 [Berners-Lee, 1998].
+In order for this statement to be machine readable, each of the entities in the statement ("Sam," "is a professor for," and "6.830") will be described by a Uniform Resource Identifier (URI) \cite{Manola}.
+The Web Ontology Language (OWL) is a language that permits the definition of ontologies, allowing additional properties and constraints on relationships among data to be represented \cite{Smith}.
+Some parallels between relational databases and RDF data on the semantic web are illustrated in table 1 \cite{Berners-Lee}.
 
 Although most web data are not currently available in semantic web formats, the biomedical community has been extremely proactive in making structured biomedical data available in RDF.
-Thus, the biomedical knowledge domain has been an early success of the semantic web [Yip, 2009].
+Thus, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
 Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
 Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists, and these semantic web resources represent valuable standardized repositories of information that can be automatically mined for biological knowledge.
-The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain [HCLSIG].
+The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
 
@@
 It is important to design user-friendly systems that allow these users to take advantage of the wealth of structured biological knowledge available on the semantic web.
-Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data is often distributed among many sources [Pasquier, 2008].
+Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data is often distributed among many sources \cite{Pasquier}.
 
@@
 Like RDF triples, a basic graph pattern contains a subject, an object, and a predicate; however, in the basic graph pattern, the subject, object, or predicate may be replaced by variables.
-The query result is the set of tuples matching the query pattern [Prud'hommeaux, 2008].
+The query result is the set of tuples matching the query pattern \cite{Prud'hommeaux}.
 
@@
 QueryMed is implemented in Java in the backend and JavaScript in the frontend.
-The JQuery library [Resig] was used to develop an attractive user interface.
-In the backend, the Jena library [McBride] is used to run the SPARQL queries.
+The JQuery library \cite{Resig} was used to develop an attractive user interface.
+In the backend, the Jena library \cite{McBride} is used to run the SPARQL queries.
 QueryMed relies on the following remote resources:
@@
 \centering
-%\epsfig{file=fly.eps, height=1in, width=1in}
+\epsfig{file=images/architecture_overview.jpg}
 \caption{QueryMed Architecture Overview}
-\end{figure}
-
-\subsection{}
-
-\section{}
-
-\subsection{}
-
-\subsection{}
-
-\subsection{Tables}
-
-\begin{table}
-\centering
-\caption{Frequency of Special Characters}
-\begin{tabular}{|c|c|l|} \hline
-Non-English or Math&Frequency&Comments\\ \hline
-\O & 1 in 1,000& For Swedish names\\ \hline
-$\pi$ & 1 in 5& Common in math\\ \hline
-\$ & 4 in 5 & Used in business\\ \hline
-$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
-\hline\end{tabular}
-\end{table}
-
-To set a wider table, which takes up the whole width of
-the page's live area, use the environment
-\textbf{table*} to enclose the table's contents and
-the table caption.  As with a single-column table, this wide
-table will ``float" to a location deemed more desirable.
-Immediately following this sentence is the point at which
-Table 2 is included in the input file; again, it is
-instructive to compare the placement of the
-table here with the table in the printed dvi
-output of this document.
-
-\begin{table*}
-\centering
-\caption{Some Typical Commands}
-\begin{tabular}{|c|c|l|} \hline
-Command&A Number&Comments\\ \hline
-\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
-\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
-\texttt{{\char'134}table}& 300 & For tables\\ \hline
-\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
-\end{table*}
-
-\subsection{Figures}
-Like tables, figures cannot be split across pages; the
-best placement for them
-is typically the top or the bottom of the page nearest
-their initial cite.  To ensure this proper ``floating'' placement
-of figures, use the environment
-\textbf{figure} to enclose the figure and its caption.
-
-This sample document contains examples of \textbf{.eps}
-and \textbf{.ps} files to be displayable with \LaTeX.  More
-details on each of these is found in the \textit{Author's Guide}.
-
-\begin{figure}
-\centering
-%\epsfig{file=fly.eps}
-\caption{A sample black and white graphic (.eps format).}
 \end{figure}
@@
 \centering
-%\epsfig{file=fly.eps, height=1in, width=1in}
-\caption{A sample black and white graphic (.eps format)
-that has been resized with the \texttt{epsfig} command.}
+\epsfig{file=images/architecture_detail.jpg}
+\caption{QueryMed Architecture details}
 \end{figure}
 
-As was the case with tables, you may want a figure
-that spans two columns.  To do this, and still to
-ensure proper ``floating'' placement of tables, use the environment
-\textbf{figure*} to enclose the figure and its caption.
-\begin{figure*}
-\centering
-%\epsfig{file=flies.eps}
-\caption{A sample black and white graphic (.eps format)
-that needs to span two columns of text.}
-\end{figure*}
-and don't forget to end the environment with
-{figure*}, not {figure}!
+\subsubsection{Proxy Server}
 
-Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
-used; use
-the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
-commands as appropriate for the different file types.
+The user submits a query to the SPARQL endpoints via the proxy server.
+The query is translated into a SPARQL query for each individual endpoint, and the results are returned to the user.
+The specific components implemented in the proxy server are as follows.
+Source Manager: The source manager reads the source list and populates the default query list on the user interface.
+It also keeps track of the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
+Translator: This component is responsible for translating the user query into valid SPARQL syntax.
+Run Query: The purpose of this service is to run SPARQL queries.
+It takes input on source and keyword from the client, and returns  the query results as a JSON object.
+The structure of the JSON object is given in figure 4.
+Get Properties: This service takes as input an individual source, and returns an array of all properties for that source.
+Get Property Values: This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
+Advanced Queries: This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
+The structure of this JSON object is given in figure 5.
+After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure 4.
+
+\subsubsection{User Interface}
+
+The query interface is implemented in JavaScript, HTML and CSS. The main components of the user interface are as follows: 
+Initial Query Interface: The user inputs their query  term into the provided text box (figure 6).
+Add New Source: Allows the user to add a new data source to extend the query. (figure 8)
+Dynamic Facet Creator/Filter: This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure 13).
+Result Browser: The user can view their query results, organized by source.
+The results are presented in a table with pagination.
+The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure 14).
+
+\subsection{System Functionality and Design Decisions}
+
+\subsubsection{Source list}
+
+Because the set of default endpoints is stored on the proxy server, the set of resources available by default to the user can easily be updated.
+Since useful biomedical resources are rapidly being developed and made available as RDF triple stores, the ease of updating the resource list ensures that the system can easily be brought up to date.
+In fact, the QueryMed system could easily be adapted to perform queries outside the biomedical domain by modifying the list of input data sources.
+
+\subsubsection{Proxy Server}
+
+While an entirely client side application is possible, we chose to have a proxy server perform the SPARQL query execution and caching. This design is advantageous for several reasons:
+1.
+Efficient cache management: The result set from running an unrestricted query can include millions of results.
+It may be infeasible to keep unfiltered query results in browser memory.
+A typical memory footprint for a browser (for e.g.
+Firefox) is usually between 20MB and 100MB.
+A poorly constructed query can result in gigabytes worth of triples returned, causing the browser to crash.
+The proxy server can cache results from initial query execution, so that only a filtered result set is subsequently returned to the client.
+2. The proxy server enables us to avoid cross domain XML-HTTP-Request errors in accessing SPARQL endpoints in various domains.
+
+\subsubsection{Data Structures}
+
+The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
+In the general ``Query All” case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
+The generated SPARQL query will be of the form illustrated in figure 3.
+The word “input” in the figure represents the keyword specified by the user.
+
+\begin{verbatim}
+SELECT ?projection_1 ?projection_2 ... ?projection_n
+WHERE {?x source:property ? projection 
+	FILTER regex(? projection, '" + input +"', 'i)
+}
+... //Other property filers are to follow
+\end{verbatim}
+Minimal SPARQL query structure in the “Query All” case
+
+When the results are returned, each result set is structured as the JSON object given in figure 4.
+This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables .
+how many results are returned and the result set.
+
+When the user runs an advanced query, the payload to be sent to the proxy server to do the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All” case.
+The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
+This information is passed as a JSON object , whose structure is given in figure 5, to the proxy server.
+
+\begin{verbatim}
+{
+     "bindings":
+     [
+     	{"source" : "source_label",
+     	 "uri" : "http://source.com/sparql.",
+     	 "vars" : ["variable1","variable2"],
+     	 "count":2,
+     	 "results" : 
+     		[
+     		 {"variable1": "Value 1", 		   "variable1": "Value 2"},
+     		 {"variable1": "Another Value 1", 		   "variable1": "Another Value 2"},
+
+     		]
+     	}
+     	,
+	{ //Another Source and Result Description
+	...
+	},
+	...
+     ]
+ }
+\end{verbatim}
+JSON returned from the RunQuery service
+
+\begin{verbatim}
+{
+   "http://some-source/sparql": {
+                            "property1": {"value1": "AND"},
+	              "property2": {"value 2": "OR"},
+	              "property2": {"value 3": "FILTER"}
+	           }
+	}
+   //Other sources ...
+}
+\end{verbatim}
+JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators
+
+\subsubsection{Query Interface}
+
+In order to make QueryMed easy to use, but still provide a flexible tool that is able to perform a range of queries, we provide an uncluttered interface with only a  blank search box and two buttons.
+The ``Query All” button will perform a keyword-based query over pre-selected data sources.
+The advanced search option ``Refine Query” allows the user to view additional query options.
+The initial query interface is shown in figure 6.
 
@@
 \centering
-%\psfig{file=rosette.ps, height=1in, width=1in,}
-\caption{A sample black and white graphic (.ps format) that has
-been resized with the \texttt{psfig} command.}
+\epsfig{file=images/new_front.jpg,  width=3in}
+\caption{Initial query interface seen by the user}
 \end{figure}
 
-\subsection{Theorem-like Constructs}
-Other common constructs that may occur in your article are
-the forms for logical constructs like theorems, axioms,
-corollaries and proofs.  There are
-two forms, one produced by the
-command \texttt{{\char'134}newtheorem} and the
-other by the command \texttt{{\char'134}newdef}; perhaps
-the clearest and easiest way to distinguish them is
-to compare the two in the output of this sample document:
+We believe that the unstructured keyword search is likely to encourage users without previous experience running queries on semantic web data to explore the additional features of our system that allow more flexible and advanced queries.
+For example, physicians who are not used to running queries on semantic web data may be more willing to invest time into learning how to find better search results once they see that the system can easily provide them with some initial useful query results using the basic keyword search option.
+However, the execution of queries using QueryMed will be very much different from a user issuing a web search on a search engine such as Google, because it exploits the underlying structure of the data as exposed in the individual data endpoints.
 
-This uses the \textbf{theorem} environment, created by
-the \texttt{{\char'134}newtheorem} command:
-\newtheorem{theorem}{Theorem}
-\begin{theorem}
-Let $f$ be continuous on $[a,b]$.  If $G$ is
-an antiderivative for $f$ on $[a,b]$, then
-\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
-\end{theorem}
+The advanced query interface provides the user with the option of selecting specific data sources to query, as shown in figure 7.
+The user can also dynamically add additional data sources, as shown in figure 8.
+This feature allows the user to perform their query only over relevant or trusted data sources.
+The ability to dynamically add additional sources increases the flexibility of  the QueryMed system, allowing users to search endpoints of interest that are not included in the default list.
 
-The other uses the \textbf{definition} environment, created
-by the \texttt{{\char'134}newdef} command:
-\newdef{definition}{Definition}
-\begin{definition}
-If $z$ is irrational, then by $e^z$ we mean the
-unique number which has
-logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
-\end{definition}
+\begin{figure}
+\centering
+\epsfig{file=images/select.jpg,  width=3in}
+\caption{The user has the option of selecting specific trusted or relevant data sources, or of adding additional data sources to query.}
+\end{figure}
 
-Two lists of constructs that use one of these
-forms is given in the
-\textit{Author's  Guidelines}.
+\begin{figure}
+\centering
+\epsfig{file=images/add_source.jpg,  width=3in}
+\caption{The user can also dynamically add new data sources}
+\end{figure}
 
-There is one other similar construct environment, which is
-already set up
-for you; i.e. you must \textit{not} use
-a \texttt{{\char'134}newdef} command to
-create it: the \textbf{proof} environment.  Here
-is a example of its use:
-\begin{proof}
-Suppose on the contrary there exists a real number $L$ such that
-\begin{displaymath}
-\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
-\end{displaymath}
-Then
-\begin{displaymath}
-l=\lim_{x\rightarrow c} f(x)
-= \lim_{x\rightarrow c}
-\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
-= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
-\frac{f(x)}{g(x)} = 0\cdot L = 0,
-\end{displaymath}
-which contradicts our assumption that $l\neq 0$.
-\end{proof}
+When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
+The properties list is generated by running a SPARQL query of the form given in figure 9 at the specified data source.
+This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
 
-Complete rules about using these environments and using the
-two different creation commands are in the
-\textit{Author's Guide}; please consult it for more
-detailed instructions.  If you need to use another construct,
-not listed therein, which you want to have the same
-formatting as the Theorem
-or the Definition\cite{salas:calculus} shown above,
-use the \texttt{{\char'134}newtheorem} or the
-\texttt{{\char'134}newdef} command,
-respectively, to create it.
+\begin{verbatim}
+SELECT DISTINCT ?property
+WHERE { [] ?property [] }
+ORDER BY ?property
+\end{verbatim}
 
-\subsection*{A {\secit Caveat} for the \TeX\ Expert}
-Because you have just been given permission to
-use the \texttt{{\char'134}newdef} command to create a
-new form, you might think you can
-use \TeX's \texttt{{\char'134}def} to create a
-new command: \textit{Please refrain from doing this!}
-Remember that your \LaTeX\ source code is primarily intended
-to create camera-ready copy, but may be converted
-to other forms -- e.g. HTML. If you inadvertently omit
-some or all of the \texttt{{\char'134}def}s recompilation will
-be, to say the least, problematic.
+Once the properties are returned, those will be displayed in the user interface as shown in figure 10.
+Clicking an individual property link displays a description of the property, allowing the user to understand the keywords that will be most useful to search on each specific property.
+A sample description for the property rdfs:label is shown in figure 11.
 
-\section{Conclusions}
+\begin{figure}
+\centering
+\epsfig{file=images/add_props.jpg, width=3.5in}
+\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
+\end{figure}
+
+The user can choose to perform exact queries, or filter results on specific keywords.
+If the user does not know the specific value for a property, they can instead specify a keywords using the “FILTER” option.
+They also can choose logical operators to connect the various parts of their query.
+When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
+as a conjunction.
+When “OR” is used the specified graph pattern is made to disjunct with the rest of the query with the SPARQL UNION operator.
+If no logical operator is specified, “AND” will be used by default.
+The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in figure 12.
+
+\begin{verbatim}
+SELECT distinct ?disease WHERE { 
+{?x <http://www.w3.org/2000/01/rdf-schema#label> 
+?disease 
+FILTER regex(?disease, "coronary artery disease", "i"). 
+?x <http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseasome/class> 
+<http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseaseClass/Cardiovascular>} 
+UNION { ?x <http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseasome/associatedGene>
+ <http://www4.wiwiss.fu-berlin.de/diseasome/resource/genes/ABCA1>.}
+}
+\end{verbatim}
+A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.
+
+The Dynamic Facet Creator shown in figure 13 allows the user to view properties from multiple endpoints, and to expand and contract properties for individual endpoints.
+This feature enables the interface to display many property lists simultaneously, grouped by source, so that only information relevant to the endpoint the user is currently examining will be visible at any time.
+
+\begin{figure}
+\centering
+\epsfig{file=images/facet_creator.jpg, width=3in}
+\caption{Dynamic Facet Creator and Filter}
+\end{figure}
+
+Query results are displayed to the user as a table, which can be filtered and searched to refine the query result, or printed as shown in in figure 14.
+This feature allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+
+\begin{figure}
+\centering
+\epsfig{file=images/results.jpg, width=3.5in}
+\caption{A sample result table}
+\end{figure}
+
+\subsection{Performance}
+
+The slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
+We compared the times required to load properties from several endpoints, by running the query illustrated in figure 9.
+Timing data are shown in table 2.
+
+\begin{table}
+\begin{tabular}{|l||l|l|l|}
+\hline
+ & Dieseasome & Dailymed & Drugbank \\
+\hline
+\hline
+1st Trial   & 3.45 s & 1.77 s  & 9.51 s  \\
+2nd Trial  & 1.61 s & 1.57 s & 9.34 s \\
+3rf Trial  & 1.71 s & 1.66 s & 9.06 s \\
+\hline
+\end{tabular}
+\caption{Running times to retrieve all the properties from selected endpoints}
+\end{table}
+
+For all selected endpoints, the property values take longer to load on the initial query than in subsequent iterations, probably due to browser caching.
+The running time to acquire the list of property values for the Drugbank endpoint is much larger than the time for the other two endpoints, probably due to the greater size of the Drugbank database.
+(As of 12/10/09, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, and DrugBank contains 765,936 triples).
+
+\subsection{QueryMed Resources}
+
+The source code for the Querymed system is available at the QueryMed Google Code project: 
+http://code.google.com/p/querymed/
+
+A video displaying a sample use case can be found at: 
+http://www.youtube.com/watch?v=JCr7vryS9Vg
+
+\section{Related Work}
+
+A number of existing tools aim to provide a user-friendly interface for browsing semantic web data, or to allow users to perform federated queries.  Several of these are described below.
+
+The SMART query tool is a web-based application designed to allow biologists to run SPARQL queries over multiple endpoints.
+Queries to the SMART system are constructed using a descriptive logic query written in the natural-language like Manchester OWL syntax \cite{Battista}.
+However, it cannot be expected that the end users know how  to use the Manchester OWL syntax to write queries, whereas in QueryMed, users can construct queries intuitively by giving keywords for user selected properties.
+
+GoWeb is another system designed for answering queries on biomedical data.
+It allows users to run traditional keyword-based web search with ontology search features \cite{Dietze}.
+After a keyword search, documents can be filtered based on the biomedical annotations they contain.
+However, in GoWeb the exact sources queried are not transparent and cannot be selected or modified by the user as in QueryMed.
+
+Twinkle offers a stand-alone graphical user interface to load and edit SPARQL queries that can be used to query online SPARQL endpoints \cite{Dodds}.
+Our system differs from the Twinkle system in several aspects.
+First of all, in Twinkle, the user is expected to know what is already available at the SPARQL endpoints to write the query.
+But in QueryMed, we only ask for specific keywords of interest, and give the option of restricting the query should the user wish to run a more precise query.
+Second, although Twinkle was designed to be a more general purpose system, it only supports a small number of specific SPARQL endpoints, while QueryMed allows the user to dynamically add SPARQL endpoints.
+
+Most SPARQL query engines are designed to run queries against individual endpoints.
+But it is often useful to draw on multiple web resources in answering a query.
+There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambrite}systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
+
+\begin{table*}
+\begin{center}
+\begin{tabular}{| p{2cm} | p{2cm}| p{2cm}| p{2cm}| p{2cm}| p{2cm} |}
+\hline
+ & Query Multiple Sources? & Dynamic Addition of Sources & Allows Keyword Queries & Open Source & GUI \\
+\hline
+\hline
+QueryMed  & Yes & Yes & Yes & Yes & Yes  \\
+\hline
+SMART  & Yes & No & No & Yes & Yes  \\
+\hline
+DARQ  & Yes & N/A & No & Yes & No  \\
+\hline
+GoWeb  & Yes & No & Yes & No & Yes  \\
+\hline
+BioGateway  & Yes & No & No & No & Yes  \\
+\hline
+Twinkle  & Yes & No & No & Yes & Yes  \\
+\hline
+\end{tabular}
+\caption{Comparison of selected features of the QueryMed system with other related systems.}
+\end{center}
+\end{table*}
+
+Table 3 compares selected features of the QueryMed system with other related systems.
+The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user.
+Other features of the QueryMed system that distinguished it from similar systems included the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the semantic web data.
+This feature increases the ease of use of our system relative to other similar systems.
+Furthermore, the Javascript based user interface of the QueryMed system, implemented using the JQuery library, makes our user interface particularly attractive, easy to interact with, and capable of handling a variety of user input events.
+Another unique feature of our system is the property-based advanced query interface.
+This enables users to take advantage of the structure of the underlying ontologies used to represent the data without prior knowledge of the ontology structures.
+
+\section{Future Work and Conclusions}
+
+The bottleneck in the running time of our system is running queries that must retrieve many triples from slow remote SPARQL endpoints.
+By caching relevant data locally, the running time could be significantly improved.
+We observed that the QueryMed system performs significantly faster the second time that a query is run on a given resource possibly due to browser caching.
+By managing our own cache to contain the data most likely to be needed on the proxy server, we could reduce running time still further.
+Another approach that could reduce running time still further, especially for users with a slow network connection, would be to create a complementary standalone application that gives the user the option at startup time of loading all the required RDF data.
+Since data is stored locally after the initial startup, using the system in subsequent queries will be rapid after an initial loading phase.
+This approach might be too memory-intensive if the user wishes to run queries over many large triple stores, but might work best in the situation where there is a small or moderate amount of data in the repositories of interest to the user.
+
+Our system currently allows only a restricted set of SPARQL queries.
+Supporting additional types of queries and including query optimization functionality could increase both the flexibility and speed of the QueryMed system.
+By drawing on the expressivity of the SPARQL language and the information contained in the ontologies used to represent the data, it would be possible to extend our system into an intelligent reasoning system on biomedical data which allows physicians to enter sophisticated, complex queries and find relevant biomedical results.
+
+We also believe that it would be useful to allow exploration of the relationships among multiple data sources containing similar resources.
+One challenge in integrating biomedical data across multiple data sources is that individual data items (i.e, a specific protein or a specific drug) may be represented by distinct URIs at different endpoints.
+The diversity of representations of identical data items across different biomedical data sources makes it difficult to automatically combine these items.
+However, there is some cross-referencing between the biomedical data sources that we examined (for example, DailyMed drugs sometimes refer to diseases in the Diseasome SPARQL endpoint by their URI).
+It might be useful to provide a representation of the relationships among search results from different sources with the query results.
+Additionally, the system could automatically detect similar data items even if they are identified by distinct URIs, and group similar items from distinct data sources together in the query results by natural language processing.
+Another useful feature might be to allow the user to view the relationships among distinct URIs (as in the relfinder system \cite{Heim} , which displays possible paths through the RDF graph between distinct resources).
+
+Another feature that we have begun to implement in our system is an autocompletion feature to automatically retrieve the distinct values for any given property.
+This feature will allow the user to automatically see all valid choices for each property, and will make the system easier to use and reduce empty query result sets.
+
+Since a major goal of QueryMed is ease of use by physicians, life scientists, and patients, we also plan to perform a user study to understand how effectively  users without knowledge of SPARQL can interact with our system.
+We plan to use data from this study to further refine our system to improve its usability.
+
+The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
+Because of the unique features of our system, we believe it will be of use to the biomedical community.
 
@@
 
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.  In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
+\bibliographystyle{abbrv}
+\bibliography{references}
 
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional
-%\balancecolumns
-\appendix
-%Appendix A
-\section{Headings in Appendices}
-The rules about hierarchical headings discussed above for
-the body of the article are different in the appendices.
-In the \textbf{appendix} environment, the command
-\textbf{section} is used to
-indicate the start of each Appendix, with alphabetic order
-designation (i.e. the first is A, the second B, etc.) and
-a title (if you include one).  So, if you need
-hierarchical structure
-\textit{within} an Appendix, start with \textbf{subsection} as the
-highest level. Here is an outline of the body of this
-document in Appendix-appropriate form:
-\subsection{Introduction}
-\subsection{The Body of the Paper}
-\subsubsection{Type Changes and  Special Characters}
-\subsubsection{Math Equations}
-\paragraph{Inline (In-text) Equations}
-\paragraph{Display Equations}
-\subsubsection{Citations}
-\subsubsection{Tables}
-\subsubsection{Figures}
-\subsubsection{Theorem-like Constructs}
-\subsubsection*{A Caveat for the \TeX\ Expert}
-\subsection{Conclusions}
-\subsection{Acknowledgments}
-\subsection{Additional Authors}
-This section is inserted by \LaTeX; you do not insert it.
-You just add the names and information in the
-\texttt{{\char'134}additionalauthors} command at the start
-of the document.
-\subsection{References}
-Generated by bibtex from your ~.bib file.  Run latex,
-then bibtex, then latex twice (to resolve references)
-to create the ~.bbl file.  Insert that ~.bbl file into
-the .tex source file and comment out
-the command \texttt{{\char'134}thebibliography}.
-%
-% This next section command marks the start of
-% Appendix B, and does not continue the present hierarchy
-\section{More Help for the Hardy}
-The www2010-submission.cls file itself is chock-full of succinct
-and helpful comments.  If you consider yourself a moderately
-experienced to expert user of \LaTeX, you may find reading
-it useful but please remember not to change it.
-\balancecolumns % GM July 2000
-% That's all folks!
 \end{document}

### 2010-02-08 docs/www2010/paper.tex
fixing up paper formatting; removing irrelevant sections

--- 2010-02-05 trunk/docs/www2010/paper.tex
@@
 \documentclass{www2010-submission}
+\usepackage{moreverb}
 
@@
 After retrieval of the initial result set, query results can be filtered to improve their relevance.
-As an advanced querying feature, the system also allows the user to exploit the underlying structure of the RDF data to improve query results.
+As an advanced search feature, the system also allows the user to exploit the underlying structure of the RDF data to improve query results.
 \end{abstract}
@@
 
-RDF data in the semantic web, like the data in relational databases, is modeled in a structured format that can be queried using a standardized query language.
 The biomedical domain is among the early successes of the semantic web, due to the rapidity with which the biomedical community has made its data available in RDF triple stores.
@@
 
-For example, a physician may know her patient’s personal information, symptoms, current medications, and genotype.
-She may wish to determine the patient’s treatment plan and identify clinical trials for which the patient is eligible.
+For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
+She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.
 Although the physician has a single question--based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?--there is no single data source that the physician can use to answer this question.
@@
 The information about the disease can be found in Diseasome.
-She might also want to know the set of drugs in DailyMed and Drugbank that can be used to treat coronary artery disease and that also will not react adversely with the patient’s current medications.
+She might also want to know the set of drugs in DailyMed and Drugbank that can be used to treat coronary artery disease.
 Information on clinical trials available for her patient can be found in LinkedCT.
@@
 
-The semantic web can be viewed as global database system for the data available on the world wide web.
-Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using a standardized query language, SPARQL \cite{Berners-Lee}.
-The addition of structure to web data is intended to allow inferences to be automatically made by intelligent agents drawing on data from multiple sources.
-
-RDF data is described by a directed, labeled graph, with relationships among data items represented  in triples containing a subject, an object and a predicate.
-For example, we might want to express the concept, "Sam is a professor for 6.830." The corresponding triple expresses the relationship "is a professor for" between two entities, “Sam” and “6.830”.
-The relationship "is a professor for" is known as the predicate of the triple; "Sam" is the subject while "6.830" is the object.
-In order for this statement to be machine readable, each of the entities in the statement ("Sam," "is a professor for," and "6.830") will be described by a Uniform Resource Identifier (URI) \cite{Manola}.
-The Web Ontology Language (OWL) is a language that permits the definition of ontologies, allowing additional properties and constraints on relationships among data to be represented \cite{Smith}.
-Some parallels between relational databases and RDF data on the semantic web are illustrated in table 1 \cite{Berners-Lee}.
-
-Although most web data are not currently available in semantic web formats, the biomedical community has been extremely proactive in making structured biomedical data available in RDF.
-Thus, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
+Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
 Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
@@
 Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data is often distributed among many sources \cite{Pasquier}.
-
-\subsection{SPARQL Query Language}
-
-A SPARQL query performs a pattern-matching search on the underlying RDF graph of the desired resource.
-Most SPARQL queries are constructed from basic graph patterns.
-Like RDF triples, a basic graph pattern contains a subject, an object, and a predicate; however, in the basic graph pattern, the subject, object, or predicate may be replaced by variables.
-The query result is the set of tuples matching the query pattern \cite{Prud'hommeaux}.
-
-The SPARQL query language is designed to query semantic web data, just as SQL is designed to query relational data.
-However unlike SQL, SPARQL supports multiple query forms appropriate for  retrieving information from structured graph data.
-In SPARQL, the following types of queries are possible:
-1. "ASK" query to test whether a graph contains  specific data of interest.  This type of query returns a boolean result.
-2. "DESCRIBE" query to generate an RDF description of one or more resources.  This type of query returns a set of RDF triples.
-3. "CONSTRUCT" query to create a custom RDF graph based on a query criteria. This can also be used to transform RDF data.  This type of query returns a graph matching the query criteria.
-4. “SELECT” query to identify triples matching a query pattern.  This type of query returns a set of matching triples.
-
-Our system supports a subset of queries of the “SELECT” type.
-
-\subsection{The Query Federation Problem}
-
-In the relational database world, data is contained within "data silos."  Schemas for different databases are likely to differ, even if the data that is stored is similar.
-For example, suppose that two companies, Company A and Company B, construct independent databases,   Database 1 and Database 2 respectively, to describe all known drugs.
-Company A may include “brand name,” “generic name,” “active compound,” “cost,” “indications,” and “protein target” information in its schema, while Company B may define the corresponding table in its database using a different schema.
-In addition to discrepancies in table schemas, the data itself may be represented differently by the two companies.
-For example, the protein target of a drug to treat cardiovascular disease may be described as “ABCA1” in Database 1, and by a synonym, “ABC1,” in Database 2.
-Therefore, it will be difficult to integrate the data between Database 1 and Database 2 without writing extensive application logic, even though the two databases contain highly overlapping information.
-In contrast, RDF data is represented in a form that is amenable for data integration spanning multiple data sources, because unique entities are described by unique Uniform Resource Identifiers (URIs).
-
-However, because of the scale free nature of the web, locating the appropriate data endpoint to execute a specific query is a challenging problem.
-This is known as the query federation problem.
-While the SPARQL query language has some support for expressing queries that span multiple data sources, there is a need for improved systems able to orchestrate queries across multiple sources.
 
@@
 QueryMed relies on the following remote resources:
-SPARQL Endpoints:  QueryMed runs queries on multiple remote SPARQL endpoints that expose biomedical data.
-Sources List: The default list of endpoints available to the user is stored in a source list on the proxy server.
+\begin{itemize}
+\item \textbf{SPARQL Endpoints}:  QueryMed runs queries on multiple remote SPARQL endpoints that expose biomedical data.
+\item \textbf{Sources List}: The default list of endpoints available to the user is stored in a source list on the proxy server.
+\end{itemize}
 
@@
 The specific components implemented in the proxy server are as follows.
-Source Manager: The source manager reads the source list and populates the default query list on the user interface.
+\begin{itemize}
+\item \textbf{Source Manager:} The source manager reads the source list and populates the default query list on the user interface.
 It also keeps track of the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
-Translator: This component is responsible for translating the user query into valid SPARQL syntax.
-Run Query: The purpose of this service is to run SPARQL queries.
+\item \textbf{Translator:} This component is responsible for translating the user query into valid SPARQL syntax.
+\item \textbf{Run Query:} The purpose of this service is to run SPARQL queries.
 It takes input on source and keyword from the client, and returns  the query results as a JSON object.
 The structure of the JSON object is given in figure 4.
-Get Properties: This service takes as input an individual source, and returns an array of all properties for that source.
-Get Property Values: This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
-Advanced Queries: This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
+\item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
+\item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
+\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
 The structure of this JSON object is given in figure 5.
 After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure 4.
+\end{itemize}
 
@@
 The query interface is implemented in JavaScript, HTML and CSS. The main components of the user interface are as follows: 
-Initial Query Interface: The user inputs their query  term into the provided text box (figure 6).
-Add New Source: Allows the user to add a new data source to extend the query. (figure 8)
-Dynamic Facet Creator/Filter: This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure 13).
-Result Browser: The user can view their query results, organized by source.
+\begin{itemize}
+\item \textbf{Initial Query Interface:} The user inputs their query  term into the provided text box (figure 6).
+\item \textbf{Add New Source:} Allows the user to add a new data source to extend the query. (figure 8)
+\item \textbf{Dynamic Facet Creator/Filter:} This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure 13).
+\item \textbf{Result Browser:} The user can view their query results, organized by source.
 The results are presented in a table with pagination.
 The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure 14).
+\end{itemize}
 
@@
 
-\begin{verbatim}
-SELECT ?projection_1 ?projection_2 ... ?projection_n
+\begin{verbatimtab}
+SELECT ?projection_1 ?projection_2 ... 
+	?projection_n
 WHERE {?x source:property ? projection 
-	FILTER regex(? projection, '" + input +"', 'i)
+	FILTER regex(? projection, '" + 
+	input +"', 'i)
 }
 ... //Other property filers are to follow
-\end{verbatim}
+\end{verbatimtab}
 Minimal SPARQL query structure in the “Query All” case
@@
 
-\begin{verbatim}
+\begin{verbatimtab}
 {
@@
      		[
-     		 {"variable1": "Value 1", 		   "variable1": "Value 2"},
-     		 {"variable1": "Another Value 1", 		   "variable1": "Another Value 2"},
+     		 {"variable1": "Value 1",
+		  "variable1": "Value 2"},
+     		 {"variable1": "Another Value 1",
+		  "variable1": "Another Value 2"},
 
@@
  }
-\end{verbatim}
+\end{verbatimtab}
 JSON returned from the RunQuery service
 
-\begin{verbatim}
+\begin{verbatimtab}
 {
@@
 }
-\end{verbatim}
+\end{verbatimtab}
 JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators
@@
 
-\begin{verbatim}
+\begin{verbatimtab}
 SELECT DISTINCT ?property
@@
 ORDER BY ?property
-\end{verbatim}
+\end{verbatimtab}
 
@@
 
-\begin{verbatim}
+\begin{verbatimtab}
 SELECT distinct ?disease WHERE { 
@@
 ?disease 
-FILTER regex(?disease, "coronary artery disease", "i"). 
-?x <http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseasome/class> 
-<http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseaseClass/Cardiovascular>} 
-UNION { ?x <http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseasome/associatedGene>
- <http://www4.wiwiss.fu-berlin.de/diseasome/resource/genes/ABCA1>.}
+FILTER regex(?disease, 
+"coronary artery disease", "i"). 
+?x <http://www4.wiwiss.fu-berlin.de/diseasome/
+resource/diseasome/class> 
+<http://www4.wiwiss.fu-berlin.de/diseasome/
+resource/diseaseClass/Cardiovascular>} 
+UNION { ?x <http://www4.wiwiss.fu-berlin.de/
+diseasome/resource/diseasome/associatedGene>
+ <http://www4.wiwiss.fu-berlin.de/diseasome/
+ resource/genes/ABCA1>.}
 }
-\end{verbatim}
+\end{verbatimtab}
 A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.
@@
 
-The source code for the Querymed system is available at the QueryMed Google Code project: 
+The source code for the Querymed system is available at the QueryMed Google Code project: \\
 http://code.google.com/p/querymed/
 
-A video displaying a sample use case can be found at: 
+A video displaying a sample use case can be found at: \\
 http://www.youtube.com/watch?v=JCr7vryS9Vg

### 2010-02-08 trunk/docs/www2010/paper.tex
adding sample use case section; updating figure references; formatting & proofreading

@@
 \documentclass{www2010-submission}
+\usepackage{fancyvrb}
 
@@
 Developing intuitive ways to query multiple data sources and display results is both an important and a challenging problem.
+Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.
 
@@
 The semantic web can be viewed as global database system for the data available on the world wide web.
-Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using a standardized query language, SPARQL \cite{Berners-Lee}.
-The addition of structure to web data is intended to allow inferences to be automatically made by intelligent agents drawing on data from multiple sources.
+Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using a standardized query language, SPARQL \cite{Berners-Lee1}.
+The addition of structure to web data allows inferences to be automatically drawn by intelligent agents integrating data from multiple sources.
 
-RDF data is described by a directed, labeled graph, with relationships among data items represented  in triples containing a subject, an object and a predicate.
-For example, we might want to express the concept, "Sam is a professor for 6.830." The corresponding triple expresses the relationship "is a professor for" between two entities, “Sam” and “6.830”.
-The relationship "is a professor for" is known as the predicate of the triple; "Sam" is the subject while "6.830" is the object.
-In order for this statement to be machine readable, each of the entities in the statement ("Sam," "is a professor for," and "6.830") will be described by a Uniform Resource Identifier (URI) \cite{Manola}.
-The Web Ontology Language (OWL) is a language that permits the definition of ontologies, allowing additional properties and constraints on relationships among data to be represented \cite{Smith}.
-Some parallels between relational databases and RDF data on the semantic web are illustrated in table 1 \cite{Berners-Lee}.
-
-Although most web data are not currently available in semantic web formats, the biomedical community has been extremely proactive in making structured biomedical data available in RDF.
-Thus, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
+Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
 Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
@@
 
-However, although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources of maximal use to physicians, patients, and life scientists.
+However, although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources accessible to physicians, patients, and life scientists.
 One challenge is constructing systems that allow end users to run intuitive queries on biomedical data.
@@
 
-\subsection{SPARQL Query Language}
-
-A SPARQL query performs a pattern-matching search on the underlying RDF graph of the desired resource.
-Most SPARQL queries are constructed from basic graph patterns.
-Like RDF triples, a basic graph pattern contains a subject, an object, and a predicate; however, in the basic graph pattern, the subject, object, or predicate may be replaced by variables.
-The query result is the set of tuples matching the query pattern \cite{Prud'hommeaux}.
-
-The SPARQL query language is designed to query semantic web data, just as SQL is designed to query relational data.
-However unlike SQL, SPARQL supports multiple query forms appropriate for  retrieving information from structured graph data.
-In SPARQL, the following types of queries are possible:
-1. "ASK" query to test whether a graph contains  specific data of interest.  This type of query returns a boolean result.
-2. "DESCRIBE" query to generate an RDF description of one or more resources.  This type of query returns a set of RDF triples.
-3. "CONSTRUCT" query to create a custom RDF graph based on a query criteria. This can also be used to transform RDF data.  This type of query returns a graph matching the query criteria.
-4. “SELECT” query to identify triples matching a query pattern.  This type of query returns a set of matching triples.
-
-Our system supports a subset of queries of the “SELECT” type.
-
-\subsection{The Query Federation Problem}
-
-In the relational database world, data is contained within "data silos."  Schemas for different databases are likely to differ, even if the data that is stored is similar.
-For example, suppose that two companies, Company A and Company B, construct independent databases,   Database 1 and Database 2 respectively, to describe all known drugs.
-Company A may include “brand name,” “generic name,” “active compound,” “cost,” “indications,” and “protein target” information in its schema, while Company B may define the corresponding table in its database using a different schema.
-In addition to discrepancies in table schemas, the data itself may be represented differently by the two companies.
-For example, the protein target of a drug to treat cardiovascular disease may be described as “ABCA1” in Database 1, and by a synonym, “ABC1,” in Database 2.
-Therefore, it will be difficult to integrate the data between Database 1 and Database 2 without writing extensive application logic, even though the two databases contain highly overlapping information.
-In contrast, RDF data is represented in a form that is amenable for data integration spanning multiple data sources, because unique entities are described by unique Uniform Resource Identifiers (URIs).
-
-However, because of the scale free nature of the web, locating the appropriate data endpoint to execute a specific query is a challenging problem.
-This is known as the query federation problem.
-While the SPARQL query language has some support for expressing queries that span multiple data sources, there is a need for improved systems able to orchestrate queries across multiple sources.
-
 \section{QueryMed}
@@
 
-A general overview of QueryMed architecture is shown in Figure 1.
+A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
 The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.
@@
 \caption{QueryMed Architecture Overview}
+\label{fig:arch_overview}
 \end{figure}
@@
 \caption{QueryMed Architecture details}
+\label{fig:arch_details}
 \end{figure}
@@
 The specific components implemented in the proxy server are as follows.
-Source Manager: The source manager reads the source list and populates the default query list on the user interface.
+\begin{itemize}
+\item \textbf{Source Manager:} The source manager reads the source list and populates the default query list on the user interface.
 It also keeps track of the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
-Translator: This component is responsible for translating the user query into valid SPARQL syntax.
-Run Query: The purpose of this service is to run SPARQL queries.
+\item \textbf{Translator:} This component is responsible for translating the user query into valid SPARQL syntax.
+\item \textbf{Run Query:} The purpose of this service is to run SPARQL queries.
 It takes input on source and keyword from the client, and returns  the query results as a JSON object.
-The structure of the JSON object is given in figure 4.
-Get Properties: This service takes as input an individual source, and returns an array of all properties for that source.
-Get Property Values: This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
-Advanced Queries: This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
-The structure of this JSON object is given in figure 5.
-After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure 4.
+The structure of the JSON object is given in figure \ref{fig:JSON_runquery}.
+\item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
+\item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
+\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
+The structure of this JSON object is given in figure \ref{fig:JSON_advquery}.
+After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure \ref{fig:JSON_runquery}.
+\end{itemize}
 
@@
 The query interface is implemented in JavaScript, HTML and CSS. The main components of the user interface are as follows: 
-Initial Query Interface: The user inputs their query  term into the provided text box (figure 6).
-Add New Source: Allows the user to add a new data source to extend the query. (figure 8)
-Dynamic Facet Creator/Filter: This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure 13).
-Result Browser: The user can view their query results, organized by source.
+\begin{itemize}
+\item \textbf{Initial Query Interface:} The user inputs their query  term into the provided text box (figure \ref{fig:initial_ui}).
+\item \textbf{Add New Source:} Allows the user to add a new data source to extend the query (figure \ref{fig:add_source})
+\item \textbf{Dynamic Facet Creator/Filter:} This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure  \ref{fig:facet_creator}).
+\item \textbf{Result Browser:} The user can view their query results, organized by source.
 The results are presented in a table with pagination.
-The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure 14).
+The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure \ref{fig:results}).
+\end{itemize}
 
@@
 The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
-In the general ``Query All” case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
-The generated SPARQL query will be of the form illustrated in figure 3.
-The word “input” in the figure represents the keyword specified by the user.
+In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
+The generated SPARQL query will be of the form illustrated in figure \ref{fig:min_query}.
+The word "input" in the figure represents the keyword specified by the user.
 
-\begin{verbatim}
-SELECT ?projection_1 ?projection_2 ... ?projection_n
+\begin{figure}
+\begin{Verbatim}[frame=single]
+SELECT ?projection_1 ?projection_2 ... 
+	?projection_n
 WHERE {?x source:property ? projection 
-	FILTER regex(? projection, '" + input +"', 'i)
+	FILTER regex(? projection, '" + 
+	input +"', 'i)
 }
 ... //Other property filers are to follow
-\end{verbatim}
-Minimal SPARQL query structure in the “Query All” case
+\end{Verbatim}
+\caption{Minimal SPARQL query structure in the “Query All” case}
+\label{fig:min_query}
+\end{figure}
 
-When the results are returned, each result set is structured as the JSON object given in figure 4.
-This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables .
-how many results are returned and the result set.
+When the results are returned, each result set is structured as the JSON object given in figure \ref{fig:JSON_runquery}.
+This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
 
-When the user runs an advanced query, the payload to be sent to the proxy server to do the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All” case.
+When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All” case.
 The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
-This information is passed as a JSON object , whose structure is given in figure 5, to the proxy server.
+This information is passed as a JSON object , whose structure is given in figure \ref{fig:JSON_advquery}, to the proxy server.
 
-\begin{verbatim}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 {
@@
      		[
-     		 {"variable1": "Value 1", 		   "variable1": "Value 2"},
-     		 {"variable1": "Another Value 1", 		   "variable1": "Another Value 2"},
+     		 {"variable1": "Value 1",
+		  "variable1": "Value 2"},
+     		 {"variable1": "Another 
+		 Value 1",
+		  "variable1": "Another Value 2"},
 
@@
  }
-\end{verbatim}
-JSON returned from the RunQuery service
+\end{Verbatim}
+\caption{JSON returned from the RunQuery service}
+\label{fig:JSON_runquery}
+\end{figure}
 
-\begin{verbatim}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 {
@@
 }
-\end{verbatim}
-JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators
+\end{Verbatim}
+\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
+\label{fig:JSON_advquery}
+\end{figure}
 
@@
 In order to make QueryMed easy to use, but still provide a flexible tool that is able to perform a range of queries, we provide an uncluttered interface with only a  blank search box and two buttons.
-The ``Query All” button will perform a keyword-based query over pre-selected data sources.
-The advanced search option ``Refine Query” allows the user to view additional query options.
-The initial query interface is shown in figure 6.
+The ``Query All" button will perform a keyword-based query over pre-selected data sources.
+The advanced search option ``Refine Query" allows the user to view additional query options.
+The initial query interface is shown in figure \ref{fig:initial_ui}.
 
@@
 \caption{Initial query interface seen by the user}
+\label{fig:initial_ui}
 \end{figure}
@@
 
-The advanced query interface provides the user with the option of selecting specific data sources to query, as shown in figure 7.
-The user can also dynamically add additional data sources, as shown in figure 8.
+The advanced query interface provides the user with the option of selecting specific data sources to query, as shown in figure \ref{fig:select}.
+The user can also dynamically add additional data sources, as shown in figure \ref{fig:add_source}.
 This feature allows the user to perform their query only over relevant or trusted data sources.
@@
 \caption{The user has the option of selecting specific trusted or relevant data sources, or of adding additional data sources to query.}
+\label{fig:select}
 \end{figure}
@@
 \caption{The user can also dynamically add new data sources}
+\label{fig:add_source}
 \end{figure}
@@
 When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
-The properties list is generated by running a SPARQL query of the form given in figure 9 at the specified data source.
+The properties list is generated by running a SPARQL query of the form given in figure \ref{fig:SPARQL_props} at the specified data source.
 This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
 
-\begin{verbatim}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 SELECT DISTINCT ?property
@@
 ORDER BY ?property
-\end{verbatim}
-
-Once the properties are returned, those will be displayed in the user interface as shown in figure 10.
+\end{Verbatim}
+\caption{SPARQL query to retrieve all properties of a given datasource}
+\label{fig:SPARQL_props}
+\end{figure}
+Once the properties are returned, those will be displayed in the user interface as shown in figure \ref{fig:add_props}.
 Clicking an individual property link displays a description of the property, allowing the user to understand the keywords that will be most useful to search on each specific property.
-A sample description for the property rdfs:label is shown in figure 11.
+A sample description for the property rdfs:label is shown in figure \ref{fig:predicate_desc}.
 
@@
 \epsfig{file=images/add_props.jpg, width=3.5in}
+\label{fig:add_props}
 \caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
+\end{figure}
+
+\begin{figure}
+\centering
+\epsfig{file=images/predicate_desc.jpg, width=3.5in}
+\caption{Clicking on a property opens a window containing a description of that property}
+\label{fig:predicate_desc}
 \end{figure}
@@
 If no logical operator is specified, “AND” will be used by default.
-The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in figure 12.
+The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in figure \ref{fig:SPARQL_adv}.
 
-\begin{verbatim}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 SELECT distinct ?disease WHERE { 
@@
 ?disease 
-FILTER regex(?disease, "coronary artery disease", "i"). 
-?x <http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseasome/class> 
-<http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseaseClass/Cardiovascular>} 
-UNION { ?x <http://www4.wiwiss.fu-berlin.de/diseasome/resource/diseasome/associatedGene>
- <http://www4.wiwiss.fu-berlin.de/diseasome/resource/genes/ABCA1>.}
+FILTER regex(?disease, 
+"coronary artery disease", "i"). 
+?x <http://www4.wiwiss.fu-berlin.de/diseasome/
+resource/diseasome/class> 
+<http://www4.wiwiss.fu-berlin.de/diseasome/
+resource/diseaseClass/Cardiovascular>} 
+UNION { ?x <http://www4.wiwiss.fu-berlin.de/
+diseasome/resource/diseasome/associatedGene>
+ <http://www4.wiwiss.fu-berlin.de/diseasome/
+ resource/genes/ABCA1>.}
 }
-\end{verbatim}
-A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.
+\end{Verbatim}
+\caption{A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.}
+\label{fig:SPARQL_adv}
+\end{figure}
 
-The Dynamic Facet Creator shown in figure 13 allows the user to view properties from multiple endpoints, and to expand and contract properties for individual endpoints.
+The Dynamic Facet Creator shown in figure \ref{fig:facet_creator} allows the user to view properties from multiple endpoints, and to expand and contract properties for individual endpoints.
 This feature enables the interface to display many property lists simultaneously, grouped by source, so that only information relevant to the endpoint the user is currently examining will be visible at any time.
@@
 \caption{Dynamic Facet Creator and Filter}
+\label{fig:facet_creator}
 \end{figure}
 
-Query results are displayed to the user as a table, which can be filtered and searched to refine the query result, or printed as shown in in figure 14.
+Query results are displayed to the user as a table, which can be filtered and searched to refine the query result, or printed as shown in in figure \ref{fig:results}.
 This feature allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
@@
 \caption{A sample result table}
+\label{fig:results}
 \end{figure}
@@
 The slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
-We compared the times required to load properties from several endpoints, by running the query illustrated in figure 9.
-Timing data are shown in table 2.
+We compared the times required to load properties from several endpoints, by running the query illustrated in figure \ref{fig:SPARQL_props}.
+Timing data are shown in table \ref{timing}.
 
@@
 \caption{Running times to retrieve all the properties from selected endpoints}
+\label{timing}
 \end{table}
@@
 
+\subsection{Sample Use Case}
+
+A physician is interested in finding semantic web resources related to coronary artery disease.
+She first tries a basic search over all default resources, by entering "coronary artery disease" in the input search box.
+She then sees a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease displayed in a table.
+She can then filter the results using additional search terms.
+For example, she knows that the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term "injection."  She then prints the table of results.
+
+She is now interested in finding relevant clinical trials for her patient.
+However, the clinical trial database LinkedCt is not in the default set of endpoints, so she selects the "Refine Query" option to choose additional endpoints to search.
+She sees a list of default endpoints, and selects the "Add" option to include an additional endpoint.
+After entering the name and URL of the LinkedCT endpoint, she is able to search for clinical trials for which her patient may be eligible.
+
+She is also interested in further refining her search.
+She uses the advanced search option to construct a boolean query that takes advantage of the underlying structure of the RDF data in the database.
+She searches diseasome for a list of diseases whose class is "Cardiovascular" or for which the associated gene is "ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to her question is automatically constructed, and she can view the query results conveniently displayed in a table.
+
+A video illustrating this sample use case can be found at: \\
+http://www.youtube.com/watch?v=JCr7vryS9Vg
+
 \subsection{QueryMed Resources}
 
-The source code for the Querymed system is available at the QueryMed Google Code project: 
+The source code for the Querymed system is available at the QueryMed Google Code project: \\
 http://code.google.com/p/querymed/
-
-A video displaying a sample use case can be found at: 
-http://www.youtube.com/watch?v=JCr7vryS9Vg
 
@@
 But it is often useful to draw on multiple web resources in answering a query.
-There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambrite}systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
+There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambite} systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
 
@@
 \caption{Comparison of selected features of the QueryMed system with other related systems.}
+\label{comparison}
 \end{center}
@@
 
-Table 3 compares selected features of the QueryMed system with other related systems.
+Table \ref{comparison} compares selected features of the QueryMed system with other related systems.
 The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user.

### 2010-02-09 trunk/docs/www2010/paper.tex
added labels for the sections

@@
 %
-\title{QueryMed: An Intuitive Federated SPARQL Query Builder for Biomedical RDF Data}
+\title{QueryMed: An Intuitive SPARQL Query Builder for Biomedical RDF Data}
 %\subtitle{[Extended Abstract]
@@
 We have developed an open-source SPARQL query builder and result set visualizer for biomedical data, QueryMed, that allows end users to easily construct and run translational medicine queries across multiple data sources.
-QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, runs federated queries across multiple SPARQL endpoints, and is designed to be accessible to users who do not know the structure of the underlying ontologies or the SPARQL query language.
+
+QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, runs queries across multiple SPARQL endpoints, and is designed to be accessible to users who do not know the structure of the underlying ontologies or the SPARQL query language.
 The system allows users to select the data sources that they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.
@@
 She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.
-Although the physician has a single question--based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?--there is no single data source that the physician can use to answer this question.
+Although the physician has a single question--``based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?"--there is no single data source that the physician can use to answer this question.
 The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and GO \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
 Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
-For example, to address her question, the physician might be interested in information on “coronary artery disease” (which she thinks is the cause of her patient’s symptoms).
+For example, to address her question, the physician might be interested in information on “coronary artery disease” (which she thinks is the cause of her patient's symptoms).
 The information about the disease can be found in Diseasome.
@@
 
-This paper is organized as follows: Section 2 provides background information on the semantic web and its relevance for the biomedical domain.
-Section 3 describes our system.
-Section 4 discusses related work and illustrates how QueryMed differs from previous systems.
-Finally, section 5 outlines future work and summarizes the contributions of our system.
+This paper is organized as follows: Section \ref{background} provides background information on the semantic web and its relevance for the biomedical domain.
+Section \ref{system} describes our system.
+Section \ref{related} discusses related work and illustrates how QueryMed differs from previous systems.
+Finally, section \ref{conclusion} outlines future work and summarizes the contributions of our system.
 
 \section{Background}
-
-\subsection{Biomedical Ontologies and Semantic Web}
+\label{background}
 
@@
 \section{QueryMed}
+\label{system}
 
@@
 \section{Related Work}
+\label{related}
 
@@
 \section{Future Work and Conclusions}
+\label{conclusion}
 

### 2010-02-09 trunk/docs/www2010/paper.tex
fixed some minor formatting errors

@@
 
-The semantic web can be viewed as global database system for the data available on the world wide web.
-Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using a standardized query language, SPARQL \cite{Berners-Lee1}.
+The semantic web can be viewed as a global database system for the data available on the world wide web.
+Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using the SPARQL query language \cite{Berners-Lee1}.
 The addition of structure to web data allows inferences to be automatically drawn by intelligent agents integrating data from multiple sources.
-
 Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
-Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
+Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF.
 Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists, and these semantic web resources represent valuable standardized repositories of information that can be automatically mined for biological knowledge.
-The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
+%The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
 
@@
 \epsfig{file=images/architecture_detail.jpg}
-\caption{QueryMed Architecture details}
+\caption{QueryMed Architecture Details}
 \label{fig:arch_details}
@@
 \item \textbf{Translator:} This component is responsible for translating the user query into valid SPARQL syntax.
-\item \textbf{Run Query:} The purpose of this service is to run SPARQL queries.
+\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries.
 It takes input on source and keyword from the client, and returns  the query results as a JSON object.
@@
 
-\subsubsection{Data Structures}
-
-The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
-In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
-The generated SPARQL query will be of the form illustrated in figure \ref{fig:min_query}.
-The word "input" in the figure represents the keyword specified by the user.
-
 \begin{figure}
@@
 \end{Verbatim}
-\caption{Minimal SPARQL query structure in the “Query All” case}
+\caption{Minimal SPARQL Query Structure in the ``Query All" Case}
 \label{fig:min_query}
 \end{figure}
-
-When the results are returned, each result set is structured as the JSON object given in figure \ref{fig:JSON_runquery}.
-This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
-
-When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All” case.
-The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
-This information is passed as a JSON object , whose structure is given in figure \ref{fig:JSON_advquery}, to the proxy server.
 
@@
 
+\subsubsection{Data Structures}
+
+The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
+In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
+The generated SPARQL query will be of the form illustrated in figure \ref{fig:min_query}.
+The word ``input" in the figure represents the keyword specified by the user.
+
+When the results are returned, each result set is structured as the JSON object given in figure \ref{fig:JSON_runquery}.
+This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
+
+When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All" case.
+The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
+This information is passed as a JSON object, whose structure is given in figure \ref{fig:JSON_advquery}, to the proxy server.
+
 \subsubsection{Query Interface}
@@
 
-When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
-The properties list is generated by running a SPARQL query of the form given in figure \ref{fig:SPARQL_props} at the specified data source.
-This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
-
 \begin{figure}
@@
 \end{figure}
+
+\begin{figure*}
+\centering
+\epsfig{file=images/add_props.jpg, width=6in}
+\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
+\label{fig:add_props}
+\end{figure*}
+
+When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
+The properties list is generated by running a SPARQL query of the form given in figure \ref{fig:SPARQL_props} at the specified data source.
+This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
+
 Once the properties are returned, those will be displayed in the user interface as shown in figure \ref{fig:add_props}.
 Clicking an individual property link displays a description of the property, allowing the user to understand the keywords that will be most useful to search on each specific property.
-A sample description for the property rdfs:label is shown in figure \ref{fig:predicate_desc}.
+%A sample description for the property rdfs:label is shown in figure \ref{fig:predicate_desc}.
 
-\begin{figure}
-\centering
-\epsfig{file=images/add_props.jpg, width=3.5in}
-\label{fig:add_props}
-\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=images/predicate_desc.jpg, width=3.5in}
-\caption{Clicking on a property opens a window containing a description of that property}
-\label{fig:predicate_desc}
-\end{figure}
+%\begin{figure}
+%\centering
+%\epsfig{file=images/predicate_desc.jpg, width=3.5in}
+%\caption{Clicking on a property opens a window containing a description of that property}
+%\label{fig:predicate_desc}
+%\end{figure}
 
@@
 \hline
- & Dieseasome & Dailymed & Drugbank \\
+ & \textbf{Dieseasome} & \textbf{Dailymed} & \textbf{Drugbank} \\
 \hline
 \hline
-1st Trial   & 3.45 s & 1.77 s  & 9.51 s  \\
-2nd Trial  & 1.61 s & 1.57 s & 9.34 s \\
-3rf Trial  & 1.71 s & 1.66 s & 9.06 s \\
+\textbf{1st Trial}   & 3.45 s & 1.77 s  & 9.51 s  \\
+\textbf{2nd Trial}  & 1.61 s & 1.57 s & 9.34 s \\
+\textbf{3rd Trial}  & 1.71 s & 1.66 s & 9.06 s \\
 \hline
@@
 
-\subsection{Sample Use Case}
+\subsection{A Sample Use Case}
 
 A physician is interested in finding semantic web resources related to coronary artery disease.
-She first tries a basic search over all default resources, by entering "coronary artery disease" in the input search box.
+She first tries a basic search over all default resources, by entering ``coronary artery disease" in the input search box.
 She then sees a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease displayed in a table.
 She can then filter the results using additional search terms.
-For example, she knows that the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term "injection."  She then prints the table of results.
+For example, she knows that the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term ``injection."  She then prints the table of results.
 
 She is now interested in finding relevant clinical trials for her patient.
-However, the clinical trial database LinkedCt is not in the default set of endpoints, so she selects the "Refine Query" option to choose additional endpoints to search.
-She sees a list of default endpoints, and selects the "Add" option to include an additional endpoint.
+However, the clinical trial database LinkedCt is not in the default set of endpoints, so she selects the ``Refine Query" option to choose additional endpoints to search.
+She sees a list of default endpoints, and selects the ``Add" option to include an additional endpoint.
 After entering the name and URL of the LinkedCT endpoint, she is able to search for clinical trials for which her patient may be eligible.
@@
 She uses the advanced search option to construct a boolean query that takes advantage of the underlying structure of the RDF data in the database.
-She searches diseasome for a list of diseases whose class is "Cardiovascular" or for which the associated gene is "ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to her question is automatically constructed, and she can view the query results conveniently displayed in a table.
+She searches diseasome for a list of diseases whose class is ``Cardiovascular" or for which the associated gene is ``ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to her question is automatically constructed, and she can view the query results conveniently displayed in a table.
 
@@
 \label{related}
+
+\begin{table*}
+\begin{center}
+\begin{tabular}{| p{2cm} | p{2cm}| p{2cm}| p{2cm}| p{2cm}| p{2cm} |}
+\hline
+ & \textbf{Query Multiple Sources?} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
+\hline
+\hline
+\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes  \\
+\hline
+\textbf{SMART}  & Yes & No & No & Yes & Yes  \\
+\hline
+\textbf{DARQ}  & Yes & N/A & No & Yes & No  \\
+\hline
+\textbf{GoWeb}  & Yes & No & Yes & No & Yes  \\
+\hline
+\textbf{BioGateway}  & Yes & No & No & No & Yes  \\
+\hline
+\textbf{Twinkle}  & Yes & No & No & Yes & Yes  \\
+\hline
+\end{tabular}
+\caption{Comparison of selected features of the QueryMed system with other related systems.}
+\label{comparison}
+\end{center}
+\end{table*}
 
@@
 There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambite} systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
-
-\begin{table*}
-\begin{center}
-\begin{tabular}{| p{2cm} | p{2cm}| p{2cm}| p{2cm}| p{2cm}| p{2cm} |}
-\hline
- & Query Multiple Sources? & Dynamic Addition of Sources & Allows Keyword Queries & Open Source & GUI \\
-\hline
-\hline
-QueryMed  & Yes & Yes & Yes & Yes & Yes  \\
-\hline
-SMART  & Yes & No & No & Yes & Yes  \\
-\hline
-DARQ  & Yes & N/A & No & Yes & No  \\
-\hline
-GoWeb  & Yes & No & Yes & No & Yes  \\
-\hline
-BioGateway  & Yes & No & No & No & Yes  \\
-\hline
-Twinkle  & Yes & No & No & Yes & Yes  \\
-\hline
-\end{tabular}
-\caption{Comparison of selected features of the QueryMed system with other related systems.}
-\label{comparison}
-\end{center}
-\end{table*}
 

### 2010-02-11 docs/www2010/paper.tex
integrated some of Ian's comments

@@
 \documentclass{www2010-submission}
-\usepackage{moreverb}
+\usepackage{fancyvrb}
 
@@
 
-The biomedical domain is among the early successes of the semantic web, due to the rapidity with which the biomedical community has made its data available in RDF triple stores.
-However, although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
+The biomedical domain is among the early successes of the semantic web, due to the rapidity with which the biomedical community has made its data available in RDF triple stores \cite{Yip}.
+Although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and also allow users to run federated queries on multiple SPARQL endpoints.
 
@@
 She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.
-Although the physician has a single question--based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?--there is no single data source that the physician can use to answer this question.
+Although the physician has a single question--``based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?"--there is no single data source that the physician can use to answer this question.
 The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and GO \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
 Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
-For example, to address her question, the physician might be interested in information on “coronary artery disease” (which she thinks is the cause of her patient’s symptoms).
-The information about the disease can be found in Diseasome.
-She might also want to know the set of drugs in DailyMed and Drugbank that can be used to treat coronary artery disease.
-Information on clinical trials available for her patient can be found in LinkedCT.
+
+Suppose the physician might be interested in information on “coronary artery disease” (which she thinks is the cause of her patient's symptoms).
+The information about the disease can be found in Diseasome, a database that lists diseases, disease genes, and associations between them~\cite{Diseasome}.
+The physician might also want to know the set of drugs in DailyMed~\cite{DailyMed} and Drugbank~\cite{Drugbank} that can be used to treat coronary artery disease.
+Information on clinical trials available for her patient can be found in LinkedCT~\cite{LinkedCT}.
 Due to the large number of databases  that the physician needs to query in order to find an answer to her single question, she is likely to find useful a system that can automatically run queries over multiple data sources.
@@
 Developing intuitive ways to query multiple data sources and display results is both an important and a challenging problem.
+Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.
 
-This paper is organized as follows: Section 2 provides background information on the semantic web and its relevance for the biomedical domain.
-Section 3 describes our system.
-Section 4 discusses related work and illustrates how QueryMed differs from previous systems.
-Finally, section 5 outlines future work and summarizes the contributions of our system.
+This paper is organized as follows: Section \ref{background} provides background information on the semantic web and its relevance for the biomedical domain.
+Section \ref{system} describes our system.
+Section \ref{related} discusses related work and illustrates how QueryMed differs from previous systems.
+Finally, section \ref{conclusion} outlines future work and summarizes the contributions of our system.
 
 \section{Background}
+\label{background}
 
-\subsection{Biomedical Ontologies and Semantic Web}
+The semantic web can be viewed as a global database system for the data available on the world wide web.
+Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using the SPARQL query language.
+The addition of structure to web data allows inferences to be automatically drawn by intelligent agents integrating data from multiple sources~\cite{Berners-Lee1}.
+%Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
 
-Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
-Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
-Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists, and these semantic web resources represent valuable standardized repositories of information that can be automatically mined for biological knowledge.
-The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
+Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF.
+Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists.
+These semantic web resources represent valuable repositories of information that can be automatically mined for applications that require biological knowledge.
+%The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
 
-However, although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources of maximal use to physicians, patients, and life scientists.
+Although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources accessible to physicians, patients, and life scientists.
 One challenge is constructing systems that allow end users to run intuitive queries on biomedical data.
 Users of biomedical resources are likely to have extensive domain knowledge, but be unfamiliar with the SPARQL query language syntax and with the structure of biomedical ontologies.
-It is important to design user-friendly systems that allow these users to take advantage of the wealth of structured biological knowledge available on the semantic web.
+Almost all the SPARQL endpoints available today are offering a generic query interface that requires users to write SPARQL queries.
+Also, no hints about the structure of data available from the SPARQL endpoint is given as the user is constructing the SPARQL query.
+The user has to essentially construct the queries that are needed for her purpose by trial and error.
+This can especially be a daunting task for someone who is new to semantic web technologies.
+Therefore, it is important to design user-friendly systems that allow users to take advantage of the wealth of structured biological knowledge available on the semantic web.
 Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data is often distributed among many sources \cite{Pasquier}.
@@
 \section{QueryMed}
+\label{system}
+
+\begin{figure}
+\centering
+\epsfig{file=images/architecture_overview.jpg}
+\caption{QueryMed Architecture Overview}
+\label{fig:arch_overview}
+\end{figure}
+
+QueryMed allows users to query many biomedical data sources from a single interface.
+The data sources that the user wishes to query are not fixed, and can be extended very easily.
+The system was designed with non-expert users of SPARQL in mind, and presents a very intuitive user interface that anybody with basic biomedical domain knowledge can use.
+
+In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind some of the crucial components of the system.
 
@@
 
-A general overview of QueryMed architecture is shown in Figure 1.
+A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
 The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.
-
-QueryMed is implemented in Java in the backend and JavaScript in the frontend.
-The JQuery library \cite{Resig} was used to develop an attractive user interface.
-In the backend, the Jena library \cite{McBride} is used to run the SPARQL queries.
 QueryMed relies on the following remote resources:
@@
 \centering
-\epsfig{file=images/architecture_overview.jpg}
-\caption{QueryMed Architecture Overview}
-\end{figure}
-
-\begin{figure}
-\centering
 \epsfig{file=images/architecture_detail.jpg}
-\caption{QueryMed Architecture details}
+\caption{QueryMed Architecture Details}
+\label{fig:arch_details}
 \end{figure}
@@
 \item \textbf{Translator:} This component is responsible for translating the user query into valid SPARQL syntax.
-\item \textbf{Run Query:} The purpose of this service is to run SPARQL queries.
+\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries.
 It takes input on source and keyword from the client, and returns  the query results as a JSON object.
-The structure of the JSON object is given in figure 4.
+The structure of the JSON object is given in figure \ref{fig:JSON_runquery}.
 \item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
@@
 \item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
-The structure of this JSON object is given in figure 5.
-After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure 4.
+The structure of this JSON object is given in figure \ref{fig:JSON_advquery}.
+After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure \ref{fig:JSON_runquery}.
 \end{itemize}
@@
 \begin{itemize}
-\item \textbf{Initial Query Interface:} The user inputs their query  term into the provided text box (figure 6).
-\item \textbf{Add New Source:} Allows the user to add a new data source to extend the query. (figure 8)
-\item \textbf{Dynamic Facet Creator/Filter:} This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure 13).
+\item \textbf{Initial Query Interface:} The user inputs their query  term into the provided text box (figure \ref{fig:initial_ui}).
+\item \textbf{Add New Source:} Allows the user to add a new data source to extend the query (figure \ref{fig:add_source})
+\item \textbf{Dynamic Facet Creator/Filter:} This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure  \ref{fig:facet_creator}).
 \item \textbf{Result Browser:} The user can view their query results, organized by source.
 The results are presented in a table with pagination.
-The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure 14).
+The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure \ref{fig:results}).
 \end{itemize}
@@
 
-\subsubsection{Data Structures}
-
-The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
-In the general ``Query All” case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
-The generated SPARQL query will be of the form illustrated in figure 3.
-The word “input” in the figure represents the keyword specified by the user.
-
-\begin{verbatimtab}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 SELECT ?projection_1 ?projection_2 ... 
@@
 ... //Other property filers are to follow
-\end{verbatimtab}
-Minimal SPARQL query structure in the “Query All” case
+\end{Verbatim}
+\caption{Minimal SPARQL Query Structure in the ``Query All" Case}
+\label{fig:min_query}
+\end{figure}
 
-When the results are returned, each result set is structured as the JSON object given in figure 4.
-This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables .
-how many results are returned and the result set.
-
-When the user runs an advanced query, the payload to be sent to the proxy server to do the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All” case.
-The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
-This information is passed as a JSON object , whose structure is given in figure 5, to the proxy server.
-
-\begin{verbatimtab}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 {
@@
 		  "variable1": "Value 2"},
-     		 {"variable1": "Another Value 1",
+     		 {"variable1": "Another 
+		 Value 1",
 		  "variable1": "Another Value 2"},
@@
  }
-\end{verbatimtab}
-JSON returned from the RunQuery service
+\end{Verbatim}
+\caption{JSON returned from the RunQuery service}
+\label{fig:JSON_runquery}
+\end{figure}
 
-\begin{verbatimtab}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 {
@@
 }
-\end{verbatimtab}
-JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators
+\end{Verbatim}
+\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
+\label{fig:JSON_advquery}
+\end{figure}
+
+\subsubsection{Data Structures}
+
+The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
+In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
+The generated SPARQL query will be of the form illustrated in figure \ref{fig:min_query}.
+The word ``input" in the figure represents the keyword specified by the user.
+
+When the results are returned, each result set is structured as the JSON object given in figure \ref{fig:JSON_runquery}.
+This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
+
+When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All" case.
+The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
+This information is passed as a JSON object, whose structure is given in figure \ref{fig:JSON_advquery}, to the proxy server.
 
@@
 \caption{The user can also dynamically add new data sources}
+\label{fig:add_source}
 \end{figure}
 
-When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
-The properties list is generated by running a SPARQL query of the form given in figure 9 at the specified data source.
-This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
-
-\begin{verbatimtab}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 SELECT DISTINCT ?property
@@
 ORDER BY ?property
-\end{verbatimtab}
+\end{Verbatim}
+\caption{SPARQL query to retrieve all properties of a given datasource}
+\label{fig:SPARQL_props}
+\end{figure}
 
-Once the properties are returned, those will be displayed in the user interface as shown in figure 10.
+\begin{figure*}
+\centering
+\epsfig{file=images/add_props.jpg, width=6in}
+\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
+\label{fig:add_props}
+\end{figure*}
+
+When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
+The properties list is generated by running a SPARQL query of the form given in figure \ref{fig:SPARQL_props} at the specified data source.
+This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
+
+Once the properties are returned, those will be displayed in the user interface as shown in figure \ref{fig:add_props}.
 Clicking an individual property link displays a description of the property, allowing the user to understand the keywords that will be most useful to search on each specific property.
-A sample description for the property rdfs:label is shown in figure 11.
+%A sample description for the property rdfs:label is shown in figure \ref{fig:predicate_desc}.
 
-\begin{figure}
-\centering
-\epsfig{file=images/add_props.jpg, width=3.5in}
-\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
-\end{figure}
+%\begin{figure}
+%\centering
+%\epsfig{file=images/predicate_desc.jpg, width=3.5in}
+%\caption{Clicking on a property opens a window containing a description of that property}
+%\label{fig:predicate_desc}
+%\end{figure}
 
@@
 If no logical operator is specified, “AND” will be used by default.
-The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in figure 12.
+The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in figure \ref{fig:SPARQL_adv}.
 
-\begin{verbatimtab}
+\begin{figure}
+\begin{Verbatim}[frame=single]
 SELECT distinct ?disease WHERE { 
@@
 }
-\end{verbatimtab}
-A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.
+\end{Verbatim}
+\caption{A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.}
+\label{fig:SPARQL_adv}
+\end{figure}
 
-The Dynamic Facet Creator shown in figure 13 allows the user to view properties from multiple endpoints, and to expand and contract properties for individual endpoints.
+The Dynamic Facet Creator shown in figure \ref{fig:facet_creator} allows the user to view properties from multiple endpoints, and to expand and contract properties for individual endpoints.
 This feature enables the interface to display many property lists simultaneously, grouped by source, so that only information relevant to the endpoint the user is currently examining will be visible at any time.
@@
 \caption{A sample result table}
+\label{fig:results}
 \end{figure}
+
+\subsubsection{Implementation}
+
+QueryMed is implemented in Java in the backend and JavaScript in the frontend.
+The JQuery library \cite{Resig} was used to develop an attractive user interface.
+In the backend, the Jena library \cite{McBride} is used to run the SPARQL queries.
+
+\subsection{A Sample Use Case}
+
+A physician is interested in finding semantic web resources related to coronary artery disease.
+She first tries a basic search over all default resources, by entering ``coronary artery disease" in the input search box.
+She then sees a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease displayed in a table.
+She can then filter the results using additional search terms.
+For example, she knows that the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term ``injection."  She then prints the table of results.
+
+She is now interested in finding relevant clinical trials for her patient.
+However, the clinical trial database LinkedCt is not in the default set of endpoints, so she selects the ``Refine Query" option to choose additional endpoints to search.
+She sees a list of default endpoints, and selects the ``Add" option to include an additional endpoint.
+After entering the name and URL of the LinkedCT endpoint, she is able to search for clinical trials for which her patient may be eligible.
+
+She is also interested in further refining her search.
+She uses the advanced search option to construct a boolean query that takes advantage of the underlying structure of the RDF data in the database.
+She searches diseasome for a list of diseases whose class is ``Cardiovascular" or for which the associated gene is ``ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to her question is automatically constructed, and she can view the query results conveniently displayed in a table.
+
+A video illustrating this sample use case can be found at: \\
+http://www.youtube.com/watch?v=JCr7vryS9Vg
 
@@
 
-A video displaying a sample use case can be found at: \\
-http://www.youtube.com/watch?v=JCr7vryS9Vg
+\section{Related Work}
+\label{related}
 
-\section{Related Work}
+\begin{table*}
+\begin{center}
+\begin{tabular}{| p{2cm} | p{2cm}| p{2cm}| p{2cm}| p{2cm}| p{2cm} |}
+\hline
+ & \textbf{Query Multiple Sources?} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
+\hline
+\hline
+\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes  \\
+\hline
+\textbf{SMART}  & Yes & No & No & Yes & Yes  \\
+\hline
+\textbf{DARQ}  & Yes & N/A & No & Yes & No  \\
+\hline
+\textbf{GoWeb}  & Yes & No & Yes & No & Yes  \\
+\hline
+\textbf{BioGateway}  & Yes & No & No & No & Yes  \\
+\hline
+\textbf{Twinkle}  & Yes & No & No & Yes & Yes  \\
+\hline
+\end{tabular}
+\caption{Comparison of selected features of the QueryMed system with other related systems.}
+\label{comparison}
+\end{center}
+\end{table*}
 
@@
 But it is often useful to draw on multiple web resources in answering a query.
-There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambrite}systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
+There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambite} systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
 
-\begin{table*}
-\begin{center}
-\begin{tabular}{| p{2cm} | p{2cm}| p{2cm}| p{2cm}| p{2cm}| p{2cm} |}
-\hline
- & Query Multiple Sources? & Dynamic Addition of Sources & Allows Keyword Queries & Open Source & GUI \\
-\hline
-\hline
-QueryMed  & Yes & Yes & Yes & Yes & Yes  \\
-\hline
-SMART  & Yes & No & No & Yes & Yes  \\
-\hline
-DARQ  & Yes & N/A & No & Yes & No  \\
-\hline
-GoWeb  & Yes & No & Yes & No & Yes  \\
-\hline
-BioGateway  & Yes & No & No & No & Yes  \\
-\hline
-Twinkle  & Yes & No & No & Yes & Yes  \\
-\hline
-\end{tabular}
-\caption{Comparison of selected features of the QueryMed system with other related systems.}
-\end{center}
-\end{table*}
-
-Table 3 compares selected features of the QueryMed system with other related systems.
+Table \ref{comparison} compares selected features of the QueryMed system with other related systems.
 The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user.

### 2010-02-11 docs/www2010/paper.tex
added a new reference, and restructured the paper - not finished yet

@@
 
+\cite{Jentzsch}
+
 This paper is organized as follows: Section \ref{background} provides background information on the semantic web and its relevance for the biomedical domain.
@@
 
-\section{QueryMed}
+\section{System Design \& Implementation}
 \label{system}
@@
 
+The system functionality in a nutshell is as follows: A user submits a query from the Query Interface to the SPARQL endpoints via the proxy server.
+The query is translated into a SPARQL query for each individual endpoint.
+Once the results are obtained from the SPARQL endpoints, they are presented to the user via the Result Browser.
+The user can then choose to refine the query by changing some of the original query terms in the Dynamic Facet Creator/Filter, and this process repeats.
+
 In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind some of the crucial components of the system.
 
-\subsection{Architecture overview}
+\subsection{QueryMed Overview}
 
@@
 The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.
-QueryMed relies on the following remote resources:
-\begin{itemize}
-\item \textbf{SPARQL Endpoints}:  QueryMed runs queries on multiple remote SPARQL endpoints that expose biomedical data.
-\item \textbf{Sources List}: The default list of endpoints available to the user is stored in a source list on the proxy server.
-\end{itemize}
+
+A more detailed illustration of the QueryMed system highlighting all the components necessary for system functionality is given in Figure \ref{fig:arch_details}.
+The system relies on two external resources:
+
+\begin{enumerate}
+\item \textbf{SPARQL endpoints} that expose biomedical data.
+\item \textbf{Sources List} that gives a list of endpoints available to the user. This list is stored on the proxy server, but can be replaced by the user. 
+\end{enumerate}
 
@@
 
+\subsubsection{User Interface}
+
+The user interface is one of the primary components of the QueryMed system. It comprises of the following:
+
+\begin{itemize}
+
+\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but still provide a flexible tool that is able to perform a range of queries, we provide an uncluttered interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
+The ``Query All" button will perform a keyword-based query over pre-selected data sources.
+The advanced search option ``Refine Query" allows the user to view additional query options that are handled in the Advanced Query Interface.
+
+\begin{figure}
+\centering
+\epsfig{file=images/new_front.jpg,  width=3in}
+\caption{Initial query interface seen by the user}
+\label{fig:initial_ui}
+\end{figure}
+
+\item \textbf{Advanced Query Interface:} This allows the user to add new data sources, and give values for properties found at a particular selected data source.
+Therefore, the user is able to extend the query from a basic keyword search over all the properties as with the basic query interface, to a more targeted query.
+
+When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources in the system as shown in Figure \ref{fig:select}.
+This feature allows the user to perform their query only over relevant or trusted data sources.
+If none of the desired data sources are available in this list, the user can dynamically add additional data sources as shown in Figure \ref{fig:add_source}.
+The ability to dynamically add additional sources increases the flexibility of  the QueryMed system, allowing users to query endpoints of interest that are not included in the default list.
+
+\begin{figure}
+\centering
+\epsfig{file=images/select.jpg,  width=2.5in}
+\caption{The user has the option of selecting specific trusted or relevant data sources, or of adding additional data sources to query.}
+\label{fig:select}
+\end{figure}
+
+\begin{figure}
+\centering
+\epsfig{file=images/add_source.jpg,  width=3in}
+\caption{The user can also dynamically add new data sources}
+\label{fig:add_source}
+\end{figure}
+
+When a data source is selected, the properties list is automatically populated with all the distinct properties available at the selected endpoint.
+If the user wants to restrict her query based on one or more properties available, she can specify those.
+This allows users to improve their queries using the underlying structure of the data without having prior knowledge of this structure to obtain more relevant query results.
+%The properties list is generated by running a SPARQL query of the form given in Figure \ref{fig:SPARQL_props} at the specified data source.
+
+%\begin{figure}
+%\begin{Verbatim}[frame=single]
+%SELECT DISTINCT ?property
+%WHERE { [] ?property [] }
+%ORDER BY ?property
+%\end{Verbatim}
+%\caption{SPARQL query to retrieve all properties of a given datasource}
+%\label{fig:SPARQL_props}
+%\end{figure}
+
+Once the properties are returned, they will be displayed in the user interface as shown in Figure \ref{fig:add_props}.
+Clicking an individual property link displays a description of the property.
+This allows the user to understand the keywords that will be most useful to search on each specific property.
+
+\begin{figure*}
+\centering
+\epsfig{file=images/add_props.jpg, width=6in}
+\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
+\label{fig:add_props}
+\end{figure*}
+
+\item \textbf{Dynamic Facet Creator/Filter:} This allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint as shown in Figure  \ref{fig:facet_creator}.
+The interface allows displaying of many property lists from different sources simultaneously (grouped by source).
+Therefore, only information relevant to the endpoint the user is currently examining will be visible at any given time.
+
+\begin{figure}
+\centering
+\epsfig{file=images/facet_creator.jpg, width=3in}
+\caption{Dynamic Facet Creator and Filter}
+\label{fig:facet_creator}
+\end{figure}
+
+\item \textbf{Result Browser:} The user can view their query results organized by source in the Result Browser.
+The results are presented in a table with pagination.
+The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
+This feature allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+
+\begin{figure}
+\centering
+\epsfig{file=images/results.jpg, width=3.5in}
+\caption{A sample result table}
+\label{fig:results}
+\end{figure}
+
+\end{itemize}
+
 \subsubsection{Proxy Server}
 
-The user submits a query to the SPARQL endpoints via the proxy server.
-The query is translated into a SPARQL query for each individual endpoint, and the results are returned to the user.
-The specific components implemented in the proxy server are as follows.
+The functionality of the proxy server is to act as an intermediary between the user interface and the remote SPARQL endpoints.  Its functionality is twofold:
+
+\begin{enumerate}
+\item Execute the SPARQL queries at the relevant remote SPARQL endpoints and consolidate the results to be presented in the user interface.
+\item Cache the results for subsequent queries for refinements to the same query to reduce the network and query execution latency.
+\end{enumerate}
+
+The specific components in the Proxy Server are as follows.
+
 \begin{itemize}
-\item \textbf{Source Manager:} The source manager reads the source list and populates the default query list on the user interface.
+
+\item \textbf{Source Manager:} The Source Manager reads the source list and populates the default query list on the user interface.
 It also keeps track of the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
-\item \textbf{Translator:} This component is responsible for translating the user query into valid SPARQL syntax.
-\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries.
-It takes input on source and keyword from the client, and returns  the query results as a JSON object.
-The structure of the JSON object is given in figure \ref{fig:JSON_runquery}.
+
+\item \textbf{Translator:} The Translator is responsible for translating the user query into valid SPARQL syntax.
+Translator obtains the parameters to construct the query from the input the user specifies in the Query Interface, and plugs them in to a relevant template.
+It specifically relies on two services:
+
+\begin{enumerate}
+
+\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries generated by the Translator, and return  the query results as a JSON object given in Figure \ref{fig:JSON_runquery}.
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
+{
+   "bindings":
+     [
+        {"source" : "source_label",
+         "uri" : "http://source.com/sparql.",
+     	"vars" : ["variable1","variable2"],
+     	"count":2,
+     	"results" : 
+     	 [
+     	    {"variable1": "Value 1",
+	      "variable1": "Value 2"},
+     	    {"variable1": "Another Value 1",
+               "variable1": "Another Value 2"},
+   	    ...	  		
+     	 ]
+     	}
+     	,
+	{ //Another Source and Result Description
+	...
+	},
+	...
+     ]
+ }
+\end{Verbatim}
+\caption{Structure of the JSON ``result" object}
+\label{fig:JSON_runquery}
+\end{figure}
+
+\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object given in Figure \ref{fig:JSON_advquery}.
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
+{
+ "http://some-source/sparql": 
+  {
+     "property1": {"value1":  "logical operator"},
+     "property2": {"value 2": "logical operator"},
+     "property2": {"value 3": "logical operator"}
+  }
+}
+   //Other sources ...
+
+\end{Verbatim}
+\caption{Structure of the JSON ``query" object}
+%\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
+\label{fig:JSON_advquery}
+\end{figure}
+
+\end{enumerate}
+
+% The user has to specify the data source(s) to query, and can give specific values for one or many of the properties available for the data available to restrict the query space. 
+\item \textbf{Additional Services:} In order to make the user interface more user friendly, the following two services perform some ``behind the scenes" work.
+
+\begin{enumerate}
 \item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
 \item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
-\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
-The structure of this JSON object is given in figure \ref{fig:JSON_advquery}.
-After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure \ref{fig:JSON_runquery}.
+\end{enumerate}
+
 \end{itemize}
 
-\subsubsection{User Interface}
-
-The query interface is implemented in JavaScript, HTML and CSS. The main components of the user interface are as follows: 
-\begin{itemize}
-\item \textbf{Initial Query Interface:} The user inputs their query  term into the provided text box (figure \ref{fig:initial_ui}).
-\item \textbf{Add New Source:} Allows the user to add a new data source to extend the query (figure \ref{fig:add_source})
-\item \textbf{Dynamic Facet Creator/Filter:} This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure  \ref{fig:facet_creator}).
-\item \textbf{Result Browser:} The user can view their query results, organized by source.
-The results are presented in a table with pagination.
-The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure \ref{fig:results}).
-\end{itemize}
-
-\subsection{System Functionality and Design Decisions}
+\subsection{Design Decisions}
 
@@
 
-\begin{figure}
-\begin{Verbatim}[frame=single]
-{
-     "bindings":
-     [
-     	{"source" : "source_label",
-     	 "uri" : "http://source.com/sparql.",
-     	 "vars" : ["variable1","variable2"],
-     	 "count":2,
-     	 "results" : 
-     		[
-     		 {"variable1": "Value 1",
-		  "variable1": "Value 2"},
-     		 {"variable1": "Another 
-		 Value 1",
-		  "variable1": "Another Value 2"},
-
-     		]
-     	}
-     	,
-	{ //Another Source and Result Description
-	...
-	},
-	...
-     ]
- }
-\end{Verbatim}
-\caption{JSON returned from the RunQuery service}
-\label{fig:JSON_runquery}
-\end{figure}
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-{
-   "http://some-source/sparql": {
-   	"property1": {"value1": "AND"},
-	"property2": {"value 2": "OR"},
-	"property2": {"value 3": "FILTER"}
-	           }
-	}
-   //Other sources ...
-}
-\end{Verbatim}
-\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
-\label{fig:JSON_advquery}
-\end{figure}
-
 \subsubsection{Data Structures}
@@
 
-In order to make QueryMed easy to use, but still provide a flexible tool that is able to perform a range of queries, we provide an uncluttered interface with only a  blank search box and two buttons.
-The ``Query All" button will perform a keyword-based query over pre-selected data sources.
-The advanced search option ``Refine Query" allows the user to view additional query options.
-The initial query interface is shown in figure \ref{fig:initial_ui}.
-
-\begin{figure}
-\centering
-\epsfig{file=images/new_front.jpg,  width=3in}
-\caption{Initial query interface seen by the user}
-\label{fig:initial_ui}
-\end{figure}
-
 We believe that the unstructured keyword search is likely to encourage users without previous experience running queries on semantic web data to explore the additional features of our system that allow more flexible and advanced queries.
 For example, physicians who are not used to running queries on semantic web data may be more willing to invest time into learning how to find better search results once they see that the system can easily provide them with some initial useful query results using the basic keyword search option.
-However, the execution of queries using QueryMed will be very much different from a user issuing a web search on a search engine such as Google, because it exploits the underlying structure of the data as exposed in the individual data endpoints.
-
-The advanced query interface provides the user with the option of selecting specific data sources to query, as shown in figure \ref{fig:select}.
-The user can also dynamically add additional data sources, as shown in figure \ref{fig:add_source}.
-This feature allows the user to perform their query only over relevant or trusted data sources.
-The ability to dynamically add additional sources increases the flexibility of  the QueryMed system, allowing users to search endpoints of interest that are not included in the default list.
-
-\begin{figure}
-\centering
-\epsfig{file=images/select.jpg,  width=3in}
-\caption{The user has the option of selecting specific trusted or relevant data sources, or of adding additional data sources to query.}
-\label{fig:select}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=images/add_source.jpg,  width=3in}
-\caption{The user can also dynamically add new data sources}
-\label{fig:add_source}
-\end{figure}
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-SELECT DISTINCT ?property
-WHERE { [] ?property [] }
-ORDER BY ?property
-\end{Verbatim}
-\caption{SPARQL query to retrieve all properties of a given datasource}
-\label{fig:SPARQL_props}
-\end{figure}
-
-\begin{figure*}
-\centering
-\epsfig{file=images/add_props.jpg, width=6in}
-\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
-\label{fig:add_props}
-\end{figure*}
-
-When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
-The properties list is generated by running a SPARQL query of the form given in figure \ref{fig:SPARQL_props} at the specified data source.
-This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
-
-Once the properties are returned, those will be displayed in the user interface as shown in figure \ref{fig:add_props}.
-Clicking an individual property link displays a description of the property, allowing the user to understand the keywords that will be most useful to search on each specific property.
-%A sample description for the property rdfs:label is shown in figure \ref{fig:predicate_desc}.
-
-%\begin{figure}
-%\centering
-%\epsfig{file=images/predicate_desc.jpg, width=3.5in}
-%\caption{Clicking on a property opens a window containing a description of that property}
-%\label{fig:predicate_desc}
-%\end{figure}
+However, the execution of queries using QueryMed will be very different from a user issuing a web search on a search engine such as Google because it exploits the underlying structure of the data exposed by the individual data endpoints.
 
@@
 
-The Dynamic Facet Creator shown in figure \ref{fig:facet_creator} allows the user to view properties from multiple endpoints, and to expand and contract properties for individual endpoints.
-This feature enables the interface to display many property lists simultaneously, grouped by source, so that only information relevant to the endpoint the user is currently examining will be visible at any time.
-
-\begin{figure}
-\centering
-\epsfig{file=images/facet_creator.jpg, width=3in}
-\caption{Dynamic Facet Creator and Filter}
-\label{fig:facet_creator}
-\end{figure}
-
-Query results are displayed to the user as a table, which can be filtered and searched to refine the query result, or printed as shown in in figure \ref{fig:results}.
-This feature allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
-
-\begin{figure}
-\centering
-\epsfig{file=images/results.jpg, width=3.5in}
-\caption{A sample result table}
-\label{fig:results}
-\end{figure}
-
 \subsubsection{Implementation}
 
-QueryMed is implemented in Java in the backend and JavaScript in the frontend.
+QueryMed is implemented in Java in the backend and JavaScript, HTML and CSS in the frontend.
 The JQuery library \cite{Resig} was used to develop an attractive user interface.
@@
 
+For all selected endpoints, the property values take longer to load on the initial query than in subsequent iterations, probably due to browser caching.
+The running time to acquire the list of property values for the Drugbank endpoint is much larger than the time for the other two endpoints, probably due to the greater size of the Drugbank database.
+(As of 12/10/09, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, and DrugBank contains 765,936 triples).
+
 \begin{table}
@@
 \end{table}
-
-For all selected endpoints, the property values take longer to load on the initial query than in subsequent iterations, probably due to browser caching.
-The running time to acquire the list of property values for the Drugbank endpoint is much larger than the time for the other two endpoints, probably due to the greater size of the Drugbank database.
-(As of 12/10/09, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, and DrugBank contains 765,936 triples).
 
@@
 The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user.
-Other features of the QueryMed system that distinguished it from similar systems included the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the semantic web data.
+Other features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the semantic web data.
 This feature increases the ease of use of our system relative to other similar systems.
@@
 Since data is stored locally after the initial startup, using the system in subsequent queries will be rapid after an initial loading phase.
-This approach might be too memory-intensive if the user wishes to run queries over many large triple stores, but might work best in the situation where there is a small or moderate amount of data in the repositories of interest to the user.
+This approach might be too memory-intensive if the user wishes to run queries over many large triple stores, but might work best in a situation where there is a small or moderate amount of data in the repositories of interest to the user.
 

### 2010-02-11 trunk/docs/www2010/paper.tex
Added all of Ian's comments

@@
 
-QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, runs queries across multiple SPARQL endpoints, and is designed to be accessible to users who do not know the structure of the underlying ontologies or the SPARQL query language.
+QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, runs queries across multiple SPARQL endpoints, and is designed to be accessible to users who do not know the structure of the underlying ontologies used in describing the datasets, or the SPARQL query language to query the data.
 The system allows users to select the data sources that they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.
@@
 After retrieval of the initial result set, query results can be filtered to improve their relevance.
-As an advanced search feature, the system also allows the user to exploit the underlying structure of the RDF data to improve query results.
+The system also allows the user to exploit the underlying structure of the RDF data to improve query results.
 \end{abstract}
@@
 
-The biomedical domain is among the early successes of the semantic web, due to the rapidity with which the biomedical community has made its data available in RDF triple stores.
-However, although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
+Publicly available data in the biomedical domain has grown significantly over the recent years.
+For example, this data include data on drug discovery~\cite{Sharp, Goble}, clinical trial information that are publicly available at ClinicalTrails.gov, disease genes, phenotypes published by the Online Mendelian Inheritance in Man (OMIM), and etc.
+With the linked open data movement, the semantic web community has been very proactive in converting these datasets to RDF~\cite{LinkingData}.
+In fact, the biomedical domain is among the early successes of adopting semantic web due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
+%The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
 
-For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
+Although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and also allow users to run federated queries on multiple SPARQL endpoints.
+The state of the art seem to be browsing the data using a linked data browser such as the Tabulator~\cite{Tabulator}.
+The linked data approach allows a user to navigate the data in an exploratory manner, but lacks support for efficient filtering and querying of the data in a manner that is amenable for domain experts.
+Therefore, there is a need to have efficient hybrid interfaces that combine both browsing and querying~\cite{Jentzsch}.
+
+For an example in the biomedical domain, a physician may know her patient's personal information, symptoms, current medications, and genotype.
 She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.
@@
 Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
-For example, to address her question, the physician might be interested in information on “coronary artery disease” (which she thinks is the cause of her patient's symptoms).
-The information about the disease can be found in Diseasome.
-She might also want to know the set of drugs in DailyMed and Drugbank that can be used to treat coronary artery disease.
-Information on clinical trials available for her patient can be found in LinkedCT.
-Due to the large number of databases  that the physician needs to query in order to find an answer to her single question, she is likely to find useful a system that can automatically run queries over multiple data sources.
+Suppose the physician is interested in information on “coronary artery disease” (which she thinks is the cause of her patient's symptoms).
+The information about the disease can be found in Diseasome, a database that lists diseases, disease genes, and associations between them~\cite{Diseasome}.
+The physician might also want to know the set of drugs in DailyMed~\cite{DailyMed}, and Drugbank~\cite{Drugbank} that can be used to treat coronary artery disease.
+Information on clinical trials that were successful in treating this particular disease of her patient can be found in LinkedCT~\cite{LinkedCT}.
+
+Due to the large number of databases  that the physician needs to query in order to find an answer to her single question, she is likely to find useful, a system that can automatically run queries over multiple data sources.
 Also, the physician may not know SPARQL query syntax, the location of the SPARQL endpoints, or the structure of the relevant ontologies.
@@
 The semantic web can be viewed as a global database system for the data available on the world wide web.
-Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using the SPARQL query language \cite{Berners-Lee1}.
-The addition of structure to web data allows inferences to be automatically drawn by intelligent agents integrating data from multiple sources.
-Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
+Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using the SPARQL query language.
+The addition of structure to web data allows inferences to be automatically drawn by intelligent agents integrating data from multiple sources~\cite{Berners-Lee1}.
+%Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
+
 Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF.
-Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists, and these semantic web resources represent valuable standardized repositories of information that can be automatically mined for biological knowledge.
-%The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
+Almost all of these data sources are interlinked with each other.
+Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists.
+These semantic web resources represent valuable repositories of information that can be automatically mined for applications that require biological knowledge.
 
-However, although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources accessible to physicians, patients, and life scientists.
+Although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources accessible to physicians, patients, and life scientists.
 One challenge is constructing systems that allow end users to run intuitive queries on biomedical data.
 Users of biomedical resources are likely to have extensive domain knowledge, but be unfamiliar with the SPARQL query language syntax and with the structure of biomedical ontologies.
-It is important to design user-friendly systems that allow these users to take advantage of the wealth of structured biological knowledge available on the semantic web.
-Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data is often distributed among many sources \cite{Pasquier}.
+Almost all the SPARQL endpoints available today are offering a generic query interface that requires users to write SPARQL queries.
+Also, no hints about the structure of data available from the SPARQL endpoint is given as the user is constructing the SPARQL query.
+The user has to essentially construct the queries that are needed for her purpose by trial and error.
+This can be a daunting task  especially for someone who is new to semantic web technologies.
+Therefore, it is important to design user-friendly systems that allow users to take advantage of the wealth of structured biological knowledge available on the semantic web.
+Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data are often distributed among many sources \cite{Pasquier}.
 
-\section{QueryMed}
+\section{Design \& Implementation}
 \label{system}
 
-\subsection{Architecture overview}
+QueryMed allows users to query many biomedical data sources from a single interface.
+Although there is a default list of sources that the user can choose to query from, this list is  not fixed, and can be extended very easily.
+The system was designed with non-expert users of SPARQL in mind, and presents a very intuitive user interface that anybody with basic biomedical domain knowledge can use.
 
-A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
-The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.
+The system functionality in a nutshell is as follows: A user submits a query from the Query Interface to the SPARQL endpoints via the proxy server.
+This is translated into a SPARQL query for each individual endpoint.
+Once the results are obtained from the SPARQL endpoints, they are presented to the user via the Result Browser.
+The user can then choose to refine the query by changing some of the original query terms in the Dynamic Facet Creator/Filter.
+This is an iterative process, where the user can refine the query by adding new sources, and tweaking values given to the properties.
 
-QueryMed is implemented in Java in the backend and JavaScript in the frontend.
-The JQuery library \cite{Resig} was used to develop an attractive user interface.
-In the backend, the Jena library \cite{McBride} is used to run the SPARQL queries.
-QueryMed relies on the following remote resources:
-\begin{itemize}
-\item \textbf{SPARQL Endpoints}:  QueryMed runs queries on multiple remote SPARQL endpoints that expose biomedical data.
-\item \textbf{Sources List}: The default list of endpoints available to the user is stored in a source list on the proxy server.
-\end{itemize}
+In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind some of the crucial components of the system.
+
+\subsection{QueryMed Overview}
 
@@
 
-\subsubsection{Proxy Server}
+A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
+The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.
+A more detailed illustration of the QueryMed system highlighting all the components necessary for system functionality is given in Figure \ref{fig:arch_details}.
 
-The user submits a query to the SPARQL endpoints via the proxy server.
-The query is translated into a SPARQL query for each individual endpoint, and the results are returned to the user.
-The specific components implemented in the proxy server are as follows.
-\begin{itemize}
-\item \textbf{Source Manager:} The source manager reads the source list and populates the default query list on the user interface.
-It also keeps track of the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
-\item \textbf{Translator:} This component is responsible for translating the user query into valid SPARQL syntax.
-\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries.
-It takes input on source and keyword from the client, and returns  the query results as a JSON object.
-The structure of the JSON object is given in figure \ref{fig:JSON_runquery}.
-\item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
-\item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
-\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
-The structure of this JSON object is given in figure \ref{fig:JSON_advquery}.
-After constructing and executing the SPARQL query, it returns a set of query results as a JSON object in the format given in figure \ref{fig:JSON_runquery}.
-\end{itemize}
+%The system relies on two external resources:
+
+%\begin{enumerate}
+%\item \textbf{SPARQL endpoints} that expose biomedical data.
+%\item \textbf{Sources List} that gives a list of endpoints available to the user. This list is stored on the proxy server, but can be replaced by the user. 
+%\end{enumerate}
 
@@
 
-The query interface is implemented in JavaScript, HTML and CSS. The main components of the user interface are as follows: 
+The user interface is one of the primary components of the QueryMed system. It comprises of the following:
+
 \begin{itemize}
-\item \textbf{Initial Query Interface:} The user inputs their query  term into the provided text box (figure \ref{fig:initial_ui}).
-\item \textbf{Add New Source:} Allows the user to add a new data source to extend the query (figure \ref{fig:add_source})
-\item \textbf{Dynamic Facet Creator/Filter:} This component allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (figure  \ref{fig:facet_creator}).
-\item \textbf{Result Browser:} The user can view their query results, organized by source.
-The results are presented in a table with pagination.
-The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns (figure \ref{fig:results}).
-\end{itemize}
 
-\subsection{System Functionality and Design Decisions}
-
-\subsubsection{Source list}
-
-Because the set of default endpoints is stored on the proxy server, the set of resources available by default to the user can easily be updated.
-Since useful biomedical resources are rapidly being developed and made available as RDF triple stores, the ease of updating the resource list ensures that the system can easily be brought up to date.
-In fact, the QueryMed system could easily be adapted to perform queries outside the biomedical domain by modifying the list of input data sources.
-
-\subsubsection{Proxy Server}
-
-While an entirely client side application is possible, we chose to have a proxy server perform the SPARQL query execution and caching. This design is advantageous for several reasons:
-1.
-Efficient cache management: The result set from running an unrestricted query can include millions of results.
-It may be infeasible to keep unfiltered query results in browser memory.
-A typical memory footprint for a browser (for e.g.
-Firefox) is usually between 20MB and 100MB.
-A poorly constructed query can result in gigabytes worth of triples returned, causing the browser to crash.
-The proxy server can cache results from initial query execution, so that only a filtered result set is subsequently returned to the client.
-2. The proxy server enables us to avoid cross domain XML-HTTP-Request errors in accessing SPARQL endpoints in various domains.
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-SELECT ?projection_1 ?projection_2 ... 
-	?projection_n
-WHERE {?x source:property ? projection 
-	FILTER regex(? projection, '" + 
-	input +"', 'i)
-}
-... //Other property filers are to follow
-\end{Verbatim}
-\caption{Minimal SPARQL Query Structure in the ``Query All" Case}
-\label{fig:min_query}
-\end{figure}
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-{
-     "bindings":
-     [
-     	{"source" : "source_label",
-     	 "uri" : "http://source.com/sparql.",
-     	 "vars" : ["variable1","variable2"],
-     	 "count":2,
-     	 "results" : 
-     		[
-     		 {"variable1": "Value 1",
-		  "variable1": "Value 2"},
-     		 {"variable1": "Another 
-		 Value 1",
-		  "variable1": "Another Value 2"},
-
-     		]
-     	}
-     	,
-	{ //Another Source and Result Description
-	...
-	},
-	...
-     ]
- }
-\end{Verbatim}
-\caption{JSON returned from the RunQuery service}
-\label{fig:JSON_runquery}
-\end{figure}
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-{
-   "http://some-source/sparql": {
-   	"property1": {"value1": "AND"},
-	"property2": {"value 2": "OR"},
-	"property2": {"value 3": "FILTER"}
-	           }
-	}
-   //Other sources ...
-}
-\end{Verbatim}
-\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
-\label{fig:JSON_advquery}
-\end{figure}
-
-\subsubsection{Data Structures}
-
-The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
-In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
-The generated SPARQL query will be of the form illustrated in figure \ref{fig:min_query}.
-The word ``input" in the figure represents the keyword specified by the user.
-
-When the results are returned, each result set is structured as the JSON object given in figure \ref{fig:JSON_runquery}.
-This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
-
-When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All" case.
-The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
-This information is passed as a JSON object, whose structure is given in figure \ref{fig:JSON_advquery}, to the proxy server.
-
-\subsubsection{Query Interface}
-
-In order to make QueryMed easy to use, but still provide a flexible tool that is able to perform a range of queries, we provide an uncluttered interface with only a  blank search box and two buttons.
+\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but still provide a flexible tool that is able to perform a range of queries, we provide an uncluttered interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
 The ``Query All" button will perform a keyword-based query over pre-selected data sources.
-The advanced search option ``Refine Query" allows the user to view additional query options.
-The initial query interface is shown in figure \ref{fig:initial_ui}.
+The advanced search option ``Refine Query" allows the user to view additional query options that are handled in the Advanced Query Interface.
 
@@
 \epsfig{file=images/new_front.jpg,  width=3in}
-\caption{Initial query interface seen by the user}
+\caption{Initial Basic Query Interface}
 \label{fig:initial_ui}
@@
 
-We believe that the unstructured keyword search is likely to encourage users without previous experience running queries on semantic web data to explore the additional features of our system that allow more flexible and advanced queries.
-For example, physicians who are not used to running queries on semantic web data may be more willing to invest time into learning how to find better search results once they see that the system can easily provide them with some initial useful query results using the basic keyword search option.
-However, the execution of queries using QueryMed will be very much different from a user issuing a web search on a search engine such as Google, because it exploits the underlying structure of the data as exposed in the individual data endpoints.
+\item \textbf{Advanced Query Interface:} This allows the user to add new data sources, and give values for properties found at a particular selected data source.
+Therefore, the user is able to extend the query from a basic keyword search over all the properties as with the basic query interface, to a more targeted query.
 
-The advanced query interface provides the user with the option of selecting specific data sources to query, as shown in figure \ref{fig:select}.
-The user can also dynamically add additional data sources, as shown in figure \ref{fig:add_source}.
+When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources in the system as shown in Figure \ref{fig:select}.
 This feature allows the user to perform their query only over relevant or trusted data sources.
-The ability to dynamically add additional sources increases the flexibility of  the QueryMed system, allowing users to search endpoints of interest that are not included in the default list.
+If none of the desired data sources are available in this list, the user can dynamically add additional data sources as shown in Figure \ref{fig:add_source}.
+The ability to dynamically add additional data sources increases the flexibility of  the QueryMed system, allowing users to query endpoints of interest that are not included in the default list.
 
@@
 \centering
-\epsfig{file=images/select.jpg,  width=3in}
+\epsfig{file=images/select.jpg,  width=2.5in}
 \caption{The user has the option of selecting specific trusted or relevant data sources, or of adding additional data sources to query.}
@@
 \epsfig{file=images/add_source.jpg,  width=3in}
-\caption{The user can also dynamically add new data sources}
+\caption{Add New Data Source}
 \label{fig:add_source}
@@
 
-\begin{figure}
-\begin{Verbatim}[frame=single]
-SELECT DISTINCT ?property
-WHERE { [] ?property [] }
-ORDER BY ?property
-\end{Verbatim}
-\caption{SPARQL query to retrieve all properties of a given datasource}
-\label{fig:SPARQL_props}
-\end{figure}
+When a data source is selected, the properties list is automatically populated with all the distinct properties available at the selected endpoint.
+If the user wants to restrict her query based on one or more properties available, she can specify those.
+This allows users to improve their queries using the underlying structure of the data without having prior knowledge of this structure to obtain more relevant query results.
+%The properties list is generated by running a SPARQL query of the form given in Figure \ref{fig:SPARQL_props} at the specified data source.
+
+Once the properties are returned, they will be displayed in the user interface as shown in Figure \ref{fig:add_props}.
+Clicking an individual property link displays a description of the property.
+This allows the user to understand the keywords that will be most useful to search on each specific property.
+
+The user can choose to perform exact queries, or filter results on specific keywords.
+If the user does not know the specific value for a property, they can instead specify a keywords using the “FILTER” option.
+They also can choose logical operators to connect the various parts of their query.
+When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
+as a conjunction.
+When “OR” is used, the specified graph pattern is made to disjunct with the rest of the query with the SPARQL UNION operator.
+If no logical operator is specified, “AND” will be used by default.
+The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in Figure \ref{fig:SPARQL_adv}.
 
@@
 \end{figure*}
-
-When a data source is selected, the properties list is automatically populated with the properties available at the selected endpoint so that the user can run queries specific to the underlying structure of the RDF data in order to find more relevant query results.
-The properties list is generated by running a SPARQL query of the form given in figure \ref{fig:SPARQL_props} at the specified data source.
-This allows users to improve their queries using the underlying structure of the data without prior knowledge of this structure.
-
-Once the properties are returned, those will be displayed in the user interface as shown in figure \ref{fig:add_props}.
-Clicking an individual property link displays a description of the property, allowing the user to understand the keywords that will be most useful to search on each specific property.
-%A sample description for the property rdfs:label is shown in figure \ref{fig:predicate_desc}.
-
-%\begin{figure}
-%\centering
-%\epsfig{file=images/predicate_desc.jpg, width=3.5in}
-%\caption{Clicking on a property opens a window containing a description of that property}
-%\label{fig:predicate_desc}
-%\end{figure}
-
-The user can choose to perform exact queries, or filter results on specific keywords.
-If the user does not know the specific value for a property, they can instead specify a keywords using the “FILTER” option.
-They also can choose logical operators to connect the various parts of their query.
-When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
-as a conjunction.
-When “OR” is used the specified graph pattern is made to disjunct with the rest of the query with the SPARQL UNION operator.
-If no logical operator is specified, “AND” will be used by default.
-The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in figure \ref{fig:SPARQL_adv}.
 
@@
 
-The Dynamic Facet Creator shown in figure \ref{fig:facet_creator} allows the user to view properties from multiple endpoints, and to expand and contract properties for individual endpoints.
-This feature enables the interface to display many property lists simultaneously, grouped by source, so that only information relevant to the endpoint the user is currently examining will be visible at any time.
+\item \textbf{Dynamic Facet Creator/Filter:} This allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint as shown in Figure  \ref{fig:facet_creator}.
+The interface allows displaying of many property lists from different sources simultaneously (grouped by source).
+Therefore, only information relevant to the endpoint the user is currently examining will be visible at any given time.
 
@@
 
-Query results are displayed to the user as a table, which can be filtered and searched to refine the query result, or printed as shown in in figure \ref{fig:results}.
-This feature allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+\item \textbf{Result Browser:} The user can view their query results organized by source in the Result Browser.
+The results are presented in a table with pagination.
+The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
+This allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
 
@@
 
-\subsection{Performance}
+\end{itemize}
 
-The slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
-We compared the times required to load properties from several endpoints, by running the query illustrated in figure \ref{fig:SPARQL_props}.
-Timing data are shown in table \ref{timing}.
+\subsubsection{Proxy Server}
 
-\begin{table}
-\begin{tabular}{|l||l|l|l|}
-\hline
- & \textbf{Dieseasome} & \textbf{Dailymed} & \textbf{Drugbank} \\
-\hline
-\hline
-\textbf{1st Trial}   & 3.45 s & 1.77 s  & 9.51 s  \\
-\textbf{2nd Trial}  & 1.61 s & 1.57 s & 9.34 s \\
-\textbf{3rd Trial}  & 1.71 s & 1.66 s & 9.06 s \\
-\hline
-\end{tabular}
-\caption{Running times to retrieve all the properties from selected endpoints}
-\label{timing}
-\end{table}
+The functionality of the proxy server is to act as an intermediary between the user interface and the remote SPARQL endpoints.  Its functionality is twofold:
 
-For all selected endpoints, the property values take longer to load on the initial query than in subsequent iterations, probably due to browser caching.
-The running time to acquire the list of property values for the Drugbank endpoint is much larger than the time for the other two endpoints, probably due to the greater size of the Drugbank database.
-(As of 12/10/09, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, and DrugBank contains 765,936 triples).
+\begin{enumerate}
+\item Execute the SPARQL queries at the relevant remote SPARQL endpoints and consolidate the results to be presented in the user interface.
+\item Cache the results for subsequent queries for refinements to the same query to reduce the network and query execution latency.
+\end{enumerate}
+
+The specific components in the Proxy Server are as follows.
+
+\begin{itemize}
+
+\item \textbf{Source Manager:} The Source Manager reads the source list and populates the default query list on the user interface.
+It also keeps track of the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
+
+\item \textbf{Translator:} The Translator is responsible for translating the user query into valid SPARQL syntax.
+Translator obtains the parameters to construct the query from the input the user specifies in the Query Interface, and plugs them in to a relevant template.
+It specifically relies on two services:
+
+\begin{enumerate}
+
+\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries generated by the Translator, and return  the query results as a JSON object given in Figure \ref{fig:JSON_runquery}.
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
+{
+   "bindings":
+     [
+        {"source" : "source_label",
+         "uri" : "http://source.com/sparql.",
+     	"vars" : ["variable1","variable2"],
+     	"count":2,
+     	"results" : 
+     	 [
+     	    {"variable1": "Value 1",
+	      "variable1": "Value 2"},
+     	    {"variable1": "Another Value 1",
+               "variable1": "Another Value 2"},
+   	    ...	  		
+     	 ]
+     	}
+     	,
+	{ //Another Source and Result Description
+	...
+	},
+	...
+     ]
+ }
+\end{Verbatim}
+\caption{Structure of the JSON ``result" object}
+\label{fig:JSON_runquery}
+\end{figure}
+
+\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object given in Figure \ref{fig:JSON_advquery}.
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
+{
+ "http://some-source/sparql": 
+  {
+     "property1": {"value1":  "logical operator"},
+     "property2": {"value 2": "logical operator"},
+     "property2": {"value 3": "logical operator"},
+     ...
+  }
+}
+   //Other sources ...
+
+\end{Verbatim}
+\caption{Structure of the JSON ``query" object}
+%\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
+\label{fig:JSON_advquery}
+\end{figure}
+
+\end{enumerate}
+
+% The user has to specify the data source(s) to query, and can give specific values for one or many of the properties available for the data available to restrict the query space. 
+\item \textbf{Additional Services:} In order to make the user interface more user friendly, the following two services perform some ``behind the scenes" work.
+
+\begin{enumerate}
+\item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
+\item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
+\end{enumerate}
+
+\end{itemize}
+
+\subsection{Design Decisions}
+
+\subsubsection{Source list}
+
+Because the set of default endpoints is stored on the proxy server, the set of resources available by default to the user can easily be updated.
+Since useful biomedical resources are rapidly being developed and made available as RDF triple stores, the ease of updating the resource list ensures that the system can easily be brought up to date.
+In fact, the QueryMed system could easily be adapted to perform queries outside the biomedical domain by modifying the list of input data sources.
+
+\subsubsection{Proxy Server}
+
+While an entirely client side application is possible, we chose to have a proxy server perform the SPARQL query execution and caching. This design is advantageous for several reasons:
+
+\begin{enumerate}
+
+\item Efficient cache management: The result set from running an unrestricted query can include millions of results.
+It may be infeasible to keep unfiltered query results in browser memory.
+A typical memory footprint for a browser (for e.g.
+Firefox) is usually between 20MB and 100MB.
+A poorly constructed query can result in gigabytes worth of triples returned, causing the browser to crash.
+The proxy server can cache results from initial query execution, so that only a filtered result set is subsequently returned to the client.
+
+\item The proxy server enables us to avoid cross domain XML-HTTP-Request errors in accessing SPARQL endpoints in various domains.
+
+\end{enumerate}
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
+SELECT ?projection_1 ?projection_2 ... 
+	?projection_n
+WHERE {?x source:property ? projection 
+	FILTER regex(? projection, '" + 
+	input +"', 'i)
+}
+... //Other property filers are to follow
+\end{Verbatim}
+\caption{Minimal SPARQL Query Structure in the ``Query All" Case}
+\label{fig:min_query}
+\end{figure}
+
+\subsubsection{Data Structures}
+
+The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
+In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
+The generated SPARQL query will be of the form illustrated in Figure \ref{fig:min_query}.
+The word ``input" in the figure represents the keyword specified by the user.
+
+When the results are returned, each result set is structured as the JSON object given in Figure \ref{fig:JSON_runquery}.
+This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
+
+When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is bit more complex than in the ``Query All" case.
+The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
+This information is passed as a JSON object, whose structure is given in Figure \ref{fig:JSON_advquery}, to the proxy server.
+
+\subsubsection{Query Interface}
+
+We believe that the unstructured keyword search is likely to encourage users without previous experience running queries on SPARQL endpoints to explore the additional features of our system that allow more flexible and advanced queries.
+For example, physicians may be more willing to invest time into learning how to find better search results once they see that the system can easily provide them with some initial useful query results using the basic keyword search option.
+However, the execution of queries using QueryMed will be very different from a user issuing a web search on a search engine such as Google because it exploits the underlying structure of the data exposed by the individual data endpoints.
+
+\subsubsection{Implementation}
+
+QueryMed is implemented in Java in the backend and JavaScript, HTML and CSS in the frontend.
+In the backend, the Jena library \cite{McBride} is used to run the SPARQL queries and 4-store~\cite{4store} is used as a triple store to provide caching support.
+The JQuery library \cite{Resig} was used to develop an attractive user interface.
 
@@
 
+\subsection{Performance}
+
+The slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
+We compared the times required to load properties from several endpoints, by running the query illustrated in Figure \ref{fig:SPARQL_props}.
+Timing data are shown in Table \ref{timing}.
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
+SELECT DISTINCT ?property
+WHERE { [] ?property [] }
+ORDER BY ?property
+\end{Verbatim}
+\caption{SPARQL query to retrieve all properties of a given data source}
+\label{fig:SPARQL_props}
+\end{figure}
+
+For all selected endpoints, the property values take longer to load on the initial query than in subsequent iterations, probably due to browser caching.
+The running time to acquire the list of property values for the Drugbank endpoint is much larger than the time for the other two endpoints, probably due to the greater size of the Drugbank database.
+(As of 12/10/09, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, and DrugBank contains 765,936 triples).
+
+\begin{table}
+\begin{tabular}{|l||l|l|l|}
+\hline
+ & \textbf{Dieseasome} & \textbf{Dailymed} & \textbf{Drugbank} \\
+\hline
+\hline
+\textbf{1st Trial}   & 3.45 s & 1.77 s  & 9.51 s  \\
+\textbf{2nd Trial}  & 1.61 s & 1.57 s & 9.34 s \\
+\textbf{3rd Trial}  & 1.71 s & 1.66 s & 9.06 s \\
+\hline
+\end{tabular}
+\caption{Running times to retrieve all the properties from selected endpoints}
+\label{timing}
+\end{table}
+
 \subsection{QueryMed Resources}
@@
 \begin{center}
-\begin{tabular}{| p{2cm} | p{2cm}| p{2cm}| p{2cm}| p{2cm}| p{2cm} |}
+\begin{tabular}{| p{1.8cm} | p{2cm} | p{1.8cm} | p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm} |}
 \hline
- & \textbf{Query Multiple Sources?} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
+ & \textbf{Hybrid Interface (Combines Querying \& Browsing)} & \textbf{Provides Local Caching} & \textbf{Queries Multiple Sources} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
 \hline
 \hline
-\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes  \\
+\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes & Yes & Yes  \\
 \hline
-\textbf{SMART}  & Yes & No & No & Yes & Yes  \\
+\textbf{SMART}  & & & Yes & No & No & Yes & Yes  \\
 \hline
-\textbf{DARQ}  & Yes & N/A & No & Yes & No  \\
+\textbf{DARQ}  & &   & Yes & N/A & No & Yes & No  \\
 \hline
-\textbf{GoWeb}  & Yes & No & Yes & No & Yes  \\
+\textbf{GoWeb}   & &  & Yes & No & Yes & No & Yes  \\
 \hline
-\textbf{BioGateway}  & Yes & No & No & No & Yes  \\
+\textbf{BioGateway}   & &  & Yes & No & No & No & Yes  \\
 \hline
-\textbf{Twinkle}  & Yes & No & No & Yes & Yes  \\
+\textbf{Twinkle}   & &  & Yes & No & No & Yes & Yes  \\
 \hline
@@
 Table \ref{comparison} compares selected features of the QueryMed system with other related systems.
-The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user.
-Other features of the QueryMed system that distinguished it from similar systems included the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the semantic web data.
+The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user, and provides a hybrid interface that enables the users to both query and browse data.
+Other features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data.
 This feature increases the ease of use of our system relative to other similar systems.
@@
 
-\section{Future Work and Conclusions}
+\section{Future Work}
 \label{conclusion}
@@
 
+\section{Conclusion}
+
 The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
 Because of the unique features of our system, we believe it will be of use to the biomedical community.
+We also believe tools such as QueryMed will entice more people to expose their data as linked open data.
 

### 2010-02-11 trunk/docs/www2010/paper.tex
merging use case with architecture description

@@
 In fact, the biomedical domain is among the early successes of adopting semantic web due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
+
 %The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
 
-Although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and also allow users to run federated queries on multiple SPARQL endpoints.
-The state of the art seem to be browsing the data using a linked data browser such as the Tabulator~\cite{Tabulator}.
-The linked data approach allows a user to navigate the data in an exploratory manner, but lacks support for efficient filtering and querying of the data in a manner that is amenable for domain experts.
-Therefore, there is a need to have efficient hybrid interfaces that combine both browsing and querying~\cite{Jentzsch}.
+Although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
+Many currently available systems are linked data data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner, but lack support for efficiently filtering and querying the data.
+Therefore, there is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}.
 
@@
 The physician might also want to know the set of drugs in DailyMed~\cite{DailyMed}, and Drugbank~\cite{Drugbank} that can be used to treat coronary artery disease.
-Information on clinical trials that were successful in treating this particular disease of her patient can be found in LinkedCT~\cite{LinkedCT}.
+Information on clinical trial for which her patient may be eligible can be found in LinkedCT~\cite{LinkedCT}.
 
@@
 
+ For example, a physician interested in finding semantic web resources related to coronary artery disease could use “coronary artery disease”  as a search keyword in the input search box and select the "Query All" option.
+The simplicity of the basic query interface allows this end user to rapidly and easily browse semantic web resources of interest.
+
 \begin{figure}
@@
 \end{figure}
+
+\item \textbf{Result Browser:} The user can view their query results organized by source in the Result Browser.
+The results are presented in a table with pagination.
+The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
+This allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+
+For example, after searching for "coronary artery disease" in the basic query interface, the physician will see displayed in the Result Browser a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery
+disease. She can then filter the results using additional search terms. For example, she knows that 
+the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view the most relevant search terms.
 
@@
 
+For example, the physician might be interested in finding relevant clinical trials for 
+her patient. However, the clinical trial database LinkedCt is 
+not in the default set of endpoints, so she selects the “Refine 
+Query” option to choose additional endpoints to search. She 
+sees a list of default endpoints, and selects the “Add” option 
+to include an additional endpoint. After entering the name 
+and URL of the LinkedCT endpoint, she is able to search 
+for clinical trials for which her patient may be eligible.
+
 \item \textbf{Dynamic Facet Creator/Filter:} This allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint as shown in Figure  \ref{fig:facet_creator}.
@@
 Therefore, only information relevant to the endpoint the user is currently examining will be visible at any given time.
+
+For example, the physician might want to further refine her search for cardiovascular diseases. She 
+uses the advanced search option to construct a boolean query 
+that takes advantage of the underlying structure of the RDF 
+data in the database. She searches Diseasome for a list of 
+diseases whose class is “Cardiovascular” or for which the as- 
+sociated gene is “ABCA1.” Using the QueryMed advanced 
+search interface, the complex SPARQL query corresponding 
+to her question is automatically constructed (Figure \ref{fig:SPARQL_adv}), and she can 
+view the query results conveniently displayed in the Results Browser.
 
@@
 \end{figure}
-
-\item \textbf{Result Browser:} The user can view their query results organized by source in the Result Browser.
-The results are presented in a table with pagination.
-The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
-This allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
 
@@
 
-\subsection{A Sample Use Case}
+%\subsection{A Sample Use Case}
 
-A physician is interested in finding semantic web resources related to coronary artery disease.
+%A physician is interested in finding semantic web resources related to coronary artery disease.
 She first tries a basic search over all default resources, by entering ``coronary artery disease" in the input search box.
@@
 
-She is now interested in finding relevant clinical trials for her patient.
+%She is now interested in finding relevant clinical trials for her patient.
 However, the clinical trial database LinkedCt is not in the default set of endpoints, so she selects the ``Refine Query" option to choose additional endpoints to search.
@@
 
-She is also interested in further refining her search.
+%She is also interested in further refining her search.
 She uses the advanced search option to construct a boolean query that takes advantage of the underlying structure of the RDF data in the database.
@@
 
-A video illustrating this sample use case can be found at: \\
+%A video illustrating this sample use case can be found at: \\
 http://www.youtube.com/watch?v=JCr7vryS9Vg
@@
 http://code.google.com/p/querymed/
+
+A video illustrating a sample use case can be found at: [USE CASE URL]
 

### 2010-02-12 docs/www2010/paper.tex
revised intro, backround, and some of design/implementation sections

@@
 
-The biomedical domain is among the early successes of the semantic web, due to the rapidity with which the biomedical community has made its data available in RDF triple stores \cite{Yip}.
-Although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and also allow users to run federated queries on multiple SPARQL endpoints.
+The quantity of publicly available data in the biomedical domain has dramatically increased over recent years.
+Publicly available biomedical resources include data on drug discovery~\cite{Sharp, Goble}, clinical trials, diseases, disease genes, and phenotypes.
+With the linked open data movement, the semantic web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
+In fact, the biomedical domain is among the early successes of the semantic web due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
 
+%The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
+
+To allow end users to exploit the abundance of useful biomedical data that is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
+There is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}, since many currently available systems are linked data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner but lack support for filtering and querying the data.
+
+Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple endpoints.
 For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
@@
 
-Suppose the physician might be interested in information on “coronary artery disease” (which she thinks is the cause of her patient's symptoms).
-The information about the disease can be found in Diseasome, a database that lists diseases, disease genes, and associations between them~\cite{Diseasome}.
-The physician might also want to know the set of drugs in DailyMed~\cite{DailyMed} and Drugbank~\cite{Drugbank} that can be used to treat coronary artery disease.
-Information on clinical trials available for her patient can be found in LinkedCT~\cite{LinkedCT}.
-Due to the large number of databases  that the physician needs to query in order to find an answer to her single question, she is likely to find useful a system that can automatically run queries over multiple data sources.
+Since the physician must search many databases in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.
 Also, the physician may not know SPARQL query syntax, the location of the SPARQL endpoints, or the structure of the relevant ontologies.
@@
 Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.
-
-\cite{Jentzsch}
 
@@
 
-The semantic web can be viewed as a global database system for the data available on the world wide web.
+The semantic web can be viewed as a global database system for the information available on the world wide web.
 Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using the SPARQL query language.
@@
 
-Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF.
+Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
+Almost all of these data sources are interlinked with each other.
 Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists.
 These semantic web resources represent valuable repositories of information that can be automatically mined for applications that require biological knowledge.
-%The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
 
@@
 The user has to essentially construct the queries that are needed for her purpose by trial and error.
-This can especially be a daunting task for someone who is new to semantic web technologies.
+This can be a daunting task  especially for someone who is new to semantic web technologies.
 Therefore, it is important to design user-friendly systems that allow users to take advantage of the wealth of structured biological knowledge available on the semantic web.
-Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data is often distributed among many sources \cite{Pasquier}.
+Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data are often distributed among many sources \cite{Pasquier}.
 
-\section{System Design \& Implementation}
+\section{Design \& Implementation}
 \label{system}
+
+The QueryMed system allows users to easily query multiple biomedical data sources.
+Queries can be run against a default list of datasources, or against a set of user-selected SPARQL endpoint.
+The end user can easily input additional endpoints in order to utilize resources that are not included in the default list.
+The system automatically translates the user input into a SPARQL query for each individual endpoint, combines the results, and returns them to the user.
+The user can choose to refine the query by iteratively modifying the original query terms, and by filtering the result list.
+The advanced query functionality of the QueryMed system allows the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the ontologies.
+The simple user interface of the QueryMed system is designed to be intuitive for users with no knowledge of the SPARQL query language.
+
+In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind some of the crucial components of the system.
+
+\subsection{QueryMed Overview}
 
@@
 
-QueryMed allows users to query many biomedical data sources from a single interface.
-The data sources that the user wishes to query are not fixed, and can be extended very easily.
-The system was designed with non-expert users of SPARQL in mind, and presents a very intuitive user interface that anybody with basic biomedical domain knowledge can use.
-
-The system functionality in a nutshell is as follows: A user submits a query from the Query Interface to the SPARQL endpoints via the proxy server.
-The query is translated into a SPARQL query for each individual endpoint.
-Once the results are obtained from the SPARQL endpoints, they are presented to the user via the Result Browser.
-The user can then choose to refine the query by changing some of the original query terms in the Dynamic Facet Creator/Filter, and this process repeats.
-
-In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind some of the crucial components of the system.
-
-\subsection{QueryMed Overview}
-
-A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
-The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.
-
-A more detailed illustration of the QueryMed system highlighting all the components necessary for system functionality is given in Figure \ref{fig:arch_details}.
-The system relies on two external resources:
-
-\begin{enumerate}
-\item \textbf{SPARQL endpoints} that expose biomedical data.
-\item \textbf{Sources List} that gives a list of endpoints available to the user. This list is stored on the proxy server, but can be replaced by the user. 
-\end{enumerate}
-
 \begin{figure}
@@
 
+A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
+The main components of the system are the user interface and the proxy server that takes input from the user and retrieves relevant biomedical data from remote SPARQL endpoints.
+After a user submits a query using the user interface, the query is translated by the proxy server into individual SPARQL queries for each remote endpoint.
+The query results are returned from the remote endpoints, combined by the proxy server, and presented to the user.
+A detailed illustration of the parts of the QueryMed system is shown in Figure \ref{fig:arch_details}.
+
+%The system relies on two external resources:
+
+%\begin{enumerate}
+%\item \textbf{SPARQL endpoints} that expose biomedical data.
+%\item \textbf{Sources List} that gives a list of endpoints available to the user. This list is stored on the proxy server, but can be replaced by the user. 
+%\end{enumerate}
+
 \subsubsection{User Interface}
 
-The user interface is one of the primary components of the QueryMed system. It comprises of the following:
+The QueryMed user interface is designed to be intuitive for the end user, yet flexible enough to permit a broad range of interesting queries.
+The basic query interface allows the user to run simple queries, and is designed for maximal easy of use.
+The advanced search capabilities allow the user to easily construct complex logical queries that take advantage of the underlying structure of the biomedical ontologies.
+The user interface also allows the user to iteratively refine queries, and displays the query results, which have been retrieved from multiple SPARQL endpoints.
+The main components of the user interface are as follows:
 
@@
 \epsfig{file=images/new_front.jpg,  width=3in}
-\caption{Initial query interface seen by the user}
+\caption{Initial Basic Query Interface}
 \label{fig:initial_ui}
 \end{figure}
+
+\item \textbf{Result Browser:} The user can view their query results organized by source in the Result Browser.
+The results are presented in a table with pagination.
+The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
+This allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+
+For example, after searching for "coronary artery disease" in the basic query interface, the physician will see displayed in the Result Browser a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery
+disease. She can then filter the results using additional search terms. For example, she knows that 
+the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view only the most relevant query results.
 
@@
 If none of the desired data sources are available in this list, the user can dynamically add additional data sources as shown in Figure \ref{fig:add_source}.
-The ability to dynamically add additional sources increases the flexibility of  the QueryMed system, allowing users to query endpoints of interest that are not included in the default list.
+The ability to dynamically add additional data sources increases the flexibility of  the QueryMed system, allowing users to query endpoints of interest that are not included in the default list.
 
@@
 
-%\begin{figure}
-%\begin{Verbatim}[frame=single]
-%SELECT DISTINCT ?property
-%WHERE { [] ?property [] }
-%ORDER BY ?property
-%\end{Verbatim}
-%\caption{SPARQL query to retrieve all properties of a given datasource}
-%\label{fig:SPARQL_props}
-%\end{figure}
-
 Once the properties are returned, they will be displayed in the user interface as shown in Figure \ref{fig:add_props}.
@@
 This allows the user to understand the keywords that will be most useful to search on each specific property.
+
+The user can choose to perform exact queries, or filter results on specific keywords.
+If the user does not know the specific value for a property, they can instead specify a keywords using the “FILTER” option.
+They also can choose logical operators to connect the various parts of their query.
+When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
+as a conjunction.
+When “OR” is used, the specified graph pattern is made to disjunct with the rest of the query with the SPARQL UNION operator.
+If no logical operator is specified, “AND” will be used by default.
+The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in Figure \ref{fig:SPARQL_adv}.
 
@@
 
+\begin{figure}
+\begin{Verbatim}[frame=single]
+SELECT distinct ?disease WHERE { 
+{?x <http://www.w3.org/2000/01/rdf-schema#label> 
+?disease 
+FILTER regex(?disease, 
+"coronary artery disease", "i"). 
+?x <http://www4.wiwiss.fu-berlin.de/diseasome/
+resource/diseasome/class> 
+<http://www4.wiwiss.fu-berlin.de/diseasome/
+resource/diseaseClass/Cardiovascular>} 
+UNION { ?x <http://www4.wiwiss.fu-berlin.de/
+diseasome/resource/diseasome/associatedGene>
+ <http://www4.wiwiss.fu-berlin.de/diseasome/
+ resource/genes/ABCA1>.}
+}
+\end{Verbatim}
+\caption{A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.}
+\label{fig:SPARQL_adv}
+\end{figure}
+
+For example, the physician might be interested in finding relevant clinical trials for 
+her patient. However, the clinical trial database LinkedCt is 
+not in the default set of endpoints, so she selects the “Refine 
+Query” option to choose additional endpoints to search. She 
+sees a list of default endpoints, and selects the “Add” option 
+to include an additional endpoint. After entering the name 
+and URL of the LinkedCT endpoint, she is able to search 
+for clinical trials for which her patient may be eligible.
+
 \item \textbf{Dynamic Facet Creator/Filter:} This allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint as shown in Figure  \ref{fig:facet_creator}.
@@
 \end{figure}
-
-\item \textbf{Result Browser:} The user can view their query results organized by source in the Result Browser.
-The results are presented in a table with pagination.
-The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
-This feature allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
 
@@
      "property2": {"value 2": "logical operator"},
-     "property2": {"value 3": "logical operator"}
+     "property2": {"value 3": "logical operator"},
+     ...
   }
@@
 While an entirely client side application is possible, we chose to have a proxy server perform the SPARQL query execution and caching. This design is advantageous for several reasons:
-1.
-Efficient cache management: The result set from running an unrestricted query can include millions of results.
+
+\begin{enumerate}
+
+\item Efficient cache management: The result set from running an unrestricted query can include millions of results.
 It may be infeasible to keep unfiltered query results in browser memory.
@@
 The proxy server can cache results from initial query execution, so that only a filtered result set is subsequently returned to the client.
-2. The proxy server enables us to avoid cross domain XML-HTTP-Request errors in accessing SPARQL endpoints in various domains.
+
+\item The proxy server enables us to avoid cross domain XML-HTTP-Request errors in accessing SPARQL endpoints in various domains.
+
+\end{enumerate}
 
@@
 In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
-The generated SPARQL query will be of the form illustrated in figure \ref{fig:min_query}.
+The generated SPARQL query will be of the form illustrated in Figure \ref{fig:min_query}.
 The word ``input" in the figure represents the keyword specified by the user.
 
-When the results are returned, each result set is structured as the JSON object given in figure \ref{fig:JSON_runquery}.
+When the results are returned, each result set is structured as the JSON object given in Figure \ref{fig:JSON_runquery}.
 This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
 
-When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is far more complex than in the ``Query All" case.
+When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is bit more complex than in the ``Query All" case.
 The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
-This information is passed as a JSON object, whose structure is given in figure \ref{fig:JSON_advquery}, to the proxy server.
+This information is passed as a JSON object, whose structure is given in Figure \ref{fig:JSON_advquery}, to the proxy server.
 
@@
 
-We believe that the unstructured keyword search is likely to encourage users without previous experience running queries on semantic web data to explore the additional features of our system that allow more flexible and advanced queries.
-For example, physicians who are not used to running queries on semantic web data may be more willing to invest time into learning how to find better search results once they see that the system can easily provide them with some initial useful query results using the basic keyword search option.
+We believe that the unstructured keyword search is likely to encourage users without previous experience running queries on SPARQL endpoints to explore the additional features of our system that allow more flexible and advanced queries.
+For example, physicians may be more willing to invest time into learning how to find better search results once they see that the system can easily provide them with some initial useful query results using the basic keyword search option.
 However, the execution of queries using QueryMed will be very different from a user issuing a web search on a search engine such as Google because it exploits the underlying structure of the data exposed by the individual data endpoints.
-
-The user can choose to perform exact queries, or filter results on specific keywords.
-If the user does not know the specific value for a property, they can instead specify a keywords using the “FILTER” option.
-They also can choose logical operators to connect the various parts of their query.
-When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
-as a conjunction.
-When “OR” is used the specified graph pattern is made to disjunct with the rest of the query with the SPARQL UNION operator.
-If no logical operator is specified, “AND” will be used by default.
-The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in figure \ref{fig:SPARQL_adv}.
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-SELECT distinct ?disease WHERE { 
-{?x <http://www.w3.org/2000/01/rdf-schema#label> 
-?disease 
-FILTER regex(?disease, 
-"coronary artery disease", "i"). 
-?x <http://www4.wiwiss.fu-berlin.de/diseasome/
-resource/diseasome/class> 
-<http://www4.wiwiss.fu-berlin.de/diseasome/
-resource/diseaseClass/Cardiovascular>} 
-UNION { ?x <http://www4.wiwiss.fu-berlin.de/
-diseasome/resource/diseasome/associatedGene>
- <http://www4.wiwiss.fu-berlin.de/diseasome/
- resource/genes/ABCA1>.}
-}
-\end{Verbatim}
-\caption{A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.}
-\label{fig:SPARQL_adv}
-\end{figure}
 
@@
 QueryMed is implemented in Java in the backend and JavaScript, HTML and CSS in the frontend.
+In the backend, the Jena library \cite{McBride} is used to run the SPARQL queries and 4-store~\cite{4store} is used as a triple store to provide caching support.
 The JQuery library \cite{Resig} was used to develop an attractive user interface.
-In the backend, the Jena library \cite{McBride} is used to run the SPARQL queries.
 
-\subsection{A Sample Use Case}
+%\subsection{A Sample Use Case}
 
-A physician is interested in finding semantic web resources related to coronary artery disease.
+%A physician is interested in finding semantic web resources related to coronary artery disease.
 She first tries a basic search over all default resources, by entering ``coronary artery disease" in the input search box.
@@
 The slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
-We compared the times required to load properties from several endpoints, by running the query illustrated in figure \ref{fig:SPARQL_props}.
-Timing data are shown in table \ref{timing}.
+We compared the times required to load properties from several endpoints, by running the query illustrated in Figure \ref{fig:SPARQL_props}.
+Timing data are shown in Table \ref{timing}.
+
+\begin{figure}
+\begin{Verbatim}[frame=single]
+SELECT DISTINCT ?property
+WHERE { [] ?property [] }
+ORDER BY ?property
+\end{Verbatim}
+\caption{SPARQL query to retrieve all properties of a given data source}
+\label{fig:SPARQL_props}
+\end{figure}
 
@@
 
+A video illustrating a sample use case can be found at: [USE CASE URL]
+
 \section{Related Work}
@@
 Table \ref{comparison} compares selected features of the QueryMed system with other related systems.
-The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user.
-Other features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the semantic web data.
+The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user, and provides a hybrid interface that enables the users to both query and browse data.
+Other features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data.
 This feature increases the ease of use of our system relative to other similar systems.

### 2010-02-12 docs/www2010/paper.tex
general edits to paper

@@
 
-In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind some of the crucial components of the system.
+In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind crucial components of the system.
 
@@
 The QueryMed user interface is designed to be intuitive for the end user, yet flexible enough to permit a broad range of interesting queries.
-The basic query interface allows the user to run simple queries, and is designed for maximal easy of use.
+The basic query interface allows the user to run simple queries, and is designed for maximal ease of use.
 The advanced search capabilities allow the user to easily construct complex logical queries that take advantage of the underlying structure of the biomedical ontologies.
@@
 
-\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but still provide a flexible tool that is able to perform a range of queries, we provide an uncluttered interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
+\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
 The ``Query All" button will perform a keyword-based query over pre-selected data sources.
@@
 disease. She can then filter the results using additional search terms. For example, she knows that 
-the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view only the most relevant query results.
+the route of administration of the drug that she is interested in is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view only the most relevant query results.
 
-\item \textbf{Advanced Query Interface:} This allows the user to add new data sources, and give values for properties found at a particular selected data source.
-Therefore, the user is able to extend the query from a basic keyword search over all the properties as with the basic query interface, to a more targeted query.
+\item \textbf{Advanced Query Interface:} The Advanced Query Interface allows the user to add new data sources, and to construct a complex SPARQL query that takes advantage of the structure of the underlying biomedical ontologies.
+Thus, the end user can construct a targeted query without previous knowledge of the SPARQL language or of the structure of the relevant biomedical ontologies.
 
 When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources in the system as shown in Figure \ref{fig:select}.
-This feature allows the user to perform their query only over relevant or trusted data sources.
-If none of the desired data sources are available in this list, the user can dynamically add additional data sources as shown in Figure \ref{fig:add_source}.
-The ability to dynamically add additional data sources increases the flexibility of  the QueryMed system, allowing users to query endpoints of interest that are not included in the default list.
+The user may select from this list relevant or trusted data sources that they wish to query.
+The user can also dynamically add additional data sources as shown in Figure \ref{fig:add_source}, allowing users to query endpoints of interest that are not included in the default list.
+
+A physician who is interested in using the QueryMed system to find relevant clinical trials for 
+her patient can use the Advanced Query Interface to add additional relevant datasources .  For example, she might want to use the clinical trial database LinkedCt, which is 
+not in the default set of endpoints.  So, she selects the “Refine 
+Query” option to select the endpoints to search, and then the “Add” option 
+to include an additional endpoint. After entering the name 
+and URL of the LinkedCT endpoint, she can able to search 
+for clinical trials for which her patient may be eligible.
 
@@
 
-When a data source is selected, the properties list is automatically populated with all the distinct properties available at the selected endpoint.
-If the user wants to restrict her query based on one or more properties available, she can specify those.
-This allows users to improve their queries using the underlying structure of the data without having prior knowledge of this structure to obtain more relevant query results.
+When a data source is selected, the properties list is automatically populated with all distinct properties available at the selected endpoint.
+If the user wants to restrict her query based on one or more properties available, she can specify values for the desired properties.
+Because the QueryMed system automatically displays a list of properties to the user, users can improve their queries using the underlying structure of the data to obtain more relevant query results.
 %The properties list is generated by running a SPARQL query of the form given in Figure \ref{fig:SPARQL_props} at the specified data source.
@@
 The user can choose to perform exact queries, or filter results on specific keywords.
-If the user does not know the specific value for a property, they can instead specify a keywords using the “FILTER” option.
-They also can choose logical operators to connect the various parts of their query.
+If the user does not know the specific value for a property, she can instead specify a keywords using the “FILTER” option.
+She also can choose logical operators to connect the various parts of their query.
 When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
@@
 The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in Figure \ref{fig:SPARQL_adv}.
+
+For example, the physician might want to further refine her search for cardiovascular diseases. She 
+uses the advanced search option to construct a boolean query 
+that takes advantage of the underlying structure of the RDF 
+data in the database. She searches Diseasome for a list of 
+diseases whose class is “Cardiovascular” or for which the associated gene is “ABCA1.” Using the QueryMed advanced 
+search interface, the complex SPARQL query corresponding 
+to her question is automatically constructed (Figure \ref{fig:SPARQL_adv}), and she can 
+view the query results conveniently displayed in the Results Browser.
 
@@
 
-For example, the physician might be interested in finding relevant clinical trials for 
-her patient. However, the clinical trial database LinkedCt is 
-not in the default set of endpoints, so she selects the “Refine 
-Query” option to choose additional endpoints to search. She 
-sees a list of default endpoints, and selects the “Add” option 
-to include an additional endpoint. After entering the name 
-and URL of the LinkedCT endpoint, she is able to search 
-for clinical trials for which her patient may be eligible.
-
-\item \textbf{Dynamic Facet Creator/Filter:} This allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint as shown in Figure  \ref{fig:facet_creator}.
+\item \textbf{Dynamic Facet Creator/Filter:} The Dynamic Facet Creator/Filter allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (Figure  \ref{fig:facet_creator}).
 The interface allows displaying of many property lists from different sources simultaneously (grouped by source).
 Therefore, only information relevant to the endpoint the user is currently examining will be visible at any given time.
-
-For example, the physician might want to further refine her search for cardiovascular diseases. She 
-uses the advanced search option to construct a boolean query 
-that takes advantage of the underlying structure of the RDF 
-data in the database. She searches Diseasome for a list of 
-diseases whose class is “Cardiovascular” or for which the as- 
-sociated gene is “ABCA1.” Using the QueryMed advanced 
-search interface, the complex SPARQL query corresponding 
-to her question is automatically constructed (Figure \ref{fig:SPARQL_adv}), and she can 
-view the query results conveniently displayed in the Results Browser.
 
@@
 
-The functionality of the proxy server is to act as an intermediary between the user interface and the remote SPARQL endpoints.  Its functionality is twofold:
+The proxy server acts as an intermediary between the user interface and the remote SPARQL endpoints.  Its functionality is twofold:
 
@@
 \item Execute the SPARQL queries at the relevant remote SPARQL endpoints and consolidate the results to be presented in the user interface.
-\item Cache the results for subsequent queries for refinements to the same query to reduce the network and query execution latency.
+\item Cache the results of the current query so that refinements of the query will have reduced network and query execution latency.
 \end{enumerate}
@@
 \item \textbf{Translator:} The Translator is responsible for translating the user query into valid SPARQL syntax.
-Translator obtains the parameters to construct the query from the input the user specifies in the Query Interface, and plugs them in to a relevant template.
-It specifically relies on two services:
+The Translator obtains the parameters to construct the query from the input the user specifies in the Query Interface, and dynamically constructs a SPARQL query based on the user input.
+The Translator relies on two services:
 
@@
 
-\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries generated by the Translator, and return  the query results as a JSON object given in Figure \ref{fig:JSON_runquery}.
+\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries generated by the Translator, and return  the query results as a JSON object.
 
-\begin{figure}
-\begin{Verbatim}[frame=single]
-{
-   "bindings":
-     [
-        {"source" : "source_label",
-         "uri" : "http://source.com/sparql.",
-     	"vars" : ["variable1","variable2"],
-     	"count":2,
-     	"results" : 
-     	 [
-     	    {"variable1": "Value 1",
-	      "variable1": "Value 2"},
-     	    {"variable1": "Another Value 1",
-               "variable1": "Another Value 2"},
-   	    ...	  		
-     	 ]
-     	}
-     	,
-	{ //Another Source and Result Description
-	...
-	},
-	...
-     ]
- }
-\end{Verbatim}
-\caption{Structure of the JSON ``result" object}
-\label{fig:JSON_runquery}
-\end{figure}
+%\begin{figure}
+%\begin{Verbatim}[frame=single]
+%{
+%   "bindings":
+%  [
+%        {"source" : "source_label",
+%         "uri" : "http://source.com/sparql.",
+ %    	"vars" : ["variable1","variable2"],
+  %   	"count":2,
+   %  	"results" : 
+   %  	 [
+  %   	    {"variable1": "Value 1",
+%	      "variable1": "Value 2"},
+ %    	    {"variable1": "Another Value 1",
+  %             "variable1": "Another Value 2"},
+ %  	    ...	  		
+  %   	 ]
+  %   	}
+  %   	,
+%	{ //Another Source and Result Description
+%	...
+%	},
+%	...
+ %    ]
+% }
+%\end{Verbatim}
+%\caption{Structure of the JSON ``result" object}
+%\label{fig:JSON_runquery}
+%\end{figure}
 
-\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object given in Figure \ref{fig:JSON_advquery}.
+\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
 
-\begin{figure}
-\begin{Verbatim}[frame=single]
-{
- "http://some-source/sparql": 
-  {
-     "property1": {"value1":  "logical operator"},
-     "property2": {"value 2": "logical operator"},
-     "property2": {"value 3": "logical operator"},
-     ...
-  }
-}
-   //Other sources ...
+%\begin{figure}
+%\begin{Verbatim}[frame=single]
+%{
+% "http://some-source/sparql": 
+%  {
+%     "property1": {"value1":  "logical operator"},
+%     "property2": {"value 2": "logical operator"},
+%     "property2": {"value 3": "logical operator"},
+%     ...
+%  }
+%}
+%   //Other sources ...
 
-\end{Verbatim}
-\caption{Structure of the JSON ``query" object}
+%\end{Verbatim}
+%\caption{Structure of the JSON ``query" object}
 %\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
-\label{fig:JSON_advquery}
-\end{figure}
+%\label{fig:JSON_advquery}
+%\end{figure}
 
@@
 
-When the results are returned, each result set is structured as the JSON object given in Figure \ref{fig:JSON_runquery}.
+When the results are returned, each result set is structured as a JSON object.
 This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
 
-When the user runs an advanced query, the payload to be sent to the proxy server to perform the SPARQL query construction and query execution orchestration is bit more complex than in the ``Query All" case.
-The system must specify what data sources are to be queried, what properties are queried, the values for each of the selected properties, and how these property-value pairs relate to the original user-specified query term.
-This information is passed as a JSON object, whose structure is given in Figure \ref{fig:JSON_advquery}, to the proxy server.
+When the user runs an advanced query, the proxy server takes as input the data sources to be queried, the properties to be queried, the values for each of the selected properties, and the relationship between the property-value pairs and the original user-specified query term.
+This information is passed to the proxy server as a JSON object.
 
@@
 
-We believe that the unstructured keyword search is likely to encourage users without previous experience running queries on SPARQL endpoints to explore the additional features of our system that allow more flexible and advanced queries.
-For example, physicians may be more willing to invest time into learning how to find better search results once they see that the system can easily provide them with some initial useful query results using the basic keyword search option.
-However, the execution of queries using QueryMed will be very different from a user issuing a web search on a search engine such as Google because it exploits the underlying structure of the data exposed by the individual data endpoints.
+The basic query interface is designed to be as simple as possible, so that users who have little previous experience running queries on SPARQL endpoints can easily and  explore rapidly find query results.
+The advanced query functionality of our system allows users the flexibility to construct complex SPARQL queries.
+The QueryMed system allows the user to construct queries that take advantage of the underlying structure of the data exposed by the individual data endpoints, without requiring previous knowledge of this structure.
 
@@
 Queries to the SMART system are constructed using a descriptive logic query written in the natural-language like Manchester OWL syntax \cite{Battista}.
-However, it cannot be expected that the end users know how  to use the Manchester OWL syntax to write queries, whereas in QueryMed, users can construct queries intuitively by giving keywords for user selected properties.
+However, it cannot be expected that the end users know how  to use the Manchester OWL syntax to write queries, whereas in QueryMed, users can construct queries intuitively by specifying keywords for user selected properties.
 
@@
 Another unique feature of our system is the property-based advanced query interface.
-This enables users to take advantage of the structure of the underlying ontologies used to represent the data without prior knowledge of the ontology structures.
+This interface enables users to take advantage of the structure of the underlying ontologies used to represent the data without prior knowledge of the ontology structures.
 

### 2010-02-12 trunk/docs/www2010/paper.tex
added BioGateway reference; completed table; other assorted edits

@@
 
-Publicly available data in the biomedical domain has grown significantly over the recent years.
-For example, this data include data on drug discovery~\cite{Sharp, Goble}, clinical trial information that are publicly available at ClinicalTrails.gov, disease genes, phenotypes published by the Online Mendelian Inheritance in Man (OMIM), and etc.
-With the linked open data movement, the semantic web community has been very proactive in converting these datasets to RDF~\cite{LinkingData}.
-In fact, the biomedical domain is among the early successes of adopting semantic web due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
+The quantity of publicly available data in the biomedical domain has dramatically increased over recent years.
+Publicly available biomedical resources include data on drug discovery~\cite{Sharp, Goble}, clinical trials, diseases, disease genes, and phenotypes.
+With the linked open data movement, the semantic web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
+In fact, the biomedical domain is among the early successes of the semantic web due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
 
@@
 
-Although a plethora of useful biomedical data is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
-Many currently available systems are linked data data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner, but lack support for efficiently filtering and querying the data.
-Therefore, there is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}.
+To allow end users to exploit the abundance of useful biomedical data that is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
+There is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}, since many currently available systems are linked data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner but lack support for filtering and querying the data.
 
-For an example in the biomedical domain, a physician may know her patient's personal information, symptoms, current medications, and genotype.
+Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple endpoints.
+For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
 She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.
@@
 Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
-Suppose the physician is interested in information on “coronary artery disease” (which she thinks is the cause of her patient's symptoms).
-The information about the disease can be found in Diseasome, a database that lists diseases, disease genes, and associations between them~\cite{Diseasome}.
-The physician might also want to know the set of drugs in DailyMed~\cite{DailyMed}, and Drugbank~\cite{Drugbank} that can be used to treat coronary artery disease.
-Information on clinical trial for which her patient may be eligible can be found in LinkedCT~\cite{LinkedCT}.
 
-Due to the large number of databases  that the physician needs to query in order to find an answer to her single question, she is likely to find useful, a system that can automatically run queries over multiple data sources.
+Since the physician must search many databases in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.
 Also, the physician may not know SPARQL query syntax, the location of the SPARQL endpoints, or the structure of the relevant ontologies.
@@
 
-Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF.
+Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
 Almost all of these data sources are interlinked with each other.
@@
 
-QueryMed allows users to query many biomedical data sources from a single interface.
-Although there is a default list of sources that the user can choose to query from, this list is  not fixed, and can be extended very easily.
-The system was designed with non-expert users of SPARQL in mind, and presents a very intuitive user interface that anybody with basic biomedical domain knowledge can use.
+The QueryMed system allows users to easily query multiple biomedical data sources.
+Queries can be run against a default list of datasources, or against a set of user-selected SPARQL endpoint.
+The end user can easily input additional endpoints in order to utilize resources that are not included in the default list.
+The system automatically translates the user input into a SPARQL query for each individual endpoint, combines the results, and returns them to the user.
+The user can choose to refine the query by iteratively modifying the original query terms, and by filtering the result list.
+The advanced query functionality of the QueryMed system allows the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the ontologies.
+The simple user interface of the QueryMed system is designed to be intuitive for users with no knowledge of the SPARQL query language.
 
-The system functionality in a nutshell is as follows: A user submits a query from the Query Interface to the SPARQL endpoints via the proxy server.
-This is translated into a SPARQL query for each individual endpoint.
-Once the results are obtained from the SPARQL endpoints, they are presented to the user via the Result Browser.
-The user can then choose to refine the query by changing some of the original query terms in the Dynamic Facet Creator/Filter.
-This is an iterative process, where the user can refine the query by adding new sources, and tweaking values given to the properties.
-
-In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind some of the crucial components of the system.
+In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind crucial components of the system.
 
@@
 A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
-The main components of the system are the user interface and the proxy server that retrieves biomedical data from remote SPARQL endpoints.
-A more detailed illustration of the QueryMed system highlighting all the components necessary for system functionality is given in Figure \ref{fig:arch_details}.
+The main components of the system are the user interface and the proxy server that takes input from the user and retrieves relevant biomedical data from remote SPARQL endpoints.
+After a user submits a query using the user interface, the query is translated by the proxy server into individual SPARQL queries for each remote endpoint.
+The query results are returned from the remote endpoints, combined by the proxy server, and presented to the user.
+A detailed illustration of the parts of the QueryMed system is shown in Figure \ref{fig:arch_details}.
 
@@
 
-The user interface is one of the primary components of the QueryMed system. It comprises of the following:
+The QueryMed user interface is designed to be intuitive for the end user, yet flexible enough to permit a broad range of interesting queries.
+The basic query interface allows the user to run simple queries, and is designed for maximal ease of use.
+The advanced search capabilities allow the user to easily construct complex logical queries that take advantage of the underlying structure of the biomedical ontologies.
+The user interface also allows the user to iteratively refine queries, and displays the query results, which have been retrieved from multiple SPARQL endpoints.
+The main components of the user interface are as follows:
 
@@
 disease. She can then filter the results using additional search terms. For example, she knows that 
-the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view the most relevant search terms.
+the route of administration of the drug that she is interested in is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view only the most relevant query results.
 
-\item \textbf{Advanced Query Interface:} This allows the user to add new data sources, and give values for properties found at a particular selected data source.
-Therefore, the user is able to extend the query from a basic keyword search over all the properties as with the basic query interface, to a more targeted query.
+\item \textbf{Advanced Query Interface:} The Advanced Query Interface allows the user to add new data sources, and to construct a complex SPARQL query that takes advantage of the structure of the underlying biomedical ontologies.
+Thus, the end user can construct a targeted query without previous knowledge of the SPARQL language or of the structure of the relevant biomedical ontologies.
 
 When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources in the system as shown in Figure \ref{fig:select}.
-This feature allows the user to perform their query only over relevant or trusted data sources.
-If none of the desired data sources are available in this list, the user can dynamically add additional data sources as shown in Figure \ref{fig:add_source}.
-The ability to dynamically add additional data sources increases the flexibility of  the QueryMed system, allowing users to query endpoints of interest that are not included in the default list.
+The user may select from this list relevant or trusted data sources that they wish to query.
+The user can also dynamically add additional data sources as shown in Figure \ref{fig:add_source}, allowing users to query endpoints of interest that are not included in the default list.
+
+A physician who is interested in using the QueryMed system to find relevant clinical trials for 
+her patient can use the Advanced Query Interface to add additional relevant datasources .  For example, she might want to use the clinical trial database LinkedCt, which is 
+not in the default set of endpoints.  So, she selects the “Refine 
+Query” option to select the endpoints to search, and then the “Add” option 
+to include an additional endpoint. After entering the name 
+and URL of the LinkedCT endpoint, she can able to search 
+for clinical trials for which her patient may be eligible.
 
@@
 \hline
-\textbf{SMART}  & & & Yes & No & No & Yes & Yes  \\
+\textbf{SMART}  & No & Yes & Yes & No & No & Yes & Yes  \\
 \hline
-\textbf{DARQ}  & &   & Yes & N/A & No & Yes & No  \\
+\textbf{DARQ}  & No &  No & Yes & N/A & No & Yes & No  \\
 \hline
-\textbf{GoWeb}   & &  & Yes & No & Yes & No & Yes  \\
+\textbf{GoWeb}   & Yes & No  & Yes & No & Yes & No & Yes  \\
 \hline
-\textbf{BioGateway}   & &  & Yes & No & No & No & Yes  \\
+\textbf{BioGateway}   & Yes & Yes & Yes & No & No & No & Yes  \\
 \hline
-\textbf{Twinkle}   & &  & Yes & No & No & Yes & Yes  \\
+\textbf{Twinkle}   & No & No & Yes & No & No & Yes & Yes  \\
 \hline
@@
 Queries to the SMART system are constructed using a descriptive logic query written in the natural-language like Manchester OWL syntax \cite{Battista}.
-However, it cannot be expected that the end users know how  to use the Manchester OWL syntax to write queries, whereas in QueryMed, users can construct queries intuitively by giving keywords for user selected properties.
+However, it cannot be expected that the end users know how  to use the Manchester OWL syntax to write queries, whereas in QueryMed, users can construct queries intuitively by specifying keywords for user selected properties.
 
-GoWeb is another system designed for answering queries on biomedical data.
-It allows users to run traditional keyword-based web search with ontology search features \cite{Dietze}.
+GoWeb and BioGateway are two other systems designed for answering queries on biomedical data.
+GoWeb allows users to run traditional keyword-based web search with ontology search features \cite{Dietze, Antezana}.
 After a keyword search, documents can be filtered based on the biomedical annotations they contain.
 However, in GoWeb the exact sources queried are not transparent and cannot be selected or modified by the user as in QueryMed.
+BioGateway provides a web interface to query multiple SPARQL endpoints, but does not allow the user to dynamically add endpoints.
 
@@
 Because of the unique features of our system, we believe it will be of use to the biomedical community.
-We also believe tools such as QueryMed will entice more people to expose their data as linked open data.
+We also believe that systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.
 

### 2010-02-12 docs/www2010/paper.tex
additional minor edits

@@
 The main components of the system are the user interface and the proxy server that takes input from the user and retrieves relevant biomedical data from remote SPARQL endpoints.
-After a user submits a query using the user interface, the query is translated by the proxy server into individual SPARQL queries for each remote endpoint.
+After a user submits a query from the user interface, the query is translated by the proxy server into individual SPARQL queries for each remote endpoint.
 The query results are returned from the remote endpoints, combined by the proxy server, and presented to the user.
@@
 
-\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
+\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered basic search interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
 The ``Query All" button will perform a keyword-based query over pre-selected data sources.
-The advanced search option ``Refine Query" allows the user to view additional query options that are handled in the Advanced Query Interface.
+The advanced search option ``Refine Query" allows the user to view additional query options that are handled by the Advanced Query Interface.
 
  For example, a physician interested in finding semantic web resources related to coronary artery disease could use “coronary artery disease”  as a search keyword in the input search box and select the "Query All" option.
-The simplicity of the basic query interface allows this end user to rapidly and easily browse semantic web resources of interest.
+The simplicity of the basic query interface allows the physician to rapidly and easily browse semantic web resources of interest.
 
@@
 
-\item \textbf{Result Browser:} The user can view their query results organized by source in the Result Browser.
+\item \textbf{Result Browser:} The user can view his query results organized by source in the Result Browser.
 The results are presented in a table with pagination.
@@
 \item \textbf{Advanced Query Interface:} The Advanced Query Interface allows the user to add new data sources, and to construct a complex SPARQL query that takes advantage of the structure of the underlying biomedical ontologies.
-Thus, the end user can construct a targeted query without previous knowledge of the SPARQL language or of the structure of the relevant biomedical ontologies.
+Thus, the end user can construct a targeted query without previous knowledge of the SPARQL language or the structure of the relevant biomedical ontologies.
 
 When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources in the system as shown in Figure \ref{fig:select}.
-The user may select from this list relevant or trusted data sources that they wish to query.
+The user may select from this list relevant trusted data sources that they wish to query.
 The user can also dynamically add additional data sources as shown in Figure \ref{fig:add_source}, allowing users to query endpoints of interest that are not included in the default list.
@@
 
-The basic query interface is designed to be as simple as possible, so that users who have little previous experience running queries on SPARQL endpoints can easily and  explore rapidly find query results.
+The basic query interface is designed to be as simple as possible, so that users who have little previous experience running queries on SPARQL endpoints can easily and rapidly find query results.
 The advanced query functionality of our system allows users the flexibility to construct complex SPARQL queries.
@@
 \hline
-\textbf{SMART}  & & & Yes & No & No & Yes & Yes  \\
+\textbf{SMART}  & No & Yes & Yes & No & No & Yes & Yes  \\
 \hline
-\textbf{DARQ}  & &   & Yes & N/A & No & Yes & No  \\
+\textbf{DARQ}  & No &  No & Yes & N/A & No & Yes & No  \\
 \hline
-\textbf{GoWeb}   & &  & Yes & No & Yes & No & Yes  \\
+\textbf{GoWeb}   & Yes & No  & Yes & No & Yes & No & Yes  \\
 \hline
-\textbf{BioGateway}   & &  & Yes & No & No & No & Yes  \\
+\textbf{BioGateway}   & Yes & Yes & Yes & No & No & Yes & Yes  \\
 \hline
-\textbf{Twinkle}   & &  & Yes & No & No & Yes & Yes  \\
+\textbf{Twinkle}   & No & No & Yes & No & No & Yes & Yes  \\
 \hline
@@
 Queries to the SMART system are constructed using a descriptive logic query written in the natural-language like Manchester OWL syntax \cite{Battista}.
-However, it cannot be expected that the end users know how  to use the Manchester OWL syntax to write queries, whereas in QueryMed, users can construct queries intuitively by specifying keywords for user selected properties.
+Major differences between the QueryMed system and SMART include the ability to construct queries intuitively by specifying keywords for user selected properties, to iteratively refine query results, and to dynamically add additional endpoints in the QueryMed system.
 
-GoWeb is another system designed for answering queries on biomedical data.
-It allows users to run traditional keyword-based web search with ontology search features \cite{Dietze}.
-After a keyword search, documents can be filtered based on the biomedical annotations they contain.
-However, in GoWeb the exact sources queried are not transparent and cannot be selected or modified by the user as in QueryMed.
+GoWeb and BioGateway are two additional systems designed for answering queries on biomedical data  \cite{Dietze, Antezana}.
+GoWeb allows users to a hybrid search, running keyword-based queries and then filtering based on ontological concepts.
+However, while GoWeb functions as a search engine that incorporates ontological background knowledge to improve search results, QueryMed is a query translation system.
+BioGateway provides a web interface to query a provided SPARQL endpoint that integrates multiple biomedical resources, but does not allow the user to run queries over multiple endpoints or to dynamically add endpoints.
 
@@
 This feature increases the ease of use of our system relative to other similar systems.
-Furthermore, the Javascript based user interface of the QueryMed system, implemented using the JQuery library, makes our user interface particularly attractive, easy to interact with, and capable of handling a variety of user input events.
+Furthermore, the Javascript-based user interface of the QueryMed system, implemented using the JQuery library, makes our user interface particularly attractive, easy to interact with, and capable of handling a variety of user input events.
 Another unique feature of our system is the property-based advanced query interface.

### 2010-02-15 trunk/docs/www2010/paper.tex
added cache stuff

@@
 We have developed an open-source SPARQL query builder and result set visualizer for biomedical data, QueryMed, that allows end users to easily construct and run translational medicine queries across multiple data sources.
+\smallskip
 
@@
 %The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
+\smallskip
 
 To allow end users to exploit the abundance of useful biomedical data that is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
-There is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}, since many currently available systems are linked data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner but lack support for filtering and querying the data.
+There is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}, since many currently available systems are linked data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner but lack support for filtering and querying the data in a user friendly manner.
+
+\smallskip
 
@@
 
+\smallskip
+
 Since the physician must search many databases in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.
@@
 
-This paper is organized as follows: Section \ref{background} provides background information on the semantic web and its relevance for the biomedical domain.
+\smallskip
+
+This paper is organized as follows: Section \ref{background} provides background information on the semantic web and its relevance for modeling data in the biomedical domain.
 Section \ref{system} describes our system.
 Section \ref{related} discusses related work and illustrates how QueryMed differs from previous systems.
-Finally, section \ref{conclusion} outlines future work and summarizes the contributions of our system.
+Section \ref{future-work} outlines future work giving an outlook for Querymed.
+Finally section \ref{conclusion} summarizes the contributions of our system.
 
@@
 
+\smallskip
+
 Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
@@
 These semantic web resources represent valuable repositories of information that can be automatically mined for applications that require biological knowledge.
+
+\smallskip
 
@@
 
+\smallskip
+
 \section{Design \& Implementation}
@@
 The simple user interface of the QueryMed system is designed to be intuitive for users with no knowledge of the SPARQL query language.
+
+\smallskip
 
@@
 
-\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
+\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered basic search interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
 The ``Query All" button will perform a keyword-based query over pre-selected data sources.
-The advanced search option ``Refine Query" allows the user to view additional query options that are handled in the Advanced Query Interface.
+The advanced search option ``Refine Query" allows the user to view additional query options that are handled by the Advanced Query Interface.
+
+\smallskip
 
  For example, a physician interested in finding semantic web resources related to coronary artery disease could use “coronary artery disease”  as a search keyword in the input search box and select the "Query All" option.
-The simplicity of the basic query interface allows this end user to rapidly and easily browse semantic web resources of interest.
+The simplicity of the basic query interface allows the physician to rapidly and easily browse semantic web resources of interest.
 
@@
 This allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+
+\smallskip
 
@@
 
+\smallskip
+
 Once the properties are returned, they will be displayed in the user interface as shown in Figure \ref{fig:add_props}.
@@
 This allows the user to understand the keywords that will be most useful to search on each specific property.
+
+\smallskip
 
@@
 The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in Figure \ref{fig:SPARQL_adv}.
+
+\smallskip
 
@@
 
+\smallskip
+
 When the results are returned, each result set is structured as a JSON object.
 This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
+
+\smallskip
 
@@
 
-%A video illustrating this sample use case can be found at: \\
-http://www.youtube.com/watch?v=JCr7vryS9Vg
-
 \subsection{Performance}
 
-The slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
+\begin{table*}
+\begin{center}
+%\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
+\begin{tabular}{| l | p{1.2cm}| p{1.2cm} | p{1.2cm} | p{1.2cm} | p{1.2cm} | p{1.2cm} | p{1.2cm} | p{1.2cm} |}
+\hline
+& \multicolumn{2}{c}{\textbf{Dieseasome} } \vline & \multicolumn{2}{c}{\textbf{Dailymed}} \vline & \multicolumn{2}{c}{\textbf{Drugbank}} \vline & \multicolumn{2}{c}{\textbf{Sider}} \vline \\
+\hline
+& \textbf{Without Caching} & \textbf{With Caching} & \textbf{Without Caching} & \textbf{With Caching} & \textbf{Without Caching} & \textbf{With Caching} & \textbf{Without Caching} & \textbf{With Caching} \\
+\hline
+\textbf{1st Trial}   & 3.45 s &  134 ms & 1.77 s &   384 ms & 9.51 s & 60 ms & 14.64 s & 23 ms \\
+\hline
+\textbf{2nd Trial}  & 1.61 s & 31 ms & 1.57 s &  32 ms  & 9.34 s & 31 ms & 2.94 s & 7 ms \\
+\hline
+\textbf{3rd Trial}  & 1.71 s &  7 ms & 1.66 s &   11 ms & 9.06 s & 23 ms & 2.86 s & 6 ms\\
+\hline
+\end{tabular}
+\caption{Running times to retrieve all the properties from selected endpoints}
+\label{timing}
+\end{center}
+\end{table*}
+
+We observed that the slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
 We compared the times required to load properties from several endpoints, by running the query illustrated in Figure \ref{fig:SPARQL_props}.
 Timing data are shown in Table \ref{timing}.
+
+\smallskip
+
+For all selected endpoints, without local data (i.e.
+no cache on the proxy server), the property values took longer to load.
+The difference in running time between \emph{with local data} and \emph{without local data} is approximately three orders of magnitude.
+The network latency and the slow performance of the remote SPARQL endpoints as compared to the local data cache can be accounted for the increase in running time for the queries in the two cases.
+We also noted that the running time of the subsequent iterations of the same query is significantly less than the initial query, probably due to browser caching.
+When comparing the query execution time amongst the data sources, we see that the Drugbank takes the longest, probably due to the greater size of the Drugbank data as compared to the other three data sources~\footnote{As of February 15\textsuperscript{th} 2010, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, Sider contains 192,515 triples, and DrugBank contains 765,936 triples) \cite{Diseasome, DailyMed, Sider, Drugbank}}.
+
+\smallskip
+
+Therefore, it is fair to say that bottleneck in the running time of our system is running queries that must retrieve many triples from slow remote SPARQL endpoints.
+By managing our own cache to contain the data most likely to be needed on the proxy server, we were able to reduce running time further.
+
+\smallskip
 
@@
 \end{Verbatim}
-\caption{SPARQL query to retrieve all properties of a given data source}
+\caption{SPARQL query to retrieve all Properties. This was used to measure the performance of implementing a local cache.}
 \label{fig:SPARQL_props}
@@
 
-For all selected endpoints, the property values take longer to load on the initial query than in subsequent iterations, probably due to browser caching.
-The running time to acquire the list of property values for the Drugbank endpoint is much larger than the time for the other two endpoints, probably due to the greater size of the Drugbank database.
-(As of 12/10/09, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, and DrugBank contains 765,936 triples).
-
-\begin{table}
-\begin{tabular}{|l||l|l|l|}
-\hline
- & \textbf{Dieseasome} & \textbf{Dailymed} & \textbf{Drugbank} \\
-\hline
-\hline
-\textbf{1st Trial}   & 3.45 s & 1.77 s  & 9.51 s  \\
-\textbf{2nd Trial}  & 1.61 s & 1.57 s & 9.34 s \\
-\textbf{3rd Trial}  & 1.71 s & 1.66 s & 9.06 s \\
-\hline
-\end{tabular}
-\caption{Running times to retrieve all the properties from selected endpoints}
-\label{timing}
-\end{table}
+\begin{figure}
+\centering
+\epsfig{file=images/chart.jpg, width=3in}
+\caption{Comparison of Query Execution Times at Various SPARQL Endpoints With and Without Local Data (Cache)}
+\label{fig:chart}
+\end{figure}
 
@@
 The source code for the Querymed system is available at the QueryMed Google Code project: \\
-http://code.google.com/p/querymed/
+http://code.google.com/p/querymed
 
-A video illustrating a sample use case can be found at: [USE CASE URL]
+\smallskip
+
+A video illustrating a sample use case can be found at: \\
+http://dig.csail.mit.edu/2010/Papers/www-ws-colab-science/\\
+videos
 
@@
 \hline
-\hline
 \textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes & Yes & Yes  \\
@@
 \hline
-\textbf{BioGateway}   & Yes & Yes & Yes & No & No & No & Yes  \\
+\textbf{BioGateway}   & Yes & Yes & Yes & No & No & Yes & Yes  \\
 \hline
@@
 
+\smallskip
+
 The SMART query tool is a web-based application designed to allow biologists to run SPARQL queries over multiple endpoints.
 Queries to the SMART system are constructed using a descriptive logic query written in the natural-language like Manchester OWL syntax \cite{Battista}.
-However, it cannot be expected that the end users know how  to use the Manchester OWL syntax to write queries, whereas in QueryMed, users can construct queries intuitively by specifying keywords for user selected properties.
+Major differences between the QueryMed system and SMART include the ability to construct queries intuitively by specifying keywords for user selected properties, to iteratively refine query results, and to dynamically add additional endpoints in the QueryMed system.
 
-GoWeb and BioGateway are two other systems designed for answering queries on biomedical data.
-GoWeb allows users to run traditional keyword-based web search with ontology search features \cite{Dietze, Antezana}.
-After a keyword search, documents can be filtered based on the biomedical annotations they contain.
-However, in GoWeb the exact sources queried are not transparent and cannot be selected or modified by the user as in QueryMed.
-BioGateway provides a web interface to query multiple SPARQL endpoints, but does not allow the user to dynamically add endpoints.
+\smallskip
+
+GoWeb and BioGateway are two additional systems designed for answering queries on biomedical data  \cite{Dietze, Antezana}.
+GoWeb allows users to a hybrid search, running keyword-based queries and then filtering based on ontological concepts.
+However, while GoWeb functions as a search engine that incorporates ontological background knowledge to improve search results, QueryMed is a query translation system.
+BioGateway provides a web interface to query a provided SPARQL endpoint that integrates multiple biomedical resources, but does not allow the user to run queries over multiple endpoints or to dynamically add endpoints.
+
+\smallskip
 
@@
 
+\smallskip
+
 Most SPARQL query engines are designed to run queries against individual endpoints.
@@
 There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambite} systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
+
+\smallskip
 
@@
 \section{Future Work}
-\label{conclusion}
+\label{future-work}
 
-The bottleneck in the running time of our system is running queries that must retrieve many triples from slow remote SPARQL endpoints.
-By caching relevant data locally, the running time could be significantly improved.
-We observed that the QueryMed system performs significantly faster the second time that a query is run on a given resource possibly due to browser caching.
-By managing our own cache to contain the data most likely to be needed on the proxy server, we could reduce running time still further.
+Our system currently allows only a restricted set of SPARQL queries.
+Supporting additional types of queries and including query optimization functionality could increase both the flexibility and speed of the QueryMed system.
+By drawing on the expressivity of the SPARQL language and the information contained in the ontologies used to represent the data, it would be possible to extend our system into an intelligent reasoning system on biomedical data which allows physicians to enter sophisticated, complex queries and find relevant biomedical results.
+
+\smallskip
+
 Another approach that could reduce running time still further, especially for users with a slow network connection, would be to create a complementary standalone application that gives the user the option at startup time of loading all the required RDF data.
@@
 
-Our system currently allows only a restricted set of SPARQL queries.
-Supporting additional types of queries and including query optimization functionality could increase both the flexibility and speed of the QueryMed system.
-By drawing on the expressivity of the SPARQL language and the information contained in the ontologies used to represent the data, it would be possible to extend our system into an intelligent reasoning system on biomedical data which allows physicians to enter sophisticated, complex queries and find relevant biomedical results.
+\smallskip
 
@@
 
+\smallskip
+
 Another feature that we have begun to implement in our system is an autocompletion feature to automatically retrieve the distinct values for any given property.
 This feature will allow the user to automatically see all valid choices for each property, and will make the system easier to use and reduce empty query result sets.
+
+\smallskip
 
@@
 \section{Conclusion}
+\label{conclusion}
 
@@
 
-\section{Acknowledgments}
+%\section{Acknowledgments}
 

### 2010-02-15 trunk/docs/www2010/paper.tex
a couple of minor changes

@@
 Therefore, it is fair to say that bottleneck in the running time of our system is running queries that must retrieve many triples from slow remote SPARQL endpoints.
-By managing our own cache to contain the data most likely to be needed on the proxy server, we were able to reduce running time further.
+By managing our own cache to contain the data most likely to be needed on the proxy server, we were able to dramatically decrease running time.
 
@@
 \epsfig{file=images/chart.jpg, width=3in}
-\caption{Comparison of Query Execution Times at Various SPARQL Endpoints With and Without Local Data (Cache)}
+\caption{Comparison of Query Execution Times (measured in milliseconds) at Various SPARQL Endpoints With and Without Local Data (Cache) (average of three trials)}
 \label{fig:chart}
@@
 The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
-Because of the unique features of our system, we believe it will be of use to the biomedical community.
-We also believe that systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.
+Because our system is flexible and easy to use, we believe it will be of use to the biomedical community.
+We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.
 

### 2010-02-15 trunk/docs/www2010/paper.tex
final submission

@@
 
-QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, runs queries across multiple SPARQL endpoints, and is designed to be accessible to users who do not know the structure of the underlying ontologies used in describing the datasets, or the SPARQL query language to query the data.
+QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, and runs queries across multiple SPARQL endpoints.
+It is designed to be accessible to users who are not familiar with the structure of the underlying ontologies used in describing the datasets, or with the SPARQL query language.
 The system allows users to select the data sources that they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.
@@
 
-\keywords{Biomedical Ontologies, SPARQL, Query Federation, Query Building, Semantic Web, User Interfaces}
+\keywords{Biomedical Ontologies, SPARQL Query Building, Query Federation, Semantic Web, User Interfaces}
 
@@
 With the linked open data movement, the semantic web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
-In fact, the biomedical domain is among the early successes of the semantic web due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
+In fact, the biomedical domain is among the early successes of the semantic web, due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
 
@@
 
-To allow end users to exploit the abundance of useful biomedical data that is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
-There is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}, since many currently available systems are linked data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner but lack support for filtering and querying the data in a user friendly manner.
+To allow end users to exploit the abundance of biomedical data that is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
+There is also a  need for efficient hybrid interfaces that allow both browsing and querying data~\cite{Jentzsch}.
+Many currently available systems are linked data browsers such as the Tabulator~\cite{Tabulator} which permit users to navigate the data in an exploratory manner, but lack support for filtering and querying the data in a user friendly manner.
 
@@
 
-Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple endpoints.
+Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple data sources.
 For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
@@
 Although the physician has a single question--``based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?"--there is no single data source that the physician can use to answer this question.
-The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and GO \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
+The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and Gene Ontology \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
 Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
@@
 
-Since the physician must search many databases in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.
+Since the physician must search many data sources in order to find an answer to her single question, she requires a system that can automatically run queries over multiple SPARQL endpoints.
 Also, the physician may not know SPARQL query syntax, the location of the SPARQL endpoints, or the structure of the relevant ontologies.
-She is likely to want an intuitive way to query and to display the query result.
-Developing intuitive ways to query multiple data sources and display results is both an important and a challenging problem.
+She is likely to want an intuitive way to query and to display the query results.
+Developing intuitive ways to query multiple SPARQL endpoints and to display results is both an important and a challenging problem.
 Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.
@@
 Section \ref{related} discusses related work and illustrates how QueryMed differs from previous systems.
-Section \ref{future-work} outlines future work giving an outlook for Querymed.
+Section \ref{future-work} outlines future work.
 Finally section \ref{conclusion} summarizes the contributions of our system.
+
+\pagebreak
 
@@
 \label{background}
+
+\begin{figure}
+\centering
+\epsfig{file=images/architecture_overview.jpg}
+\caption{QueryMed Architecture Overview}
+\label{fig:arch_overview}
+\end{figure}
 
@@
 Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
-Almost all of these data sources are interlinked with each other.
+Almost all of these data sources are interlinked.
 Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists.
@@
 One challenge is constructing systems that allow end users to run intuitive queries on biomedical data.
-Users of biomedical resources are likely to have extensive domain knowledge, but be unfamiliar with the SPARQL query language syntax and with the structure of biomedical ontologies.
-Almost all the SPARQL endpoints available today are offering a generic query interface that requires users to write SPARQL queries.
-Also, no hints about the structure of data available from the SPARQL endpoint is given as the user is constructing the SPARQL query.
-The user has to essentially construct the queries that are needed for her purpose by trial and error.
+Users of biomedical resources are likely to have extensive domain knowledge, but be unfamiliar with query language syntax and with the structure of biomedical ontologies.
+Almost all the SPARQL endpoints available today offer generic query interfaces that require users to manually write their queries.
 This can be a daunting task  especially for someone who is new to semantic web technologies.
-Therefore, it is important to design user-friendly systems that allow users to take advantage of the wealth of structured biological knowledge available on the semantic web.
-Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data are often distributed among many sources \cite{Pasquier}.
+Therefore, it is important to design user-friendly systems that allow users to take advantage of the wealth of structured biomedical knowledge available on the semantic web.
+Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biomedical data are often distributed among many sources \cite{Pasquier}.
 
@@
 
-The QueryMed system allows users to easily query multiple biomedical data sources.
-Queries can be run against a default list of datasources, or against a set of user-selected SPARQL endpoint.
+The QueryMed system allows the user to easily query multiple biomedical data sources.
+Queries can be run against a default list of SPARQL endpoints, or against a set of user-selected endpoints.
 The end user can easily input additional endpoints in order to utilize resources that are not included in the default list.
-The system automatically translates the user input into a SPARQL query for each individual endpoint, combines the results, and returns them to the user.
-The user can choose to refine the query by iteratively modifying the original query terms, and by filtering the result list.
+The system automatically translates the user input into a SPARQL query for each individual endpoint, executes the query, combines the results, and returns them to the user.
+The user can choose to refine the query by iteratively modifying the original query terms, and by filtering the result set.
 The advanced query functionality of the QueryMed system allows the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the ontologies.
@@
 \centering
-\epsfig{file=images/architecture_overview.jpg}
-\caption{QueryMed Architecture Overview}
-\label{fig:arch_overview}
-\end{figure}
-
-\begin{figure}
-\centering
 \epsfig{file=images/architecture_detail.jpg}
@@
 A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
-The main components of the system are the user interface and the proxy server that takes input from the user and retrieves relevant biomedical data from remote SPARQL endpoints.
+The main components of the system are the user interface and the proxy server.
 After a user submits a query from the user interface, the query is translated by the proxy server into individual SPARQL queries for each remote endpoint.
 The query results are returned from the remote endpoints, combined by the proxy server, and presented to the user.
-A detailed illustration of the parts of the QueryMed system is shown in Figure \ref{fig:arch_details}.
+Figure \ref{fig:arch_details} presents a detailed illustration of the parts of the QueryMed system.
 
@@
 The basic query interface allows the user to run simple queries, and is designed for maximal ease of use.
-The advanced search capabilities allow the user to easily construct complex logical queries that take advantage of the underlying structure of the biomedical ontologies.
+The advanced search capabilities enable the user to easily construct complex logical queries that take advantage of the underlying structure of the biomedical ontologies.
 The user interface also allows the user to iteratively refine queries, and displays the query results, which have been retrieved from multiple SPARQL endpoints.
+
+\smallskip
+
 The main components of the user interface are as follows:
@@
 
-\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered basic search interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
-The ``Query All" button will perform a keyword-based query over pre-selected data sources.
-The advanced search option ``Refine Query" allows the user to view additional query options that are handled by the Advanced Query Interface.
+\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered basic search interface as shown in Figure \ref{fig:initial_ui}.
+The ``Query All" button performs a keyword-based query over the default set of  data sources.
+The advanced search option ``Refine Query" allows the user to access the Advanced Query Interface.
 
@@
 
- For example, a physician interested in finding semantic web resources related to coronary artery disease could use “coronary artery disease”  as a search keyword in the input search box and select the "Query All" option.
+ For example, a physician interested in finding semantic web resources related to coronary artery disease could use ``coronary artery disease"  as a search keyword in the input search box and select the ``Query All" option.
+This query searches all default data sources for resources related to coronary artery disease.
 The simplicity of the basic query interface allows the physician to rapidly and easily browse semantic web resources of interest.
@@
 
-\item \textbf{Result Browser:} The user can view his query results organized by source in the Result Browser.
+\item \textbf{Result Browser:} The user can view her query results organized by source in the Result Browser.
 The results are presented in a table with pagination.
 The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
-This allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+This allows the user to perform additional filtering on the query results to display the most relevant results.
+This feature is particularly useful for refining queries that return a large number of results.
+
+\begin{figure}
+\centering
+\epsfig{file=images/results.jpg, width=3.5in}
+\caption{Sample Result Table}
+\label{fig:results}
+\end{figure}
 
@@
 
-For example, after searching for "coronary artery disease" in the basic query interface, the physician will see displayed in the Result Browser a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery
+For example, after searching for ``coronary artery disease" in the basic query interface, the physician will see displayed in the Result Browser a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery
 disease. She can then filter the results using additional search terms. For example, she knows that 
-the route of administration of the drug that she is interested in is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view only the most relevant query results.
+the route of administration of the drug that she is interested in is injection, so she filters the drug query results on the route of administration field using the query term ``injection."   The additional filtering capabilities of the Result Browser allow her  either to browse a large number of query results, or refine her search to view only the most relevant results.
 
 \item \textbf{Advanced Query Interface:} The Advanced Query Interface allows the user to add new data sources, and to construct a complex SPARQL query that takes advantage of the structure of the underlying biomedical ontologies.
-Thus, the end user can construct a targeted query without previous knowledge of the SPARQL language or the structure of the relevant biomedical ontologies.
+Thus, the end user can construct a targeted query without previous knowledge of the SPARQL language or the structure of the relevant ontologies.
 
-When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources in the system as shown in Figure \ref{fig:select}.
-The user may select from this list relevant trusted data sources that they wish to query.
-The user can also dynamically add additional data sources as shown in Figure \ref{fig:add_source}, allowing users to query endpoints of interest that are not included in the default list.
+When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources as shown in Figure \ref{fig:select}.
+The user may select from this list relevant data sources that they wish to query.
+The user can also dynamically add additional data sources as shown in Figure \ref{fig:add_source}, allowing her to query endpoints of interest that are not included in the default list.
 
 A physician who is interested in using the QueryMed system to find relevant clinical trials for 
-her patient can use the Advanced Query Interface to add additional relevant datasources .  For example, she might want to use the clinical trial database LinkedCt, which is 
+her patient can use the Advanced Query Interface to add additional relevant datasources.  For example, she might want to use the clinical trial database LinkedCT, which is 
 not in the default set of endpoints.  So, she selects the “Refine 
@@
 \epsfig{file=images/add_source.jpg,  width=3in}
-\caption{Add New Data Source}
+\caption{The user can dynamically add additional SPARQL endpoints.}
 \label{fig:add_source}
@@
 When a data source is selected, the properties list is automatically populated with all distinct properties available at the selected endpoint.
-If the user wants to restrict her query based on one or more properties available, she can specify values for the desired properties.
-Because the QueryMed system automatically displays a list of properties to the user, users can improve their queries using the underlying structure of the data to obtain more relevant query results.
+The user can specify values for relevant properties to restrict the search space.
 %The properties list is generated by running a SPARQL query of the form given in Figure \ref{fig:SPARQL_props} at the specified data source.
@@
 If the user does not know the specific value for a property, she can instead specify a keywords using the “FILTER” option.
-She also can choose logical operators to connect the various parts of their query.
+She also can choose logical operators to connect the various parts of her query.
 When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
@@
 that takes advantage of the underlying structure of the RDF 
-data in the database. She searches Diseasome for a list of 
-diseases whose class is “Cardiovascular” or for which the associated gene is “ABCA1.” Using the QueryMed advanced 
+data in the database. She searches Diseasome for 
+diseases whose class is ``Cardiovascular" or for which the associated gene is “ABCA1.” Using the QueryMed advanced 
 search interface, the complex SPARQL query corresponding 
@@
 view the query results conveniently displayed in the Results Browser.
-
-\begin{figure*}
-\centering
-\epsfig{file=images/add_props.jpg, width=6in}
-\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
-\label{fig:add_props}
-\end{figure*}
 
@@
 \end{Verbatim}
-\caption{A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.}
+\caption{A complex SPARQL query that was dynamically constructed using the advanced query feature of the QueryMed system.}
 \label{fig:SPARQL_adv}
 \end{figure}
+
+\begin{figure*}
+\centering
+\epsfig{file=images/add_props.jpg, width=6in}
+\caption{The advanced search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
+\label{fig:add_props}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 \label{fig:facet_creator}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=images/results.jpg, width=3.5in}
-\caption{A sample result table}
-\label{fig:results}
 \end{figure}
@@
 \item \textbf{Source Manager:} The Source Manager reads the source list and populates the default query list on the user interface.
-It also keeps track of the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
+It also manages the default endpoints, the currently selected endpoints, and the endpoints that have been dynamically added.
 
@@
 % The user has to specify the data source(s) to query, and can give specific values for one or many of the properties available for the data available to restrict the query space.
-\item \textbf{Additional Services:} In order to make the user interface more user friendly, the following two services perform some ``behind the scenes" work.
 
-\begin{enumerate}
-\item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
-\item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
-\end{enumerate}
+%\item \textbf{Additional Services:} In order to make the user interface more user friendly, the following two services perform some ``behind the scenes" work.
+
+%\begin{enumerate}
+%\item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
+%\item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
+%\end{enumerate}
 
@@
 
-\item The proxy server enables us to avoid cross domain XML-HTTP-Request errors in accessing SPARQL endpoints in various domains.
+\item The proxy server, and not the Javascript client, executes SPARQL queries at the remote endpoints.
+This avoids cross domain XML-HTTP-Request errors in accessing web servers at various domains that are used to host the remote biomedical SPARQL endpoints.
 
@@
 \end{Verbatim}
-\caption{Minimal SPARQL Query Structure in the ``Query All" Case}
+\caption{The structure of the SPARQL query constructed when the user selects ``Query All".}
 \label{fig:min_query}
@@
 When the user runs an advanced query, the proxy server takes as input the data sources to be queried, the properties to be queried, the values for each of the selected properties, and the relationship between the property-value pairs and the original user-specified query term.
-This information is passed to the proxy server as a JSON object.
+This information is also passed to the proxy server as a JSON object.
 
@@
 The basic query interface is designed to be as simple as possible, so that users who have little previous experience running queries on SPARQL endpoints can easily and rapidly find query results.
-The advanced query functionality of our system allows users the flexibility to construct complex SPARQL queries.
-The QueryMed system allows the user to construct queries that take advantage of the underlying structure of the data exposed by the individual data endpoints, without requiring previous knowledge of this structure.
-
+The advanced query functionality of our system allows users the flexibility to intuitively construct complex SPARQL queries.
 \subsubsection{Implementation}
@@
 The difference in running time between \emph{with local data} and \emph{without local data} is approximately three orders of magnitude.
-The network latency and the slow performance of the remote SPARQL endpoints as compared to the local data cache can be accounted for the increase in running time for the queries in the two cases.
+The network latency and the slow performance of the remote SPARQL endpoints as compared to the local data cache explains  the increase in running time for the queries in the two cases.
 We also noted that the running time of the subsequent iterations of the same query is significantly less than the initial query, probably due to browser caching.
-When comparing the query execution time amongst the data sources, we see that the Drugbank takes the longest, probably due to the greater size of the Drugbank data as compared to the other three data sources~\footnote{As of February 15\textsuperscript{th} 2010, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, Sider contains 192,515 triples, and DrugBank contains 765,936 triples) \cite{Diseasome, DailyMed, Sider, Drugbank}}.
+When comparing query execution times across the data sources, we see that the Drugbank takes the longest, probably due to the greater size of the Drugbank dataset as compared to the other three datasets.\footnote{As of February 15\textsuperscript{th} 2010, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, Sider contains 192,515 triples, and DrugBank contains 765,936 triples) \cite{Diseasome, DailyMed, Sider, Drugbank}}
 
@@
 
-Therefore, it is fair to say that bottleneck in the running time of our system is running queries that must retrieve many triples from slow remote SPARQL endpoints.
-By managing our own cache to contain the data most likely to be needed on the proxy server, we were able to dramatically decrease running time.
+Based on the dramatic improvement in running time obtained using caching, the slowest step in executing queries with our system is retrieving triples from slow remote SPARQL endpoints.
+By managing our own cache to contain the data most likely to be needed on the proxy server, we were able to decrease running time considerably.
 
@@
 \end{Verbatim}
-\caption{SPARQL query to retrieve all Properties. This was used to measure the performance of implementing a local cache.}
+\caption{SPARQL query to retrieve all properties. This was used to measure the performance of implementing a local cache.}
 \label{fig:SPARQL_props}
@@
 \epsfig{file=images/chart.jpg, width=3in}
-\caption{Comparison of Query Execution Times (measured in milliseconds) at Various SPARQL Endpoints With and Without Local Data (Cache) (average of three trials)}
+\caption{Comparison of query execution times \emph{(ms)} with and without local data caching. The data points represent the average of three trials.}
 \label{fig:chart}
 \end{figure}
+
+\pagebreak
 
@@
 \hline
-\textbf{DARQ}  & No &  No & Yes & N/A & No & Yes & No  \\
+\textbf{DARQ}  & No &  No & Yes & Yes & No & Yes & No  \\
 \hline
@@
 
-A number of existing tools aim to provide a user-friendly interface for browsing semantic web data, or to allow users to perform federated queries.  Several of these are described below.
+A number of existing tools aim to provide a user-friendly interface for browsing biomedical semantic web data, or to allow users to perform federated queries.  Several of these are described below.
 
@@
 The SMART query tool is a web-based application designed to allow biologists to run SPARQL queries over multiple endpoints.
-Queries to the SMART system are constructed using a descriptive logic query written in the natural-language like Manchester OWL syntax \cite{Battista}.
+Queries to the SMART system are written in the natural-language like Manchester OWL syntax \cite{Battista}.
 Major differences between the QueryMed system and SMART include the ability to construct queries intuitively by specifying keywords for user selected properties, to iteratively refine query results, and to dynamically add additional endpoints in the QueryMed system.
@@
 GoWeb and BioGateway are two additional systems designed for answering queries on biomedical data  \cite{Dietze, Antezana}.
-GoWeb allows users to a hybrid search, running keyword-based queries and then filtering based on ontological concepts.
-However, while GoWeb functions as a search engine that incorporates ontological background knowledge to improve search results, QueryMed is a query translation system.
-BioGateway provides a web interface to query a provided SPARQL endpoint that integrates multiple biomedical resources, but does not allow the user to run queries over multiple endpoints or to dynamically add endpoints.
+GoWeb allows users to perform a hybrid search, running keyword-based queries and then filtering based on ontological concepts.
+However, while GoWeb functions as a search engine that incorporates ontological background knowledge to improve search results, QueryMed is a query translation system that utilizes user input in constructing SPARQL queries.
+BioGateway provides a web interface to query a provided SPARQL endpoint that includes graphs from several biomedical resources.
+However, BioGateway does not allow the user to run queries over multiple endpoints or to dynamically add endpoints.
 
@@
 First of all, in Twinkle, the user is expected to know what is already available at the SPARQL endpoints to write the query.
-But in QueryMed, we only ask for specific keywords of interest, and give the option of restricting the query should the user wish to run a more precise query.
+But in QueryMed, the user can provide input in the form of keywords, and has the option to restrict the query if she wishes to run a more precise query.
 Second, although Twinkle was designed to be a more general purpose system, it only supports a small number of specific SPARQL endpoints, while QueryMed allows the user to dynamically add SPARQL endpoints.
@@
 But it is often useful to draw on multiple web resources in answering a query.
-There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambite} systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
+The DARQ  system \cite{Quilitz} is designed to allow the user to run integrated queries against multiple SPARQL endpoints.
+But it does not offer a graphical user interface to facilitate use by biomedical domain experts who are not familiar with SPARQL query syntax.
 
@@
 Table \ref{comparison} compares selected features of the QueryMed system with other related systems.
-The QueryMed system was unique among the systems that we found in that it allows endpoints to be dynamically added by the user, and provides a hybrid interface that enables the users to both query and browse data.
-Other features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data.
-This feature increases the ease of use of our system relative to other similar systems.
-Furthermore, the Javascript-based user interface of the QueryMed system, implemented using the JQuery library, makes our user interface particularly attractive, easy to interact with, and capable of handling a variety of user input events.
-Another unique feature of our system is the property-based advanced query interface.
-This interface enables users to take advantage of the structure of the underlying ontologies used to represent the data without prior knowledge of the ontology structures.
 
@@
 Our system currently allows only a restricted set of SPARQL queries.
-Supporting additional types of queries and including query optimization functionality could increase both the flexibility and speed of the QueryMed system.
+Supporting additional types of queries and including query optimization functionality could increase both the flexibility and the speed of the QueryMed system.
 By drawing on the expressivity of the SPARQL language and the information contained in the ontologies used to represent the data, it would be possible to extend our system into an intelligent reasoning system on biomedical data which allows physicians to enter sophisticated, complex queries and find relevant biomedical results.
@@
 We also believe that it would be useful to allow exploration of the relationships among multiple data sources containing similar resources.
-One challenge in integrating biomedical data across multiple data sources is that individual data items (i.e, a specific protein or a specific drug) may be represented by distinct URIs at different endpoints.
+One challenge in integrating biomedical data across multiple sources is that individual data items (i.e, a specific protein or a specific drug) may be represented by distinct URIs at different endpoints.
 The diversity of representations of identical data items across different biomedical data sources makes it difficult to automatically combine these items.
-However, there is some cross-referencing between the biomedical data sources that we examined (for example, DailyMed drugs sometimes refer to diseases in the Diseasome SPARQL endpoint by their URI).
+However, there is some cross-referencing between the biomedical data sources that we included by default in our system.
+For example, DailyMed drugs sometimes refer to diseases in the Diseasome data source by their URI.
 It might be useful to provide a representation of the relationships among search results from different sources with the query results.
-Additionally, the system could automatically detect similar data items even if they are identified by distinct URIs, and group similar items from distinct data sources together in the query results by natural language processing.
-Another useful feature might be to allow the user to view the relationships among distinct URIs (as in the relfinder system \cite{Heim} , which displays possible paths through the RDF graph between distinct resources).
+Additionally, the system could automatically detect similar data items even if they are identified by different URIs, and group similar items from distinct data sources together in the query results by natural language processing.
+Another useful feature might be to allow the user to view the relationships among URIs (as in the relfinder system \cite{Heim} , which displays possible paths through the RDF graph between distinct resources).
 
@@
 
-Another feature that we have begun to implement in our system is an autocompletion feature to automatically retrieve the distinct values for any given property.
+We are also working on an autocompletion feature to automatically retrieve the values for any given property.
 This feature will allow the user to automatically see all valid choices for each property, and will make the system easier to use and reduce empty query result sets.
@@
 
-The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
+The QueryMed system  allows endpoints to be dynamically added by the user, and provides a hybrid interface that enables the user to both query and browse data.
+QueryMed also allows the user both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data.
+Furthermore, the Javascript-based user interface of the QueryMed system, implemented using the JQuery UI library, is particularly attractive, easy to interact with, and capable of handling a variety of user input events.
+We believe that developing systems such as such as QueryMed, which makes SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data sources.
+
+\smallskip
+
+The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and the ability to run queries over multiple endpoints.
 Because our system is flexible and easy to use, we believe it will be of use to the biomedical community.
-We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.
+
+\smallskip
 

### 2010-11-17 docs/www2010/www2011poster.tex
shortened paper for www2111

--- 2010-02-12 docs/www2010/paper.tex
@@
 We have developed an open-source SPARQL query builder and result set visualizer for biomedical data, QueryMed, that allows end users to easily construct and run translational medicine queries across multiple data sources.
+\smallskip
 
-QueryMed is flexible enough to allow queries relevant to a wide range of biomedical topics, runs queries across multiple SPARQL endpoints, and is designed to be accessible to users who do not know the structure of the underlying ontologies used in describing the datasets, or the SPARQL query language to query the data.
+QueryMed allows a flexible range of queries relevant to a wide range of biomedical topics and runs queries across multiple SPARQL endpoints.
+The system is accessible for users who are unfamiliar with the SPARQL query language or the structure of the underlying ontologies.
 The system allows users to select the data sources that they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.
 Users can add additional data sources if they are interested in querying endpoints that are not in the default list.
-After retrieval of the initial result set, query results can be filtered to improve their relevance.
-The system also allows the user to exploit the underlying structure of the RDF data to improve query results.
+The system automatically translates the user input into a SPARQL query for each individual endpoint, combines the results, and returns them to the user.
+After retrieval of the initial result set, query results can be improved  by iteratively modifying the original query terms, and by filtering the result list.
+The advanced query functionality of the system allows the user to easily construct complex logical SPARQL queries that exploit the underlying structure of the RDF data.
 \end{abstract}
-
-% A category with only the three required fields
-\category{J.3}{Life and Medical Sciences}{Computer Applications}
-\category{H.3.3}{Information Search and Retrieval}{Information Systems}
 
@@
 
-\section{Introduction}
-
-The quantity of publicly available data in the biomedical domain has dramatically increased over recent years.
-Publicly available biomedical resources include data on drug discovery~\cite{Sharp, Goble}, clinical trials, diseases, disease genes, and phenotypes.
-With the linked open data movement, the semantic web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
-In fact, the biomedical domain is among the early successes of the semantic web due to the rapidity with which the community has made its data available in RDF triple stores \cite{Yip}.
-
-%The Semantic Web Health Care and Life Sciences Interest Group (HCLSIG) has been formed with the purpose of exploring the applications of the semantic web to the biomedical domain \cite{HCLSIG}.
-
-To allow end users to exploit the abundance of useful biomedical data that is currently available in RDF, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow users to run federated queries on multiple SPARQL endpoints.
-There is also a  need for efficient hybrid interfaces that allow both browsing and querying~\cite{Jentzsch}, since many currently available systems are linked data browsers such as the Tabulator~\cite{Tabulator}, which allow a user to navigate the data in an exploratory manner but lack support for filtering and querying the data.
-
-Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple endpoints.
-For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
-She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.
-Although the physician has a single question--``based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?"--there is no single data source that the physician can use to answer this question.
-The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and GO \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
-Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
-
-Since the physician must search many databases in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.
-Also, the physician may not know SPARQL query syntax, the location of the SPARQL endpoints, or the structure of the relevant ontologies.
-She is likely to want an intuitive way to query and to display the query result.
-Developing intuitive ways to query multiple data sources and display results is both an important and a challenging problem.
-Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.
-
-This paper is organized as follows: Section \ref{background} provides background information on the semantic web and its relevance for the biomedical domain.
-Section \ref{system} describes our system.
-Section \ref{related} discusses related work and illustrates how QueryMed differs from previous systems.
-Finally, section \ref{conclusion} outlines future work and summarizes the contributions of our system.
-
-\section{Background}
-\label{background}
-
-The semantic web can be viewed as a global database system for the information available on the world wide web.
-Semantic web data is modeled by structured languages such as RDF and OWL, and can be queried using the SPARQL query language.
-The addition of structure to web data allows inferences to be automatically drawn by intelligent agents integrating data from multiple sources~\cite{Berners-Lee1}.
-%Although most web data are not currently available in semantic web formats, the biomedical knowledge domain has been an early success of the semantic web \cite{Yip}.
-
-Many major biological and biomedical data resources, including Gene Ontology, DailyMed, LinkedCT, and Diseasome, are currently available as RDF triplestores.
-Almost all of these data sources are interlinked with each other.
-Integrating biomedical data across multiple data sources and automatically extracting specific  knowledge from web resources are crucial tasks for physicians and biologists.
-These semantic web resources represent valuable repositories of information that can be automatically mined for applications that require biological knowledge.
-
-Although many valuable resources in the biomedical domain are available in RDF, there are a number of challenges that must be addressed in order to make such resources accessible to physicians, patients, and life scientists.
-One challenge is constructing systems that allow end users to run intuitive queries on biomedical data.
-Users of biomedical resources are likely to have extensive domain knowledge, but be unfamiliar with the SPARQL query language syntax and with the structure of biomedical ontologies.
-Almost all the SPARQL endpoints available today are offering a generic query interface that requires users to write SPARQL queries.
-Also, no hints about the structure of data available from the SPARQL endpoint is given as the user is constructing the SPARQL query.
-The user has to essentially construct the queries that are needed for her purpose by trial and error.
-This can be a daunting task  especially for someone who is new to semantic web technologies.
-Therefore, it is important to design user-friendly systems that allow users to take advantage of the wealth of structured biological knowledge available on the semantic web.
-Another central challenge is designing systems that permit users to query multiple data sources simultaneously, since relevant biological data are often distributed among many sources \cite{Pasquier}.
-
-\section{Design \& Implementation}
 \label{system}
 
-The QueryMed system allows users to easily query multiple biomedical data sources.
-Queries can be run against a default list of datasources, or against a set of user-selected SPARQL endpoint.
-The end user can easily input additional endpoints in order to utilize resources that are not included in the default list.
-The system automatically translates the user input into a SPARQL query for each individual endpoint, combines the results, and returns them to the user.
-The user can choose to refine the query by iteratively modifying the original query terms, and by filtering the result list.
-The advanced query functionality of the QueryMed system allows the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the ontologies.
-The simple user interface of the QueryMed system is designed to be intuitive for users with no knowledge of the SPARQL query language.
+\section{QueryMed Overview}
 
-In the following sections we explain the system functionality by first giving an overview of the QueryMed system architecture and the design decisions behind crucial components of the system.
+\smallskip
 
@@
 
-\begin{figure}
-\centering
-\epsfig{file=images/architecture_overview.jpg}
-\caption{QueryMed Architecture Overview}
-\label{fig:arch_overview}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=images/architecture_detail.jpg}
-\caption{QueryMed Architecture Details}
-\label{fig:arch_details}
-\end{figure}
-
-A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_overview}.
+A general overview of QueryMed architecture is shown in Figure \ref{fig:architecture_detail.jpg}.
 The main components of the system are the user interface and the proxy server that takes input from the user and retrieves relevant biomedical data from remote SPARQL endpoints.
@@
 The query results are returned from the remote endpoints, combined by the proxy server, and presented to the user.
-A detailed illustration of the parts of the QueryMed system is shown in Figure \ref{fig:arch_details}.
 
@@
 The QueryMed user interface is designed to be intuitive for the end user, yet flexible enough to permit a broad range of interesting queries.
-The basic query interface allows the user to run simple queries, and is designed for maximal ease of use.
+The basic query interface allows the user to run simple queries, and is designed for maximal ease of use (Figure \ref{fig:initial_ui}).
 The advanced search capabilities allow the user to easily construct complex logical queries that take advantage of the underlying structure of the biomedical ontologies.
 The user interface also allows the user to iteratively refine queries, and displays the query results, which have been retrieved from multiple SPARQL endpoints.
-The main components of the user interface are as follows:
 
-\begin{itemize}
+\begin{figure}
+\centering
+\epsfig{file=images/architecture_detail.jpg}
+\caption{QueryMed Architecture}
+\label{fig:arch_details}
+\end{figure}
 
-\item \textbf{Basic Query Interface:}  In order to make QueryMed easy to use, but also provide a flexible system that is capable of performing a broad variety of biomedically relevant queries, we provide an uncluttered basic search interface with only a  blank search box and two buttons as shown in Figure \ref{fig:initial_ui}.
-The ``Query All" button will perform a keyword-based query over pre-selected data sources.
-The advanced search option ``Refine Query" allows the user to view additional query options that are handled by the Advanced Query Interface.
-
- For example, a physician interested in finding semantic web resources related to coronary artery disease could use “coronary artery disease”  as a search keyword in the input search box and select the "Query All" option.
-The simplicity of the basic query interface allows the physician to rapidly and easily browse semantic web resources of interest.
+\smallskip
 
@@
 
-\item \textbf{Result Browser:} The user can view his query results organized by source in the Result Browser.
-The results are presented in a table with pagination.
-The user can choose the number of results viewed at a time, search the columns based on some text value and also sort the columns as shown in Figure \ref{fig:results}.
-This allows the user to perform additional filtering on the query results to display the most relevant results, and is particularly useful for refining queries that return a large number of results.
+\smallskip
 
-For example, after searching for "coronary artery disease" in the basic query interface, the physician will see displayed in the Result Browser a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery
-disease. She can then filter the results using additional search terms. For example, she knows that 
-the route of administration of the drug that she is interested in is injection, so she filters the drug query results on the route of administration field using the query term “injection.”   The additional filtering capabilities of the Result Browser allows her  to either browse a large number of query results, or refine her search to view only the most relevant query results.
-
-\item \textbf{Advanced Query Interface:} The Advanced Query Interface allows the user to add new data sources, and to construct a complex SPARQL query that takes advantage of the structure of the underlying biomedical ontologies.
-Thus, the end user can construct a targeted query without previous knowledge of the SPARQL language or the structure of the relevant biomedical ontologies.
-
-When the ``Refine Query" in the Basic Query Interface is invoked, the user is provided with the default list of data sources in the system as shown in Figure \ref{fig:select}.
-The user may select from this list relevant trusted data sources that they wish to query.
-The user can also dynamically add additional data sources as shown in Figure \ref{fig:add_source}, allowing users to query endpoints of interest that are not included in the default list.
-
-A physician who is interested in using the QueryMed system to find relevant clinical trials for 
-her patient can use the Advanced Query Interface to add additional relevant datasources .  For example, she might want to use the clinical trial database LinkedCt, which is 
-not in the default set of endpoints.  So, she selects the “Refine 
-Query” option to select the endpoints to search, and then the “Add” option 
-to include an additional endpoint. After entering the name 
-and URL of the LinkedCT endpoint, she can able to search 
-for clinical trials for which her patient may be eligible.
-
-\begin{figure}
-\centering
-\epsfig{file=images/select.jpg,  width=2.5in}
-\caption{The user has the option of selecting specific trusted or relevant data sources, or of adding additional data sources to query.}
-\label{fig:select}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=images/add_source.jpg,  width=3in}
-\caption{Add New Data Source}
-\label{fig:add_source}
-\end{figure}
-
-When a data source is selected, the properties list is automatically populated with all distinct properties available at the selected endpoint.
-If the user wants to restrict her query based on one or more properties available, she can specify values for the desired properties.
-Because the QueryMed system automatically displays a list of properties to the user, users can improve their queries using the underlying structure of the data to obtain more relevant query results.
-%The properties list is generated by running a SPARQL query of the form given in Figure \ref{fig:SPARQL_props} at the specified data source.
-
-Once the properties are returned, they will be displayed in the user interface as shown in Figure \ref{fig:add_props}.
-Clicking an individual property link displays a description of the property.
-This allows the user to understand the keywords that will be most useful to search on each specific property.
-
-The user can choose to perform exact queries, or filter results on specific keywords.
-If the user does not know the specific value for a property, she can instead specify a keywords using the “FILTER” option.
-She also can choose logical operators to connect the various parts of their query.
-When “AND” is used it will be appended as a basic triple pattern to the SPARQL query, i.e.
-as a conjunction.
-When “OR” is used, the specified graph pattern is made to disjunct with the rest of the query with the SPARQL UNION operator.
-If no logical operator is specified, “AND” will be used by default.
-The advanced query feature is capable of dynamically constructing complex SPARQL queries, such as the query shown in Figure \ref{fig:SPARQL_adv}.
-
-For example, the physician might want to further refine her search for cardiovascular diseases. She 
-uses the advanced search option to construct a boolean query 
-that takes advantage of the underlying structure of the RDF 
-data in the database. She searches Diseasome for a list of 
-diseases whose class is “Cardiovascular” or for which the associated gene is “ABCA1.” Using the QueryMed advanced 
-search interface, the complex SPARQL query corresponding 
-to her question is automatically constructed (Figure \ref{fig:SPARQL_adv}), and she can 
-view the query results conveniently displayed in the Results Browser.
-
-\begin{figure*}
-\centering
-\epsfig{file=images/add_props.jpg, width=6in}
-\caption{The Advanced Search feature allows the user to perform exact or pattern-matching queries connected by user-specified logical operators over specific properties in given resources, taking advantage of the structure in the RDF data.}
-\label{fig:add_props}
-\end{figure*}
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-SELECT distinct ?disease WHERE { 
-{?x <http://www.w3.org/2000/01/rdf-schema#label> 
-?disease 
-FILTER regex(?disease, 
-"coronary artery disease", "i"). 
-?x <http://www4.wiwiss.fu-berlin.de/diseasome/
-resource/diseasome/class> 
-<http://www4.wiwiss.fu-berlin.de/diseasome/
-resource/diseaseClass/Cardiovascular>} 
-UNION { ?x <http://www4.wiwiss.fu-berlin.de/
-diseasome/resource/diseasome/associatedGene>
- <http://www4.wiwiss.fu-berlin.de/diseasome/
- resource/genes/ABCA1>.}
-}
-\end{Verbatim}
-\caption{A complex SPARQL query that takes advantage of the underlying data structure can be dynamically constructed using the advanced query feature of the QueryMed system.}
-\label{fig:SPARQL_adv}
-\end{figure}
-
-\item \textbf{Dynamic Facet Creator/Filter:} The Dynamic Facet Creator/Filter allows the user to select a  set of  data sources, load the properties available at these sources, and dynamically construct a complex SPARQL query connected by logical operators to take advantage of the structure of the data at each endpoint (Figure  \ref{fig:facet_creator}).
-The interface allows displaying of many property lists from different sources simultaneously (grouped by source).
-Therefore, only information relevant to the endpoint the user is currently examining will be visible at any given time.
-
-\begin{figure}
-\centering
-\epsfig{file=images/facet_creator.jpg, width=3in}
-\caption{Dynamic Facet Creator and Filter}
-\label{fig:facet_creator}
-\end{figure}
-
-\begin{figure}
-\centering
-\epsfig{file=images/results.jpg, width=3.5in}
-\caption{A sample result table}
-\label{fig:results}
-\end{figure}
-
-\end{itemize}
+\begin{table*}
+\begin{center}
+\begin{tabular}{| p{1.8cm} | p{2cm} | p{1.8cm} | p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm} |}
+\hline
+ & \textbf{Hybrid Interface (Combines Querying \& Browsing)} & \textbf{Provides Local Caching} & \textbf{Queries Multiple Sources} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
+\hline
+\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes & Yes & Yes  \\
+\hline
+\textbf{SMART}  & No & Yes & Yes & No & No & Yes & Yes  \\
+\hline
+\textbf{DARQ}  & No &  No & Yes & N/A & No & Yes & No  \\
+\hline
+\textbf{GoWeb}   & Yes & No  & Yes & No & Yes & No & Yes  \\
+\hline
+\textbf{BioGateway}   & Yes & Yes & Yes & No & No & Yes & Yes  \\
+\hline
+\textbf{Twinkle}   & No & No & Yes & No & No & Yes & Yes  \\
+\hline
+\end{tabular}
+\caption{Comparison of selected features of the QueryMed system with other related systems.}
+\label{comparison}
+\end{center}
+\end{table*}
 
@@
 The Translator obtains the parameters to construct the query from the input the user specifies in the Query Interface, and dynamically constructs a SPARQL query based on the user input.
-The Translator relies on two services:
-
-\begin{enumerate}
-
-\item \textbf{Run Query:} The purpose of this service is to execute SPARQL queries generated by the Translator, and return  the query results as a JSON object.
-
-%\begin{figure}
-%\begin{Verbatim}[frame=single]
-%{
-%   "bindings":
-%  [
-%        {"source" : "source_label",
-%         "uri" : "http://source.com/sparql.",
- %    	"vars" : ["variable1","variable2"],
-  %   	"count":2,
-   %  	"results" : 
-   %  	 [
-  %   	    {"variable1": "Value 1",
-%	      "variable1": "Value 2"},
- %    	    {"variable1": "Another Value 1",
-  %             "variable1": "Another Value 2"},
- %  	    ...	  		
-  %   	 ]
-  %   	}
-  %   	,
-%	{ //Another Source and Result Description
-%	...
-%	},
-%	...
- %    ]
-% }
-%\end{Verbatim}
-%\caption{Structure of the JSON ``result" object}
-%\label{fig:JSON_runquery}
-%\end{figure}
-
-\item \textbf{Advanced Queries:} This service takes as input list of sources, properties, query terms, and logical operators, which are passed to the proxy server as a JSON object.
-
-%\begin{figure}
-%\begin{Verbatim}[frame=single]
-%{
-% "http://some-source/sparql": 
-%  {
-%     "property1": {"value1":  "logical operator"},
-%     "property2": {"value 2": "logical operator"},
-%     "property2": {"value 3": "logical operator"},
-%     ...
-%  }
-%}
-%   //Other sources ...
-
-%\end{Verbatim}
-%\caption{Structure of the JSON ``query" object}
-%\caption{JSON constructed to send the structure of a SPARQL query for an advanced query with logical operators}
-%\label{fig:JSON_advquery}
-%\end{figure}
-
-\end{enumerate}
-
 % The user has to specify the data source(s) to query, and can give specific values for one or many of the properties available for the data available to restrict the query space.
-\item \textbf{Additional Services:} In order to make the user interface more user friendly, the following two services perform some ``behind the scenes" work.
-
-\begin{enumerate}
-\item \textbf{Get Properties:} This service takes as input an individual source, and returns an array of all properties for that source.
-\item \textbf{Get Property Values:} This service takes as input an individual source and a specific property.  It returns a list of the possible values that this property can take.
-\end{enumerate}
 
 \end{itemize}
-
-\subsection{Design Decisions}
-
-\subsubsection{Source list}
-
-Because the set of default endpoints is stored on the proxy server, the set of resources available by default to the user can easily be updated.
-Since useful biomedical resources are rapidly being developed and made available as RDF triple stores, the ease of updating the resource list ensures that the system can easily be brought up to date.
-In fact, the QueryMed system could easily be adapted to perform queries outside the biomedical domain by modifying the list of input data sources.
-
-\subsubsection{Proxy Server}
-
-While an entirely client side application is possible, we chose to have a proxy server perform the SPARQL query execution and caching. This design is advantageous for several reasons:
-
-\begin{enumerate}
-
-\item Efficient cache management: The result set from running an unrestricted query can include millions of results.
-It may be infeasible to keep unfiltered query results in browser memory.
-A typical memory footprint for a browser (for e.g.
-Firefox) is usually between 20MB and 100MB.
-A poorly constructed query can result in gigabytes worth of triples returned, causing the browser to crash.
-The proxy server can cache results from initial query execution, so that only a filtered result set is subsequently returned to the client.
-
-\item The proxy server enables us to avoid cross domain XML-HTTP-Request errors in accessing SPARQL endpoints in various domains.
-
-\end{enumerate}
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-SELECT ?projection_1 ?projection_2 ... 
-	?projection_n
-WHERE {?x source:property ? projection 
-	FILTER regex(? projection, '" + 
-	input +"', 'i)
-}
-... //Other property filers are to follow
-\end{Verbatim}
-\caption{Minimal SPARQL Query Structure in the ``Query All" Case}
-\label{fig:min_query}
-\end{figure}
-
-\subsubsection{Data Structures}
-
-The parameters required for constructing the SPARQL queries are sent from the user interface to the proxy server.
-In the general ``Query All" case, the system will take the user-specified query term as text-box input, filter on all the triples available at the SPARQL endpoint to select only those that contain the keyword, and display the results.
-The generated SPARQL query will be of the form illustrated in Figure \ref{fig:min_query}.
-The word ``input" in the figure represents the keyword specified by the user.
-
-When the results are returned, each result set is structured as a JSON object.
-This object identifies the endpoint where the query was executed, the URI of the endpoint, the query variables,  how many results are returned and the result set.
-
-When the user runs an advanced query, the proxy server takes as input the data sources to be queried, the properties to be queried, the values for each of the selected properties, and the relationship between the property-value pairs and the original user-specified query term.
-This information is passed to the proxy server as a JSON object.
-
-\subsubsection{Query Interface}
-
-The basic query interface is designed to be as simple as possible, so that users who have little previous experience running queries on SPARQL endpoints can easily and rapidly find query results.
-The advanced query functionality of our system allows users the flexibility to construct complex SPARQL queries.
-The QueryMed system allows the user to construct queries that take advantage of the underlying structure of the data exposed by the individual data endpoints, without requiring previous knowledge of this structure.
 
@@
 
-%A video illustrating this sample use case can be found at: \\
-http://www.youtube.com/watch?v=JCr7vryS9Vg
-
-\subsection{Performance}
-
-The slowest step in running queries using the QueryMed system is populating the property values for each selected endpoint.
-We compared the times required to load properties from several endpoints, by running the query illustrated in Figure \ref{fig:SPARQL_props}.
-Timing data are shown in Table \ref{timing}.
-
-\begin{figure}
-\begin{Verbatim}[frame=single]
-SELECT DISTINCT ?property
-WHERE { [] ?property [] }
-ORDER BY ?property
-\end{Verbatim}
-\caption{SPARQL query to retrieve all properties of a given data source}
-\label{fig:SPARQL_props}
-\end{figure}
-
-For all selected endpoints, the property values take longer to load on the initial query than in subsequent iterations, probably due to browser caching.
-The running time to acquire the list of property values for the Drugbank endpoint is much larger than the time for the other two endpoints, probably due to the greater size of the Drugbank database.
-(As of 12/10/09, Diseasome contains 91,182 triples, DailyMed contains 164,276 triples, and DrugBank contains 765,936 triples).
-
-\begin{table}
-\begin{tabular}{|l||l|l|l|}
-\hline
- & \textbf{Dieseasome} & \textbf{Dailymed} & \textbf{Drugbank} \\
-\hline
-\hline
-\textbf{1st Trial}   & 3.45 s & 1.77 s  & 9.51 s  \\
-\textbf{2nd Trial}  & 1.61 s & 1.57 s & 9.34 s \\
-\textbf{3rd Trial}  & 1.71 s & 1.66 s & 9.06 s \\
-\hline
-\end{tabular}
-\caption{Running times to retrieve all the properties from selected endpoints}
-\label{timing}
-\end{table}
-
 \subsection{QueryMed Resources}
@@
 
-\begin{table*}
-\begin{center}
-\begin{tabular}{| p{1.8cm} | p{2cm} | p{1.8cm} | p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm} |}
-\hline
- & \textbf{Hybrid Interface (Combines Querying \& Browsing)} & \textbf{Provides Local Caching} & \textbf{Queries Multiple Sources} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
-\hline
-\hline
-\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes & Yes & Yes  \\
-\hline
-\textbf{SMART}  & No & Yes & Yes & No & No & Yes & Yes  \\
-\hline
-\textbf{DARQ}  & No &  No & Yes & N/A & No & Yes & No  \\
-\hline
-\textbf{GoWeb}   & Yes & No  & Yes & No & Yes & No & Yes  \\
-\hline
-\textbf{BioGateway}   & Yes & Yes & Yes & No & No & Yes & Yes  \\
-\hline
-\textbf{Twinkle}   & No & No & Yes & No & No & Yes & Yes  \\
-\hline
-\end{tabular}
-\caption{Comparison of selected features of the QueryMed system with other related systems.}
-\label{comparison}
-\end{center}
-\end{table*}
-
-A number of existing tools aim to provide a user-friendly interface for browsing semantic web data, or to allow users to perform federated queries.  Several of these are described below.
-
-The SMART query tool is a web-based application designed to allow biologists to run SPARQL queries over multiple endpoints.
-Queries to the SMART system are constructed using a descriptive logic query written in the natural-language like Manchester OWL syntax \cite{Battista}.
-Major differences between the QueryMed system and SMART include the ability to construct queries intuitively by specifying keywords for user selected properties, to iteratively refine query results, and to dynamically add additional endpoints in the QueryMed system.
-
-GoWeb and BioGateway are two additional systems designed for answering queries on biomedical data  \cite{Dietze, Antezana}.
-GoWeb allows users to a hybrid search, running keyword-based queries and then filtering based on ontological concepts.
-However, while GoWeb functions as a search engine that incorporates ontological background knowledge to improve search results, QueryMed is a query translation system.
-BioGateway provides a web interface to query a provided SPARQL endpoint that integrates multiple biomedical resources, but does not allow the user to run queries over multiple endpoints or to dynamically add endpoints.
-
-Twinkle offers a stand-alone graphical user interface to load and edit SPARQL queries that can be used to query online SPARQL endpoints \cite{Dodds}.
-Our system differs from the Twinkle system in several aspects.
-First of all, in Twinkle, the user is expected to know what is already available at the SPARQL endpoints to write the query.
-But in QueryMed, we only ask for specific keywords of interest, and give the option of restricting the query should the user wish to run a more precise query.
-Second, although Twinkle was designed to be a more general purpose system, it only supports a small number of specific SPARQL endpoints, while QueryMed allows the user to dynamically add SPARQL endpoints.
-
-Most SPARQL query engines are designed to run queries against individual endpoints.
-But it is often useful to draw on multiple web resources in answering a query.
-There are a number of systems, including the DARQ \cite{Quilitz} and CALO query manager \cite{Ambite} systems, designed to allow the user to run integrated queries against multiple SPARQL endpoints.
+A number of existing tools aim to provide a user-friendly interface for browsing semantic web data, or to allow users to perform federated queries.
 
@@
 
-\section{Future Work}
+\section{Conclusion}
 \label{conclusion}
 
-The bottleneck in the running time of our system is running queries that must retrieve many triples from slow remote SPARQL endpoints.
-By caching relevant data locally, the running time could be significantly improved.
-We observed that the QueryMed system performs significantly faster the second time that a query is run on a given resource possibly due to browser caching.
-By managing our own cache to contain the data most likely to be needed on the proxy server, we could reduce running time still further.
-Another approach that could reduce running time still further, especially for users with a slow network connection, would be to create a complementary standalone application that gives the user the option at startup time of loading all the required RDF data.
-Since data is stored locally after the initial startup, using the system in subsequent queries will be rapid after an initial loading phase.
-This approach might be too memory-intensive if the user wishes to run queries over many large triple stores, but might work best in a situation where there is a small or moderate amount of data in the repositories of interest to the user.
+The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
+Because our system is flexible and easy to use, we believe it will be of use to the biomedical community.
+We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.
 
-Our system currently allows only a restricted set of SPARQL queries.
-Supporting additional types of queries and including query optimization functionality could increase both the flexibility and speed of the QueryMed system.
-By drawing on the expressivity of the SPARQL language and the information contained in the ontologies used to represent the data, it would be possible to extend our system into an intelligent reasoning system on biomedical data which allows physicians to enter sophisticated, complex queries and find relevant biomedical results.
-
-We also believe that it would be useful to allow exploration of the relationships among multiple data sources containing similar resources.
-One challenge in integrating biomedical data across multiple data sources is that individual data items (i.e, a specific protein or a specific drug) may be represented by distinct URIs at different endpoints.
-The diversity of representations of identical data items across different biomedical data sources makes it difficult to automatically combine these items.
-However, there is some cross-referencing between the biomedical data sources that we examined (for example, DailyMed drugs sometimes refer to diseases in the Diseasome SPARQL endpoint by their URI).
-It might be useful to provide a representation of the relationships among search results from different sources with the query results.
-Additionally, the system could automatically detect similar data items even if they are identified by distinct URIs, and group similar items from distinct data sources together in the query results by natural language processing.
-Another useful feature might be to allow the user to view the relationships among distinct URIs (as in the relfinder system \cite{Heim} , which displays possible paths through the RDF graph between distinct resources).
-
-Another feature that we have begun to implement in our system is an autocompletion feature to automatically retrieve the distinct values for any given property.
-This feature will allow the user to automatically see all valid choices for each property, and will make the system easier to use and reduce empty query result sets.
-
-Since a major goal of QueryMed is ease of use by physicians, life scientists, and patients, we also plan to perform a user study to understand how effectively  users without knowledge of SPARQL can interact with our system.
-We plan to use data from this study to further refine our system to improve its usability.
-
-\section{Conclusion}
-
-The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
-Because of the unique features of our system, we believe it will be of use to the biomedical community.
-We also believe that systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.
-
-\section{Acknowledgments}
+%\section{Acknowledgments}
 

### 2010-11-19 docs/www2011/sigproc-sp.tex
Used the template given for the WWW2011 posters

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

\begin{document}

\title{QueryMed: An Intuitive Query Builder for Biomedical Data available on the Web}

\numberofauthors{2}
\author{
\alignauthor Oshani Seneviratne\\
       \affaddr{Massachusetts Institute of Technology}\\
       \affaddr{Cambridge, MA}\\
       \affaddr{USA}\\
       \email{oshani@csail.mit.edu}
\alignauthor Rachel Sealfon\\
       \affaddr{Massachusetts Institute of Technology}\\
       \affaddr{Cambridge, MA}\\
       \affaddr{USA}\\
       \email{rsealfon@csail.mit.edu}
}

\maketitle
\begin{abstract}

QueryMed is a SPARQL query builder and result set visualizer for biomedical data, that allows end users to easily construct and run translational medicine queries across multiple data sources. It allows users to select the data sources they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.  User input is translated into a SPARQL query or multiple queries and executed at the relevant SPARQL endpoints for the results.

\end{abstract}

\category{H.4}{Information Systems Applications}{Miscellaneous}
\keywords{Biomedical Ontologies, SPARQL, Query Federation, Query Building, Semantic Web, User Interfaces}

\section{Introduction}

The quantity of publicly available data in the biomedical domain has dramatically increased over recent years. With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.  The biomedical domain is among the early successes of the Semantic Web, due to the rapidity with which the community has made its data available in RDF triple stores~\cite{Yip}.  However, to allow end users to exploit the abundance of biomedical RDF data, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow them to run federated queries across multiple data sources.  There is also a  need for efficient hybrid interfaces that allow both browsing and querying data~\cite{Jentzsch}.

Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple data sources.  For example, a physician may know her patient's personal information, symptoms, current medications, and genotype. She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.  Although the physician has a single question--``based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?"--there is no single data source that the physician can use to answer this question. The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and Gene Ontology \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}. Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.  Since the physician must search many data sources in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources. Also, the physician may not know SPARQL syntax, the location of the SPARQL endpoints (i.e. data sources), or the structure of the ontologies used to describe the data in the endpoints. She is likely to want an intuitive way to query and to display the query results.  Developing intuitive ways to query multiple SPARQL endpoints and to display results is both an important and a challenging problem.  Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.

\section{Overview}

QueryMed allows querying of multiple biomedical data sources very easily.  Queries can be run against a default list of SPARQL endpoints, or against a set of user-selected endpoints.  Additional endpoints can be selected in order to utilize resources that are not included in the default list.  The system automatically translates the user input into a SPARQL query for each individual endpoint, executes the query, combines the results, and returns them to the user.  If the user is not happy with the results, the query can be refined  by iteratively modifying the original terms used in the query, and by filtering the result set.  The advanced query functionality of QueryMed enable the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the ontologies. The simple user interface of QueryMed is designed to be intuitive for users with no knowledge of SPARQL. A general overview of QueryMed architecture is shown in Figure \ref{fig:architecture_detail.jpg}.

%\begin{figure}
%\centering
%\epsfig{file=images/architecture_detail.jpg}
%\caption{QueryMed Architecture}
%\label{fig:arch_details}
%\end{figure}

\section{QueryMed in Action}

Imagine a scenario in which someone, let's say a physician, is interested in finding freely available resources related to coronary artery disease.  She first tries a basic search over all default resources, by entering ``coronary artery disease" in the input search box.   She then sees a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease displayed in a table.  She can then filter the results using additional search terms.  For example, she knows that the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term ``injection."  She then prints the table of results. She is now interested in finding relevant clinical trials for her patient.  However, the clinical trial database LinkedCT is not in the default set of endpoints, so she selects the ``Refine Query" option to choose additional endpoints to search.  She sees a list of default endpoints, and selects the ``Add" option to include an additional endpoint.  After entering the name and URL of the LinkedCT endpoint, she is able to search for clinical trials for which her patient may be eligible.  She is also interested in further refining her search.  She uses the advanced search option to construct a boolean query that takes advantage of the underlying structure of the RDF data in the database.  She searches the data available in the Diseasome endpoint for a list of diseases whose class is ``Cardiovascular" or for which the associated gene is ``ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to her question is automatically constructed, and she can view the query results conveniently displayed in a table.

\section{Related Work}
\label{related}

A number of existing tools aim to provide a user-friendly interface for browsing Semantic Web data, or to allow users to perform federated queries. The SMART query tool~\cite{Battista} is a Web-based application designed to allow biologists to run queries written in Manchester OWL syntax. Unlike in SMART, QueryMed allows constructing queries intuitively by specifying keywords for user selected properties, to iteratively refine query results, and to dynamically add additional endpoints. GoWeb~\cite{Dietze} allows users to perform a hybrid search, running keyword-based queries and then filtering based on ontological concepts. However, GoWeb functions as a search engine that has in-built ontological background knowledge to improve search results, whereas QueryMed utilizes user input in constructing SPARQL queries without any ontological backing.  BioGatewa~\cite{Antezana} provides a Web interface to query a provided single SPARQL endpoint that includes graphs from several biomedical resources. BioGateway does not facilitate dynamically addition of endpoints like in QueryMed. Another, query builder, Twinkle~\cite{Dodds}, offers a stand-alone graphical user interface to load and edit SPARQL queries. Our system differs from the Twinkle system in several aspects. First of all, in Twinkle, the user is expected to know what is already available at the SPARQL endpoints to write the query. But in QueryMed, the user can provide input in the form of keywords, and has the option to restrict the query if she wishes to run a more precise query. Second, although Twinkle was designed to be a more general purpose system, it only supports a small number of specific SPARQL endpoints, while QueryMed allows the user to dynamically add SPARQL endpoints. Most SPARQL query engines are designed to run queries against individual endpoints. But it is often useful to draw on multiple web resources in answering a query. The DARQ  system \cite{Quilitz} is designed to allow the user to run integrated queries against multiple SPARQL endpoints. But it does not offer a graphical user interface to facilitate use by biomedical domain experts who are not familiar with SPARQL query syntax.

%\begin{table*}
%\begin{center}
%\begin{tabular}{| p{1.8cm} | p{2cm} | p{1.8cm} | p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm} |}
%\hline
% & \textbf{Hybrid Interface (Combines Querying \& Browsing)} & \textbf{Provides Local Caching} & \textbf{Queries Multiple Sources} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
%\hline
%\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes & Yes & Yes  \\
%\hline
%\textbf{SMART~\cite{Battista}}  & No & Yes & Yes & No & No & Yes & Yes  \\
%\hline
%\textbf{DARQ~\cite{Quilitz}}  & No &  No & Yes & Yes & No & Yes & No  \\
%\hline
%\textbf{GoWeb~\cite{Dietze}}   & Yes & No  & Yes & No & Yes & No & Yes  \\
%\hline
%\textbf{BioGateway~\cite{Antezana}}   & Yes & Yes & Yes & No & No & Yes & Yes  \\
%\hline
%\textbf{Twinkle~\cite{Dodds}}   & No & No & Yes & No & No & Yes & Yes  \\
%\hline
%\end{tabular}
%\caption{Comparison of selected features of the QueryMed system with other related systems.}
%\label{comparison}
%\end{center}
%\end{table*}

\section{Conclusion}
\label{conclusion}

One of the main goals of the system was to make it accessible for users who are unfamiliar with SPARQL or the structure of the underlying ontologies.
The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints. Because our system is flexible and easy to use, we believe it will be of use to the biomedical community.  We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.

\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
\end{document}

### 2010-11-19 trunk/docs/www2011/sigproc-sp.tex
fitted stuff in 2 pages

--- 2010-11-19 docs/www2011/sigproc-sp.tex
@@
 \documentclass{acm_proc_article-sp}
-
+\usepackage{url}
 \begin{document}
 
-\title{QueryMed: An Intuitive Query Builder for Biomedical Data available on the Web}
+\title{QueryMed: For Querying Biomedical Data on the Web\titlenote{http://code.google.com/p/querymed}}
 
@@
 
-QueryMed is a SPARQL query builder and result set visualizer for biomedical data, that allows end users to easily construct and run translational medicine queries across multiple data sources.
+QueryMed is a query builder and result set visualizer for biomedical data, that allows end-users to easily construct and run translational medicine queries across multiple data sources.
 It allows users to select the data sources they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.
-User input is translated into a SPARQL query or multiple queries and executed at the relevant SPARQL endpoints for the results.
+User input is translated into SPARQL~\cite{SPARQL} queries and executed at the relevant endpoints.
+The results are presented in an intuitive user interface that allows query refinement and filtering.
 
@@
 
-\category{H.4}{Information Systems Applications}{Miscellaneous}
+%\category{H.4}{Information Systems Applications}{Miscellaneous}
 \keywords{Biomedical Ontologies, SPARQL, Query Federation, Query Building, Semantic Web, User Interfaces}
@@
 
-The quantity of publicly available data in the biomedical domain has dramatically increased over recent years.
-With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
-The biomedical domain is among the early successes of the Semantic Web, due to the rapidity with which the community has made its data available in RDF triple stores~\cite{Yip}.
-However, to allow end users to exploit the abundance of biomedical RDF data, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow them to run federated queries across multiple data sources.
-There is also a  need for efficient hybrid interfaces that allow both browsing and querying data~\cite{Jentzsch}.
-
-Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple data sources.
+Answering many medically and biologically relevant questions with information available on the Web requires searching, filtering, and combining information from multiple data sources.
 For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
@@
 The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and Gene Ontology \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
-Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
+The quantity of publicly available data in the biomedical domain has dramatically increased over the recent years.
+With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
+The biomedical domain is in fact among the early successes of the Semantic Web, due to the rapidity with which the community has made its data available in RDF triple stores~\cite{Yip}.
+However, to exploit the abundance of this data on the Web, there is a need for easy-to-use systems that do not require the end-user to have knowledge of the underlying  structure of the data or the ontologies used in describing the data.
+These systems should also support the queries to be run across multiple data sources.
+There is also a  need for efficient hybrid interfaces that allow both browsing and querying of data~\cite{Jentzsch}.
+
+%Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
 Since the physician must search many data sources in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.
@@
 QueryMed allows querying of multiple biomedical data sources very easily.
-Queries can be run against a default list of SPARQL endpoints, or against a set of user-selected endpoints.
-Additional endpoints can be selected in order to utilize resources that are not included in the default list.
+Queries can be run against a default list of SPARQL endpoints, or against a set of user-defined endpoints.
 The system automatically translates the user input into a SPARQL query for each individual endpoint, executes the query, combines the results, and returns them to the user.
 If the user is not happy with the results, the query can be refined  by iteratively modifying the original terms used in the query, and by filtering the result set.
-The advanced query functionality of QueryMed enable the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the ontologies.
+The advanced query functionality of QueryMed enable the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the data.
 The simple user interface of QueryMed is designed to be intuitive for users with no knowledge of SPARQL.
-A general overview of QueryMed architecture is shown in Figure \ref{fig:architecture_detail.jpg}.
+A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_details}.
 
-%\begin{figure}
-%\centering
-%\epsfig{file=images/architecture_detail.jpg}
-%\caption{QueryMed Architecture}
-%\label{fig:arch_details}
-%\end{figure}
+\begin{figure}
+\centering
+\epsfig{file=images/architecture_detail.jpg, width=3in}
+\caption{QueryMed Architecture}
+\label{fig:arch_details}
+\end{figure}
 
-\section{QueryMed in Action}
+\subsection{QueryMed in Action}
 
-Imagine a scenario in which someone, let's say a physician, is interested in finding freely available resources related to coronary artery disease.
-She first tries a basic search over all default resources, by entering ``coronary artery disease" in the input search box.
-She then sees a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease displayed in a table.
-She can then filter the results using additional search terms.
-For example, she knows that the route of administration of the drug that she is looking for is injection, so she filters the drug query results on the route of administration field using the query term ``injection."  She then prints the table of results.
-She is now interested in finding relevant clinical trials for her patient.
-However, the clinical trial database LinkedCT is not in the default set of endpoints, so she selects the ``Refine Query" option to choose additional endpoints to search.
-She sees a list of default endpoints, and selects the ``Add" option to include an additional endpoint.
-After entering the name and URL of the LinkedCT endpoint, she is able to search for clinical trials for which her patient may be eligible.
-She is also interested in further refining her search.
-She uses the advanced search option to construct a boolean query that takes advantage of the underlying structure of the RDF data in the database.
-She searches the data available in the Diseasome endpoint for a list of diseases whose class is ``Cardiovascular" or for which the associated gene is ``ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to her question is automatically constructed, and she can view the query results conveniently displayed in a table.
+Imagine a scenario in which a physician called Sam, is interested in finding freely available resources related to coronary artery disease.
+Using QueryMed Sam first tries a basic search over all default resources, by entering ``coronary artery disease" in the search box.
+Sam then sees a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease displayed in a table.
+He can then filter the results using additional search terms.
+For example, he knows that the route of administration of the drug that he is looking for is injection, so he filters the drug query results on the route of administration field using the query term ``injection."  he then prints the table of results.
+he is now interested in finding relevant clinical trials for her patient.
+However, the clinical trial database (LinkedCT~\cite{LinkedCT}) is not in the default set of endpoints, so he selects the ``Refine Query" option to choose additional endpoints to search.
+he sees a list of default endpoints, and selects the ``Add" option to include an additional endpoint.
+After entering the name and URL of the LinkedCT endpoint, he is able to search for clinical trials for which her patient may be eligible.
+He is also interested in further refining her search, so he uses the advanced search option to construct a query that takes advantage of the underlying structure of the RDF data in the database.
+He searches the data available in the Diseasome endpoint for a list of diseases whose class is ``Cardiovascular" or for which the associated gene is ``ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to his question is automatically constructed, and he can view the query results conveniently displayed in a table.
+A demo describing this scenario is available at \url{http://dig.csail.mit.edu/2010/Papers/www-ws-colab-science/videos/querymed-demo.mov}.
 
@@
 GoWeb~\cite{Dietze} allows users to perform a hybrid search, running keyword-based queries and then filtering based on ontological concepts.
-However, GoWeb functions as a search engine that has in-built ontological background knowledge to improve search results, whereas QueryMed utilizes user input in constructing SPARQL queries without any ontological backing.
-BioGatewa~\cite{Antezana} provides a Web interface to query a provided single SPARQL endpoint that includes graphs from several biomedical resources.
-BioGateway does not facilitate dynamically addition of endpoints like in QueryMed.
+However, GoWeb functions as a search engine that has in-built ontological background knowledge to improve the search results, whereas QueryMed utilizes user input in constructing SPARQL queries without any ontological backing.
+BioGateway~\cite{Antezana} provides a Web interface to query a provided single SPARQL endpoint that includes graphs from several biomedical resources.
+However, BioGateway does not facilitate dynamic addition of endpoints like in QueryMed.
 Another, query builder, Twinkle~\cite{Dodds}, offers a stand-alone graphical user interface to load and edit SPARQL queries.
-Our system differs from the Twinkle system in several aspects.
-First of all, in Twinkle, the user is expected to know what is already available at the SPARQL endpoints to write the query.
-But in QueryMed, the user can provide input in the form of keywords, and has the option to restrict the query if she wishes to run a more precise query.
-Second, although Twinkle was designed to be a more general purpose system, it only supports a small number of specific SPARQL endpoints, while QueryMed allows the user to dynamically add SPARQL endpoints.
+Our system differs from the Twinkle, because in QueryMed, the user can provide input in the form of keywords, and has the option to restrict the query if she wishes to run a more precise query.
 Most SPARQL query engines are designed to run queries against individual endpoints.
@@
 The DARQ  system \cite{Quilitz} is designed to allow the user to run integrated queries against multiple SPARQL endpoints.
-But it does not offer a graphical user interface to facilitate use by biomedical domain experts who are not familiar with SPARQL query syntax.
-
-%\begin{table*}
-%\begin{center}
-%\begin{tabular}{| p{1.8cm} | p{2cm} | p{1.8cm} | p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm}| p{1.8cm} |}
-%\hline
-% & \textbf{Hybrid Interface (Combines Querying \& Browsing)} & \textbf{Provides Local Caching} & \textbf{Queries Multiple Sources} & \textbf{Dynamic Addition of Sources} & \textbf{Allows Keyword Queries} & \textbf{Open Source} & \textbf{GUI} \\
-%\hline
-%\textbf{QueryMed}  & Yes & Yes & Yes & Yes & Yes & Yes & Yes  \\
-%\hline
-%\textbf{SMART~\cite{Battista}}  & No & Yes & Yes & No & No & Yes & Yes  \\
-%\hline
-%\textbf{DARQ~\cite{Quilitz}}  & No &  No & Yes & Yes & No & Yes & No  \\
-%\hline
-%\textbf{GoWeb~\cite{Dietze}}   & Yes & No  & Yes & No & Yes & No & Yes  \\
-%\hline
-%\textbf{BioGateway~\cite{Antezana}}   & Yes & Yes & Yes & No & No & Yes & Yes  \\
-%\hline
-%\textbf{Twinkle~\cite{Dodds}}   & No & No & Yes & No & No & Yes & Yes  \\
-%\hline
-%\end{tabular}
-%\caption{Comparison of selected features of the QueryMed system with other related systems.}
-%\label{comparison}
-%\end{center}
-%\end{table*}
+But it does not offer a graphical user interface as in QueryMed to facilitate use by biomedical domain experts who are not familiar with SPARQL query syntax.
 
@@
 
-One of the main goals of the system was to make it accessible for users who are unfamiliar with SPARQL or the structure of the underlying ontologies.
-The main contributions of our system are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
-Because our system is flexible and easy to use, we believe it will be of use to the biomedical community.
-We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end users, will entice more people to expose their data as linked open data.
+The main contributions of QueryMed are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
+Because it is flexible and easy to use, we believe that it will be of immense value to the biomedical community.
+We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end-users, will entice more people to expose their data as linked open data thus promoting the growth of the LODD cloud.
 

### 2010-11-20 docs/www2011/sigproc-sp.tex
made some other changes

@@
 \documentclass{acm_proc_article-sp}
-
+\usepackage{url}
 \begin{document}
 
-\title{QueryMed: An Intuitive Query Builder for Biomedical Data available on the Web}
+\title{QueryMed: For Querying Biomedical Data on the Web\titlenote{Interested readers are advised to visit http://code.google.com/p/querymed to learn more about QueryMed.}}
 
@@
 
-The quantity of publicly available data in the biomedical domain has dramatically increased over recent years.
+Answering many medically and biologically relevant questions with information available on the Web requires searching, filtering, and combining information from multiple data sources.
+For example, a physician who knows a patient's symptoms, current medications, and genotype may wish to determine the treatment plan and identify clinical trials for which the patient is eligible for.
+However, there is no single data source that the physician can use to figure out the answers.
+The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and Gene Ontology \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
+The quantity of publicly available data in the biomedical domain has dramatically increased over the recent years.
 With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
-The biomedical domain is among the early successes of the Semantic Web, due to the rapidity with which the community has made its data available in RDF triple stores~\cite{Yip}.
-However, to allow end users to exploit the abundance of biomedical RDF data, there is a need for easy-to use systems that do not require the end user to have knowledge of the underlying  structure of the data, and that also allow them to run federated queries across multiple data sources.
-There is also a  need for efficient hybrid interfaces that allow both browsing and querying data~\cite{Jentzsch}.
+The biomedical domain is in fact among the early successes of the Semantic Web, due to the rapidity with which the community has made its data available in RDF triple stores~\cite{Yip}.
+However, to exploit the abundance of this data on the Web, there is a need for easy-to-use systems that do not require the end-user to have knowledge of the underlying  structure of the data or the ontologies used in describing the data.
+These systems should also support the queries to be run across multiple data sources.
+There is also a  need for efficient hybrid interfaces that allow both browsing and querying of data~\cite{Jentzsch}.
 
-Answering many medically and biologically relevant questions requires searching, filtering, and combining information from multiple data sources.
-For example, a physician may know her patient's personal information, symptoms, current medications, and genotype.
-She may wish to determine the patient's treatment plan and identify clinical trials for which the patient is eligible.
-Although the physician has a single question--``based on the information I have about this patient, what is the best treatment plan and set of clinical trials available?"--there is no single data source that the physician can use to answer this question.
-The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and Gene Ontology \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
-Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
+%Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
 Since the physician must search many data sources in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.

### 2010-11-20 docs/www2011/sigproc-sp.tex
more edits to www2011 poster

@@
 
-QueryMed is a query builder and result set visualizer for biomedical data, that allows end-users to easily construct and run translational medicine queries across multiple data sources.
-It allows users to select the data sources they wish to use, drawing on their specialized domain knowledge to decide the most appropriate data sources to query.
-User input is translated into SPARQL~\cite{SPARQL} queries and executed at the relevant endpoints.
-The results are presented in an intuitive user interface that allows query refinement and filtering.
+QueryMed is a query builder and result set visualizer for biomedical data that allows users to easily construct and run translational medicine queries across multiple data sources.
+The system is accessible for users who are unfamiliar with the SPARQL query language
+or the structure of the underlying ontologies.
+It permits users to retrieve data from an information-rich default set of resources or to draw on their specialized domain knowledge to determine the most appropriate data sources to query.
+User input is provided through an intuitive interface, translated into SPARQL~\cite{SPARQL} queries, and executed at the relevant endpoints.
+The results are presented in an accessible interface that allows query refinement and filtering.
 
@@
 
-Answering many medically and biologically relevant questions with information available on the Web requires searching, filtering, and combining information from multiple data sources.
-For example, a physician who knows a patient's symptoms, current medications, and genotype may wish to determine the treatment plan and identify clinical trials for which the patient is eligible for.
-However, there is no single data source that the physician can use to figure out the answers.
-The information that the physician needs must be gathered from numerous data sources such as Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome,  and Gene Ontology \cite{Pubmed, DailyMed, Drugbank, LinkedCT, Diseasome, GO}.
-The quantity of publicly available data in the biomedical domain has dramatically increased over the recent years.
-With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information resources to RDF~\cite{LinkingData}.
-The biomedical domain is in fact among the early successes of the Semantic Web, due to the rapidity with which the community has made its data available in RDF triple stores~\cite{Yip}.
-However, to exploit the abundance of this data on the Web, there is a need for easy-to-use systems that do not require the end-user to have knowledge of the underlying  structure of the data or the ontologies used in describing the data.
-These systems should also support the queries to be run across multiple data sources.
-There is also a  need for efficient hybrid interfaces that allow both browsing and querying of data~\cite{Jentzsch}.
+The quantity of publicly available data in the biomedical domain has dramatically increased in recent years.
+With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information datasources to Resource Description Framework (RDF) triplestores~\cite{LinkingData, Yip}.
+However, to exploit the abundance of biomedical data on the Semantic Web, there is a need for easy-to-use systems that do not require the end user to have knowledge of the underlying  structure of the data or of the ontologies used in describing the data.
+These systems should support queries that run across multiple data sources.
+There is also a  need for efficient hybrid interfaces that allow browsing data resources as well as performing queries~\cite{Jentzsch}.
 
@@
 
-QueryMed allows querying of multiple biomedical data sources very easily.
+The QueryMed system is designed to allow users without specialized knowledge of database query languages to browse and query the extensive biomedical data resources available in RDF triplestores.
+The simple user interface of QueryMed is intuitive for users with no knowledge of SPARQL and allows the user to run queries across multiple biomedical datasources.
 Queries can be run against a default list of SPARQL endpoints, or against a set of user-defined endpoints.
 The system automatically translates the user input into a SPARQL query for each individual endpoint, executes the query, combines the results, and returns them to the user.
-If the user is not happy with the results, the query can be refined  by iteratively modifying the original terms used in the query, and by filtering the result set.
-The advanced query functionality of QueryMed enable the user to easily construct complex logical SPARQL queries that take advantage of the underlying structure of the data.
-The simple user interface of QueryMed is designed to be intuitive for users with no knowledge of SPARQL.
+The user can choose to refine the query by iteratively modifying the original search terms and by filtering the result set.
+The system's advanced query functionality enables the user to construct complex logical SPARQL queries that take advantage of the underlying structure of the data.
 A general overview of QueryMed architecture is shown in Figure \ref{fig:arch_details}.
@@
 
-\subsection{QueryMed in Action}
+\subsection{Sample use case}
 
-Imagine a scenario in which a physician called Sam, is interested in finding freely available resources related to coronary artery disease.
-Using QueryMed Sam first tries a basic search over all default resources, by entering ``coronary artery disease" in the search box.
-Sam then sees a list of disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease displayed in a table.
+Imagine a scenario in which a physician, Sam, is interested in locating freely available resources related to coronary artery disease.
+Using QueryMed, Sam first runs a basic search by entering ``coronary artery disease" in the search box.
+He sees a table displaying disease names in the Diseasome database and drugs in DailyMed and Drugbank that relate to coronary artery disease.
 He can then filter the results using additional search terms.
-For example, he knows that the route of administration of the drug that he is looking for is injection, so he filters the drug query results on the route of administration field using the query term ``injection."  he then prints the table of results.
-he is now interested in finding relevant clinical trials for her patient.
-However, the clinical trial database (LinkedCT~\cite{LinkedCT}) is not in the default set of endpoints, so he selects the ``Refine Query" option to choose additional endpoints to search.
-he sees a list of default endpoints, and selects the ``Add" option to include an additional endpoint.
-After entering the name and URL of the LinkedCT endpoint, he is able to search for clinical trials for which her patient may be eligible.
-He is also interested in further refining her search, so he uses the advanced search option to construct a query that takes advantage of the underlying structure of the RDF data in the database.
-He searches the data available in the Diseasome endpoint for a list of diseases whose class is ``Cardiovascular" or for which the associated gene is ``ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to his question is automatically constructed, and he can view the query results conveniently displayed in a table.
-A demo describing this scenario is available at \url{http://dig.csail.mit.edu/2010/Papers/www-ws-colab-science/videos/querymed-demo.mov}.
+For example, he knows that the route of administration of the drug that he is looking for is injection, so he filters the drug query results on the route of administration field using the query term ``injection."  He is now interested in finding relevant clinical trials for his patient.
+The clinical trial database (LinkedCT~\cite{LinkedCT}) is not in the default set of endpoints, so he selects the ``Refine Query" option to choose additional endpoints to search.
+He sees a list of default endpoints, and selects the ``Add" option to include an additional endpoint.
+After entering the name and URL of the LinkedCT endpoint, he is able to search for clinical trials for which his patient may be eligible.
+He is also interested in further refining the search, so he uses the advanced search option to search the data available in the Diseasome endpoint for a list of diseases whose class is ``Cardiovascular" or for which the associated gene is ``ABCA1."  Using the QueryMed advanced search interface, the complex SPARQL query corresponding to his question is automatically constructed, and he can view the query results conveniently displayed in a table.
+A demo illustrating this scenario is available at \url{http://dig.csail.mit.edu/2010/Papers/www-ws-colab-science/videos/querymed-demo.mov}.
 
@@
 The SMART query tool~\cite{Battista} is a Web-based application designed to allow biologists to run queries written in Manchester OWL syntax.
-Unlike in SMART, QueryMed allows constructing queries intuitively by specifying keywords for user selected properties, to iteratively refine query results, and to dynamically add additional endpoints.
 GoWeb~\cite{Dietze} allows users to perform a hybrid search, running keyword-based queries and then filtering based on ontological concepts.
-However, GoWeb functions as a search engine that has in-built ontological background knowledge to improve the search results, whereas QueryMed utilizes user input in constructing SPARQL queries without any ontological backing.
-BioGateway~\cite{Antezana} provides a Web interface to query a provided single SPARQL endpoint that includes graphs from several biomedical resources.
-However, BioGateway does not facilitate dynamic addition of endpoints like in QueryMed.
-Another, query builder, Twinkle~\cite{Dodds}, offers a stand-alone graphical user interface to load and edit SPARQL queries.
-Our system differs from the Twinkle, because in QueryMed, the user can provide input in the form of keywords, and has the option to restrict the query if she wishes to run a more precise query.
-Most SPARQL query engines are designed to run queries against individual endpoints.
-But it is often useful to draw on multiple web resources in answering a query.
+BioGateway~\cite{Antezana} provides a Web interface to query a single provided SPARQL endpoint that includes graphs from several biomedical resources.
+Another query-building tool, Twinkle~\cite{Dodds}, offers a stand-alone graphical user interface to load and edit SPARQL queries.
 The DARQ  system \cite{Quilitz} is designed to allow the user to run integrated queries against multiple SPARQL endpoints.
-But it does not offer a graphical user interface as in QueryMed to facilitate use by biomedical domain experts who are not familiar with SPARQL query syntax.
+But it does not offer a graphical user interface to facilitate use by biomedical domain experts who are not familiar with SPARQL query syntax.
+
+Features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data, a hybrid interface that enables the user to both query and browse data, and a graphical user interface permitting the dynamic addition of endpoints.
+Furthermore, the Javascript-based user interface of the QueryMed system, implemented using the JQuery library, makes our user interface particularly attractive, easy to interact with, and capable of handling a flexible range of user input.
+Another unique feature of our system is the property-based advanced query interface.
+This interface enables users to take advantage of the structures of the underlying ontologies used to represent the data without prior knowledge of these structures.
 
@@
 The main contributions of QueryMed are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
-Because it is flexible and easy to use, we believe that it will be of immense value to the biomedical community.
-We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end-users, will entice more people to expose their data as linked open data thus promoting the growth of the LODD cloud.
+Because the system is flexible and easy to use, we believe that it will be valuable to the biomedical community.
+We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end-users, will entice more people to expose their biomedical data as linked open data, thus promoting the growth of the Linked Open Drug Data cloud.
 

### 2010-11-20 docs/www2011/sigproc-sp.tex
the final changes including the one that we submitted

@@
 QueryMed is a query builder and result set visualizer for biomedical data that allows users to easily construct and run translational medicine queries across multiple data sources.
-The system is accessible for users who are unfamiliar with the SPARQL query language
+The system is accessible for users who are unfamiliar with the SPARQL query language~\cite{SPARQL}
 or the structure of the underlying ontologies.
 It permits users to retrieve data from an information-rich default set of resources or to draw on their specialized domain knowledge to determine the most appropriate data sources to query.
-User input is provided through an intuitive interface, translated into SPARQL~\cite{SPARQL} queries, and executed at the relevant endpoints.
+User input is provided through an intuitive interface, translated into SPARQL queries, and executed at the relevant endpoints.
 The results are presented in an accessible interface that allows query refinement and filtering.
@@
 The quantity of publicly available data in the biomedical domain has dramatically increased in recent years.
-With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information datasources to Resource Description Framework (RDF) triplestores~\cite{LinkingData, Yip}.
+With the linked open data movement, the Semantic Web community has been very proactive in converting these rich information data sources to Resource Description Framework (RDF) triplestores~\cite{LinkingData, Yip}.
 However, to exploit the abundance of biomedical data on the Semantic Web, there is a need for easy-to-use systems that do not require the end user to have knowledge of the underlying  structure of the data or of the ontologies used in describing the data.
-These systems should support queries that run across multiple data sources.
+These systems should support queries that run across multiple datasources.
 There is also a  need for efficient hybrid interfaces that allow browsing data resources as well as performing queries~\cite{Jentzsch}.
 
-%Her question must be broken up into discrete pieces that can be executed individually at one data source at a time.
-Since the physician must search many data sources in order to find an answer to her single question, she requires a system that can automatically run queries over multiple data sources.
-Also, the physician may not know SPARQL syntax, the location of the SPARQL endpoints (i.e.
-data sources), or the structure of the ontologies used to describe the data in the endpoints.
-She is likely to want an intuitive way to query and to display the query results.
-Developing intuitive ways to query multiple SPARQL endpoints and to display results is both an important and a challenging problem.
-Our system, QueryMed, allows users with no knowledge of the SPARQL query language or the structure of the underlying ontologies to easily run queries across multiple SPARQL endpoints.
+\newpage
 
@@
 
-The QueryMed system is designed to allow users without specialized knowledge of database query languages to browse and query the extensive biomedical data resources available in RDF triplestores.
+The QueryMed system is designed to allow users without specialized knowledge of query languages to browse and query the extensive biomedical data resources available in RDF triplestores.
 The simple user interface of QueryMed is intuitive for users with no knowledge of SPARQL and allows the user to run queries across multiple biomedical datasources.
@@
 
-\subsection{Sample use case}
+\newpage
+
+\subsection{Sample Use Case}
 
@@
 Features of the QueryMed system that distinguish it from similar systems include the ability both to perform keyword queries and to construct more advanced queries taking advantage of the structure of the data, a hybrid interface that enables the user to both query and browse data, and a graphical user interface permitting the dynamic addition of endpoints.
-Furthermore, the Javascript-based user interface of the QueryMed system, implemented using the JQuery library, makes our user interface particularly attractive, easy to interact with, and capable of handling a flexible range of user input.
+Furthermore, the QueryMed user interface is attractive, easy to interact with, and capable of handling a flexible range of user input.
 Another unique feature of our system is the property-based advanced query interface.
@@
 
-The main contributions of QueryMed are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified endpoints; and ability to run queries over multiple endpoints.
+The main contributions of QueryMed are: dynamic construction of complex SPARQL queries based on intuitive user input; dynamic addition of user-specified datasources; and ability to run queries over multiple datasources.
 Because the system is flexible and easy to use, we believe that it will be valuable to the biomedical community.
-We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end-users, will entice more people to expose their biomedical data as linked open data, thus promoting the growth of the Linked Open Drug Data cloud.
+We also believe that developing systems such as such as QueryMed, which make SPARQL endpoints  easily accessible to end-users, will entice more people to expose their biomedical data as linked open data, thus promoting the growth of the Linked Open Data cloud.
 
